---
title: "COD_week14_1_MGK_BTE3207"
author: "Minsik Kim"
date: "2023-11-28"
output:
        rmdformats::downcute:
        downcute_theme: "chaos"
code_folding: show
fig_width: 6
fig_height: 6
df_print: paged
editor_options: 
        chunk_output_type: inline
markdown: 
        wrap: 72
---
        
        
        
```{r warning=FALSE, message=FALSE, echo=FALSE, results='hide', setup}
#===============================================================================
#BTC.LineZero.Header.1.1.0
#===============================================================================
#R Markdown environment setup and reporting utility.
#===============================================================================
#RLB.Dependencies:
#   knitr, magrittr, pacman, rio, rmarkdown, rmdformats, tibble, yaml
#===============================================================================
#Input for document parameters, libraries, file paths, and options.
#=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=
knitr::opts_chunk$set(message=FALSE, warning = FALSE)

path_working <- 
        ifelse(sessionInfo()[1]$R.version$platform == "aarch64-apple-darwin20",
               "/Volumes/MacMini Drive/Dropbox (Personal)/Inha/5_Lectures/Advanced biostatistics/scripts/BTE3207_Advanced_Biostatistics/", 
               "/Users/minsikkim/Dropbox (Personal)/Inha/5_Lectures/Advanced biostatistics/scripts/BTE3207_Advanced_Biostatistics/")
path_library <- 
        "/Library/Frameworks/R.framework/Resources/library"
str_libraries <- c("tidyverse", "pacman", "yaml")



YAML_header <-
        '---
title: "BTE3207 week 13-2"
author: "Minsik Kim"
date: "2032.11.24"
output:
    rmdformats::downcute:
        downcute_theme: "chaos"
        code_folding: hide
        fig_width: 6
        fig_height: 6
---'
seed <- "20230920"

#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Loads libraries, file paths, and other document options.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Boot <- function() {
        .libPaths(path_library)
        
        require(pacman)
        pacman::p_load(c("knitr", "rmarkdown", "rmdformats", "yaml", "lmerTest", "vegan", "phyloseq", "ggplot"))
        
        knitr::opts_knit$set(root.dir = path_working)
        
        str_libraries |> unique() |> sort() -> str_libraries
        pacman::p_load(char = str_libraries)
        
        set.seed(seed)
}
FUN.LineZero.Boot()
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Outputs R environment report.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Report <- function() {
        cat("Line Zero Environment:\n\n")
        paste("R:", pacman::p_version(), "\n") |> cat()
        cat("Libraries:\n")
        for (str_libraries in str_libraries) {
                paste(
                        "    ", str_libraries, ": ", pacman::p_version(package = str_libraries),
                        "\n", sep = ""
                ) |> cat()
        }
        paste("\nOperating System:", pacman::p_detectOS(), "\n") |> cat()
        paste("    Library Path:", path_library, "\n") |> cat()
        paste("    Working Path:", path_working, "\n") |> cat()
        paste("Seed:", seed, "\n\n") |> cat()
        cat("YAML Header:\n")
        cat(YAML_header)
}
FUN.LineZero.Report()


```


# Agenda

This lecture will practice running 

1. running regerssion analysis to binary outcomes (logistic regression)

and

2. Multiple hypothesis testing

# Before begin..

Let's load the data

```{r}

dataset_sbp <- read.csv("dataset/sbp_dataset_korea_2013-2014.csv")

```


```{r}

dataset_sbp$BTH_G %>% hist

```
Where, each level of BTH_G means
1: 20 to 24 years old
2: 25 to 26 years old
...
26: is 73 to 74
27: 75 years old or older.

# Data manipulation

from this dataset, we can calulate 

## Hypertension

```{r}
dataset_sbp$hypertension <- ifelse(dataset_sbp$SBP > 130 |
                                           dataset_sbp$DBP > 80,
                                   1,
                                   0) %>%
        as.factor

```



## Obesity

```{r}
dataset_sbp$obesity <- ifelse(dataset_sbp$BMI > 25,
                                   1,
                                   0) %>% 
        as.factor()
```





## Female (re-level)

```{r}
dataset_sbp$Female <- ifelse(dataset_sbp$SEX == 2, 
                          1,
                          0) %>%
        as.factor
```




## DIS (re-level)

```{r}
dataset_sbp$DIS <- factor(dataset_sbp$DIS, 
                          levels = c(1,2,3,4),
                          labels = c("Hypertension and diabetes",
                                     "Hypertension",
                                     "Diabetes",
                                     "No history"))
```


```{r}
set.seed(1)

dataset_sbp_small <- subset(dataset_sbp,
                        row.names(dataset_sbp)%in% sample(x = 1:1000000, size = 1000))

```

# Generating table

```{r}
table(dataset_sbp$obesity,dataset_sbp$Female)
```



```{r}
data <- data.frame(yearly_veg = c(56, 45, 40, 39, 29, 11, 10, 8, 5, 3, 2),
                   heartdisease = c(1,1,1,1,0,1,0,0,0,0,0))

data2 <- data.frame(yearly_veg = c(55, 40, 40, 38, 37, 21, 19, 10, 4, 2, 2),
                   heartdisease = c(1,1,1,1,0,1,1,0,0,0,0))


model <- glm(data = data, heartdisease ~ yearly_veg, family = "binomial")

model %>% summary

data2$prediction <- ifelse(predict(model, newdata = data2, type = "response") > 0.5,
                           1, 0)


view(data2)
        

```


```{r}
#install.packages("pROC")
library(pROC)

roc(data2$heartdisease,  data2$prediction, plot = TRUE,
               print.auc = TRUE)


roc(data2$heartdisease,  c(0,0,0,0,0,1,1,0,1,0,0), plot = TRUE,
               print.auc = TRUE)


roc(data2$heartdisease,  data2$heartdisease, plot = TRUE,
               print.auc = TRUE)



```

```{r}


roc(data2$heartdisease,  c(0,0,0,0,0,1,1,0,1,0,0), plot = TRUE,
               print.auc = TRUE)


roc(data2$heartdisease,  data2$heartdisease, plot = TRUE,
               print.auc = TRUE)


```


```{r}
#install.packages("pROC")
pbd <- daewr::PBDes(nruns = 12, nfactors = 5)
names(pbd) <- c("YE", "Tryptone", "Glycerol", "PBS", "KH2PO4")

pbd$DCW <- c(3.15, 2.12, 3.77, 2.54, 1.23, 2.2, 2.3, 6.12, 5.79, 4.12, 4.22, 0.08)
view(pbd)
pbd_model <- lm (data = pbd, DCW ~ YE + Tryptone + Glycerol + PBS + KH2PO4)


summary(pbd_model)$coefficients[,1] + 2.447 * summary(pbd_model)$coefficients[,2]
summary(pbd_model)$coefficients[,1] + 3.97 * summary(pbd_model)$coefficients[,2]
?confint
pbd_model %>% anova
```

# Bibliography

```{r warning=FALSE, message=FALSE, echo=FALSE}
#===============================================================================
#BTC.LineZero.Footer.1.1.0
#===============================================================================
#R markdown citation generator.
#===============================================================================
#RLB.Dependencies:
#   magrittr, pacman, stringr
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#BTC.Dependencies:
#   LineZero.Header
#===============================================================================
#Generates citations for each explicitly loaded library.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
str_libraries <- c("r", str_libraries)
for (str_libraries in str_libraries) {
        str_libraries |>
                pacman::p_citation() |>
                print(bibtex = FALSE) |>
                capture.output() %>%
                .[-1:-3] %>% .[. != ""] |>
                stringr::str_squish() |>
                stringr::str_replace("_", "") |>
                cat()
        cat("\n")
}
#===============================================================================
```
