---
title: "COD_week9_2_MGK_BTE3207"
author: "Minsik Kim"
date: "2024-10-31"
output:
        rmdformats::downcute:
        downcute_theme: "chaos"
code_folding: show
fig_width: 6
fig_height: 6
df_print: paged
editor_options: 
        chunk_output_type: inline
markdown: 
        wrap: 72
---

<!-- <style> -->
<!--   /* Default light mode styles */ -->
<!--   .reactable { -->
<!--     background-color: #ffffff !important; /* Light background */ -->
<!--     color: #000000 !important;            /* Dark text */ -->
<!--     border-color: #cccccc !important;     /* Light border */ -->
<!--   } -->

<!-- </style> -->

```{r warning=FALSE, message=FALSE, echo=FALSE, results='hide', setup}
#===============================================================================
#BTC.LineZero.Header.1.1.0
#===============================================================================
#R Markdown environment setup and reporting utility.
#===============================================================================
#RLB.Dependencies:
#   knitr, magrittr, pacman, rio, rmarkdown, rmdformats, tibble, yaml
#===============================================================================
#Input for document parameters, libraries, file paths, and options.
#=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=
knitr::opts_chunk$set(message=FALSE, warning = FALSE)

path_working <- 
        ifelse(sessionInfo()[1]$R.version$platform == "x86_64-pc-linux-gnu",
               "/mnt/4T_samsung/Dropbox/",
               ifelse(sessionInfo()[1]$R.version$platform == "aarch64-apple-darwin20",
                      "/Volumes/macdrive/Dropbox/", 
                      "/Users/minsikkim/Dropbox (Personal)/"))
path_library <- 
        ifelse(sessionInfo()[1]$R.version$platform == "x86_64-pc-linux-gnu",
               "/home/bagel/R_lib/",
               "/Library/Frameworks/R.framework/Resources/library/")


str_libraries <- c("tidyverse", "pacman", "yaml", "reactable")



YAML_header <-
        '---
title: "BTE3207 week 9-2"
author: "Minsik Kim"
date: "2024.10.31"
output:
    rmdformats::downcute:
        downcute_theme: "chaos"
        code_folding: hide
        fig_width: 6
        fig_height: 6
---'
seed <- "20241031"

#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Loads libraries, file paths, and other document options.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Boot <- function() {
        .libPaths(path_library)
        
        require(pacman)
        pacman::p_load(c("knitr", "rmarkdown", "rmdformats", "yaml"))
        
        knitr::opts_knit$set(root.dir = path_working)
        
        str_libraries |> unique() |> sort() -> str_libraries
        pacman::p_load(char = str_libraries)
        
        set.seed(seed)
}
FUN.LineZero.Boot()
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Outputs R environment report.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Report <- function() {
        cat("Line Zero Environment:\n\n")
        paste("R:", pacman::p_version(), "\n") |> cat()
        cat("Libraries:\n")
        for (str_libraries in str_libraries) {
                paste(
                        "    ", str_libraries, ": ", pacman::p_version(package = str_libraries),
                        "\n", sep = ""
                ) |> cat()
        }
        paste("\nOperating System:", pacman::p_detectOS(), "\n") |> cat()
        paste("    Library Path:", path_library, "\n") |> cat()
        paste("    Working Path:", path_working, "\n") |> cat()
        paste("Seed:", seed, "\n\n") |> cat()
        cat("YAML Header:\n")
        cat(YAML_header)
}
FUN.LineZero.Report()


```


# Before begin..



Function to be used.

```{r}

lm_eqn <- function(m){
    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2, 
         list(a = format(unname(coef(m)[1]), digits = 2),
              b = format(unname(coef(m)[2]), digits = 2),
             r2 = format(summary(m)$r.squared, digits = 3)))
    as.character(as.expression(eq));
}

```

# Example 10

```{r}

p10_data <- data.frame(subejct = 1:8,
                      before = c(126, 131, 129, 122, 130, 131, 128, 134),
                      after = c(131, 121, 131, 114, 133, 119, 127, 129)) %>% 
         
        mutate(change = before - after)

t.test(p10_data$change)
#or
t.test(p10_data$after, p10_data$before, paired = T)

```

# Example 11

```{r}
p11_data <- data.frame(subejct = 1:20,
                       group =  c(replicate(10, "low_carb"), replicate(10, "low_fat")),
                      data = c(-10.8, -10.6, 3.5, 10.4, -7.4, -2.6, -6.5, 1.7, -2.9, -7.6,
                               -1.3, -0.9, -1.2, -0.5, -1.6, -0.9, -2.5, -5, -3.4, -1.9)) 

t.test(x = subset(p11_data, p11_data$group == "low_carb")$data, 
       y = subset(p11_data, p11_data$group == "low_fat")$data,
       paired = F,
       var.equal = F)


```


# Example 12

```{r}

p12_data <- data.frame(group = c("obese", "not_obese"),
                       low_fat = c(52, 130),
                       low_carb = c(30, 209),
                       no_diet = c(90, 130)) %>%
        column_to_rownames("group") %>% t %>% data.frame()

library(epitools)

oddsratio(as.matrix(p12_data), rev = "rows")

chisq.test(p12_data)


```

# Example 13

```{r}

p13_data <- data.frame(group = c("Obese", "Not obese"),
                       hypertension = c(6, 2),
                       normal = c(5, 9)) %>%
        column_to_rownames("group")

a <- p13_data[1,1]
b <- p13_data[1,2]
c <- p13_data[2,1]
d <- p13_data[2,2]



(factorial(a+b) * factorial(a+c) * factorial(c+d) * factorial(b+d))/
(factorial(a+b+c+d) * factorial(a) * factorial(b) * factorial(c) * factorial(d))

cat("\n chi.sq-test")
chisq.test(p13_data)

# Counts of hypertension cases
hypertension_counts <- c(89, 102)
# Total counts for each group
total_counts <- c(89 + 93, 102 + 137)

cat("\n z-test")
prop.test(hypertension_counts, total_counts, alternative = "two.sided", correct = FALSE)
```



# Statistics on the results


# Score data in baxplot - by each problem set

```{r}
# 
# midterm <- readxl::read_excel("Inha/5_Lectures/2023/Advanced biostatistics/2023F/Scoreboard/BTE3207_2023F_midterm_results.xlsx") %>%
#         tidyr::gather(., problemset, score, `1`:`12`, factor_key=TRUE)

#readxl::read_excel("Inha/5_Lectures/2024/Advanced biostatistics/2024F/BTE3207_2024F_midterm_score.xlsx")

midterm <- readxl::read_excel("Inha/5_Lectures/2024/Advanced biostatistics/2024F/BTE3207_2024F_midterm_score.xlsx") %>%
        select(-name,-`87_score`, -`100_score`, -`1st_to_100`) %>% 
        subset(., .$student_id != 0)
        
full_score <- readxl::read_excel("Inha/5_Lectures/2024/Advanced biostatistics/2024F/BTE3207_2024F_midterm_score.xlsx") %>%
        select(-name,-`87_score`, -`100_score`, -`1st_to_100`) %>% 
        subset(., .$student_id == 0) %>%
        select(-student_id, ) %>%
        t %>%
        data.frame(full_score = .) %>%
        rownames_to_column("problemset")

students_list <- readxl::read_excel("Inha/5_Lectures/2024/Advanced biostatistics/2024F/BTE3207_2024F_students_list.xlsx") %>%
        rename(student_id = "학번",
               department = "학부(과)",
               major = "전공",
               year = "학년") %>% 
        select(c("student_id", "department", "major", "year")) %>%
        mutate(department = case_when(department == "생명공학과" ~ "BE",
                                      department == "식품영양학과" ~ "FS",
                                      department == "화학과" ~ "CHEM",
                                      department == "통계학과" ~ "STAT",
                                      department == "해양과학과" ~ "OS",
                                      .default = NA))



midterm_merged <- merge(midterm, students_list, by = "student_id", all = T) %>%
        tidyr::gather(., problemset, score, `p1`:`p13`, factor_key=TRUE) %>%
        merge(., full_score, by = "problemset") %>%
        mutate(score_perc = score/full_score*100, .before = "full_score") %>%
        tidyr::gather(., score_type, score, `score`:`score_perc`, factor_key=TRUE)


midterm_total <- midterm_merged %>%
        mutate(student_year = substr(student_id, 3,4)) %>% 
        group_by(student_id, score_type) %>%
        summarise(student_year = student_year[1],
                  year = year[1],
                  department = department[1],
                  major = major[1],
                  total = sum(score)) %>%
        mutate(total = case_when(score_type == "score_perc" ~ total/1300 * 100,
                                 .default = total))


cat("\n Summary of raw score")

midterm_total %>% 
        subset(., .$score_type == "score") %>%
        .$total %>%
        summary


cat("\n Summary of percent score (over 100)")

midterm_total %>% 
        subset(., .$score_type != "score") %>%
        .$total %>%
        summary


midterm_total %>% 
        #subset(., .$score_type == "score") %>%
        ggplot(aes(x=total, fill = "white")) +
        geom_histogram() +
        theme_classic() +
        theme(legend.position = "none") +
        ylab("Number of students") +
        xlab("Raw score | percent scroe") +
        facet_wrap(~score_type, ncol = 1) +
        MetBrewer::scale_fill_met_d("Lakota") 
        


```


# Boxplot of each problemset


```{r}
midterm_merged %>% 
        ggplot(aes(x = problemset, y = score, fill = problemset)) +
                       geom_boxplot() +
        scale_fill_brewer(name = "Problem set",
                          type = "qual", 
                          palette = 3) +
        theme_classic(base_family = "serif",
                      base_size = 13) +
        xlab("Problem set") +
        ylab("Score") +
        facet_wrap(~score_type, scales = "free")


```


# Score data in histogram - by each problem set

```{r}
midterm_merged %>% 
        subset(., .$score_type == "score") %>%
        ggplot(aes(x = score, fill = problemset)) +
        geom_histogram() +
        MetBrewer::scale_fill_met_d(name = "Signac") +
        facet_wrap(~problemset, scales = "free") +
        #guide_legend(title = "Problem set") +
        theme_classic(base_family = "serif",
                      base_size = 13) +
        ylab("Number of students") +
        xlab("Range of raw score")


midterm_merged %>% 
        subset(., .$score_type == "score_perc") %>%
        ggplot(aes(x = score, fill = problemset)) +
        geom_histogram() +
        MetBrewer::scale_fill_met_d(name = "Signac") +
        facet_wrap(~problemset, scales = "free") +
        #guide_legend(title = "Problem set") +
        theme_classic(base_family = "serif",
                      base_size = 13) +
        ylab("Number of students") +
        xlab("Range of % score")


                       

```

# Score total - summary statistics

```{r}

midterm_merged %>% 
        subset(., .$score_type == "score") %>% 
        select(-c(full_score, student_id, department, major, year)) %>%
        reactable(groupBy = c("problemset"),
                  columns = list(
                          score_type = colDef(aggregate = "frequency"),
                          score = colDef(aggregate = "mean", format = colFormat(digits = 2))
                  )) %>%
        reactablefmtr::add_title("Interactive table of summary statistics raw score by each problemset")
        


midterm_merged %>% 
        subset(., .$score_type != "score") %>% 
        select(-c(full_score, student_id, department, major, year)) %>%
        reactable(groupBy = c("problemset"),
                  columns = list(
                          score_type = colDef(aggregate = "frequency"),
                          score = colDef(aggregate = "mean", format = colFormat(digits = 2))
                  )) %>%
        reactablefmtr::add_title("Interactive table of summary statistics of % score by each problemset")
        


```


# Score total - boxplot and histogram

```{r}

midterm_total %>%
        ggplot(aes(x = department, y = total, fill = department)) +
        geom_boxplot() +
        MetBrewer::scale_fill_met_d(name = "Signac") +
        theme_classic(base_family = "serif",
                      base_size = 13) +
        xlab("By department") +
        theme(axis.text.x = element_blank(),
              legend.position = "none") +
        facet_wrap(~score_type, scales = "free")+
        ylab("Total score")


midterm_total %>%
        ggplot(aes(x = student_year, y = total, fill = student_year)) +
        geom_boxplot() +
        MetBrewer::scale_fill_met_d(name = "Signac") +
        theme_classic(base_family = "serif",
                      base_size = 13) +
        theme(axis.text.x = element_blank(),
              legend.position = "none") +
        guides(fill = guide_legend()) +
        xlab("By school admission year") +
        facet_wrap(~score_type, scales = "free")+
        ylab("Total score")


```

# For fun, a statistical test

Is there a different between junior and senior?

# Total score - boxplot by year

```{r}

midterm_total %>%
        subset(., .$score_type == "score") %>% 
        mutate(year = as.factor(year)) %>% 
        ggplot(aes(y = total, x = year, fill = year)) +
        geom_boxplot() +
        #facet_wrap(~year, ncol = 1) +
        theme_classic(base_family = "serif",
                      base_size = 13) +
        ylab("Score") +
        xlab("Year of study at Inha")
        


```

# Total score - histogram by year

```{r}

midterm_total %>%
        #subset(., .$score_type == "score") %>% 
        mutate(year = as.factor(year)) %>%
        ggplot(aes(x = total, fill = year)) +
        geom_histogram(aes(fill = year), position = position_identity(), alpha = 0.5) +
        scale_fill_brewer(name = "Year of study at Inha",
                          type = "qual", 
                          palette = 6) +
        #facet_wrap(~year, ncol = 1) +
        theme_classic(base_family = "serif",
                      base_size = 13) +
        theme(legend.position = "top") + 
        facet_wrap(~score_type, ncol = 1, scales = "free") + 
        xlab("Score") +
        ylab("Number of students")




```


# *t*-tests

t-test (unpaired, unequal variance)

H<sub>0</sub> (null hypothesis): true difference in means is 0, i.e., there is no difference between junior and senior
H<sub>a</sub> (alternative hypothesis): true difference in means is not equal to 0.



```{r}

raw_score_data <- midterm_total %>%
        subset(., .$score_type == "score") %>% 
        mutate(year = as.factor(year))

t.test(subset(raw_score_data, raw_score_data$year == "4") %>% .$total,
       subset(raw_score_data, raw_score_data$year == "3") %>% .$total,
       paired = F, var.equal = F)




perc_score_data <- midterm_total %>%
        subset(., .$score_type != "score") %>% 
        mutate(year = as.factor(year))

t.test(subset(perc_score_data, perc_score_data$year == "4") %>% .$total,
       subset(perc_score_data, perc_score_data$year == "3") %>% .$total,
       paired = F, var.equal = F)



        
```

**p**-value is 0.48. Therefore, we accect the null hypothesis (there is no significant difference between mean of junior and senior).


Alternatively,

```{r}

raw_score_data %>%
        lm(total ~ year, data = .) %>% 
        summary


perc_score_data %>%
        lm(total ~ year, data = .) %>% 
        summary


```

# LM on each problemset

```{r}
midterm_merged %>%
        subset(., .$score_type=="score_perc") %>%
        lm(data = ., score ~ problemset) %>%
        summary %>%
        .$coefficient %>%
        round(3) %>%
        format(digits = 3) %>%
        data.frame(check.names = F) %>%
        mutate(` ` = case_when(`Pr(>|t|)` %>%as.numeric < 0.001 ~ "***",
                               `Pr(>|t|)` %>%as.numeric < 0.01 ~ "**",
                               `Pr(>|t|)` %>%as.numeric < 0.05 ~ "*",
                               .default = NA)) %>%
        reactable(sortable = T, showPageSizeOptions = T)
```
# LM on raw score

With categorical variables (department)

```{r}

raw_score_data %>%
        lm(total ~ department, data = .) %>% 
        summary %>%
        .$coefficient %>%
        data.frame(check.names = F) %>%
        round(3) %>% 
        format(digits = 3) %>%
        mutate(row_temp = c("(Intercept)", "Dept. 2", "Dept. 3", "Dept. 4", "Dept. 5")) %>%
        remove_rownames() %>%
        column_to_rownames("row_temp") %>%
        reactable

```
# LM on percent score

```{r}

perc_score_data %>%
        lm(total ~ department, data = .) %>% 
        summary %>%
        .$coefficient %>%
        data.frame(check.names = F) %>%
        round(3) %>% 
        format(digits = 3) %>%
        mutate(row_temp = c("(Intercept)", "Dept. 2", "Dept. 3", "Dept. 4", "Dept. 5")) %>%
        remove_rownames() %>%
        column_to_rownames("row_temp") %>%
        reactable



```

We are going to learn more about the linear model, at the later half of the semester!
# Bibliography

```{r warning=FALSE, message=FALSE, echo=FALSE}
#===============================================================================
#BTC.LineZero.Footer.1.1.0
#===============================================================================
#R markdown citation generator.
#===============================================================================
#RLB.Dependencies:
#   magrittr, pacman, stringr
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#BTC.Dependencies:
#   LineZero.Header
#===============================================================================
#Generates citations for each explicitly loaded library.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
str_libraries <- c("r", str_libraries)
for (str_libraries in str_libraries) {
        str_libraries |>
                pacman::p_citation() |>
                print(bibtex = FALSE) |>
                capture.output() %>%
                .[-1:-3] %>% .[. != ""] |>
                stringr::str_squish() |>
                stringr::str_replace("_", "") |>
                cat()
        cat("\n")
}
#===============================================================================
```
