---
title: "COD_Midterm_BTE3207"
author: "Minsik Kim"
date: "2023-10-21"
output:
        rmdformats::downcute:
        downcute_theme: "chaos"
code_folding: show
fig_width: 6
fig_height: 6
df_print: paged
editor_options: 
        chunk_output_type: inline
markdown: 
        wrap: 72
---
        

        
```{r warning=FALSE, message=FALSE, echo=FALSE, results='hide', setup}
#===============================================================================
#BTC.LineZero.Header.1.1.0
#===============================================================================
#R Markdown environment setup and reporting utility.
#===============================================================================
#RLB.Dependencies:
#   knitr, magrittr, pacman, rio, rmarkdown, rmdformats, tibble, yaml
#===============================================================================
#Input for document parameters, libraries, file paths, and options.
#=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=
knitr::opts_chunk$set(message=FALSE, warning = FALSE)

path_working <- "/Users/minsikkim/Dropbox (Personal)/Inha/5_Lectures/Advanced biostatistics/scripts/BTE3207_Advanced_Biostatistics"
path_library <- "/Library/Frameworks/R.framework/Resources/library"
str_libraries <- c("tidyverse", "pacman", "yaml")



YAML_header <-
        '---
title: "BTE3207 week 7-1"
author: "Minsik Kim"
date: "2032.10.10"
output:
    rmdformats::downcute:
        downcute_theme: "chaos"
        code_folding: hide
        fig_width: 6
        fig_height: 6
---'
seed <- "20230920"

#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Loads libraries, file paths, and other document options.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Boot <- function() {
        .libPaths(path_library)
        
        require(pacman)
        pacman::p_load(c("knitr", "rmarkdown", "rmdformats", "yaml"))
        
        knitr::opts_knit$set(root.dir = path_working)
        
        str_libraries |> unique() |> sort() -> str_libraries
        pacman::p_load(char = str_libraries)
        
        set.seed(seed)
}
FUN.LineZero.Boot()
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Outputs R environment report.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Report <- function() {
        cat("Line Zero Environment:\n\n")
        paste("R:", pacman::p_version(), "\n") |> cat()
        cat("Libraries:\n")
        for (str_libraries in str_libraries) {
                paste(
                        "    ", str_libraries, ": ", pacman::p_version(package = str_libraries),
                        "\n", sep = ""
                ) |> cat()
        }
        paste("\nOperating System:", pacman::p_detectOS(), "\n") |> cat()
        paste("    Library Path:", path_library, "\n") |> cat()
        paste("    Working Path:", path_working, "\n") |> cat()
        paste("Seed:", seed, "\n\n") |> cat()
        cat("YAML Header:\n")
        cat(YAML_header)
}
FUN.LineZero.Report()


```

Function to be used.

```{r}

lm_eqn <- function(m){
    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2, 
         list(a = format(unname(coef(m)[1]), digits = 2),
              b = format(unname(coef(m)[2]), digits = 2),
             r2 = format(summary(m)$r.squared, digits = 3)))
    as.character(as.expression(eq));
}

```

# Example 9

```{r}

p9_data <- data.frame(subejct = 1:8,
                      before = c(126, 131, 129, 122, 130, 131, 128, 134),
                      after = c(131, 121, 131, 114, 133, 119, 127, 129)) %>%
        mutate(change = before - after)

t.test(p9_data$change)
#or
t.test(p9_data$after, p9_data$before, paired = T)

```

# Example 10

```{r}
p10_data <- data.frame(subejct = 1:20,
                       group =  c(replicate(10, "low_carb"), replicate(10, "low_fat")),
                      data = c(-10.8, -10.6, 3.5, 10.4, -7.4, -2.6, -6.5, 1.7, -2.9, -7.6,
                               -1.3, -0.9, -1.2, -0.5, -1.6, -0.9, -2.5, -5, -3.4, -1.9)) 

t.test(x = subset(p10_data, p10_data$group == "low_carb")$data, 
       y = subset(p10_data, p10_data$group == "low_fat")$data,
       paired = F,
       var.equal = F)


```


# Example 11

```{r}

p11_data <- data.frame(group = c("obese", "not_obese"),
                       low_fat = c(52, 130),
                       low_carb = c(30, 209),
                       no_diet = c(90, 130)) %>%
        column_to_rownames("group") %>% t %>% data.frame()

library(epitools)

oddsratio(as.matrix(p11_data), rev = "rows")

chisq.test(p11_data)


```

# Example 12

```{r}

p12_data <- data.frame(group = c("hypertension", "normal"),
                       low_fat = c(89, 102),
                       low_carb = c(93, 137)) %>%
        column_to_rownames("group")


chisq.test(p12_data)

```


# Statistics on the results


# Score data in baxplot - by each problem set

```{r}

midterm <- readxl::read_excel("/Users/minsikkim/Dropbox (Personal)/Inha/5_Lectures/Advanced biostatistics/2023F/Midterm_result/BTE3207_2023F_midterm_results.xlsx") %>%
        tidyr::gather(., problemset, score, `1`:`12`, factor_key=TRUE)

midterm %>% 
        ggplot(aes(x = problemset, y = score, fill = problemset)) +
                       geom_boxplot() +
        scale_fill_brewer(name = "Problem set",
                          type = "qual", 
                          palette = 3) +
        theme_classic(base_family = "serif",
                      base_size = 13) +
        xlab("Problem set") +
        ylab("Score")


```


# Score data in histogram - by each problem set

```{r}
midterm %>% 
        ggplot(aes(x = score, fill = problemset)) +
        geom_histogram() +
        scale_fill_brewer(name = "Problem set",
                          type = "qual", 
                          palette = 3) +
        facet_wrap(~problemset, scales = "free") +
        #guide_legend(title = "Problem set") +
        theme_classic(base_family = "serif",
                      base_size = 13) +
        ylab("Number of students")

                       

```

# Score total - summary statistics

```{r}

readxl::read_excel("/Users/minsikkim/Dropbox (Personal)/Inha/5_Lectures/Advanced biostatistics/2023F/Midterm_result/BTE3207_2023F_midterm_results.xlsx") %>% 
        .$total %>%
        summary

```


# Score total - boxplot and histogram

```{r}

a <- merge(readxl::read_excel("/Users/minsikkim/Dropbox (Personal)/Inha/5_Lectures/Advanced biostatistics/2023F/Midterm_result/BTE3207_2023F_midterm_results.xlsx"),
        readxl::read_excel("/Users/minsikkim/Dropbox (Personal)/Inha/5_Lectures/Advanced biostatistics/2023F/BTE3207_2023F_students_list.xlsx") %>% rename(student_id = "학번",
       department = "학부(과)",
       major = "전공",
       year = "학년") %>% select(c("student_id", "department", "major", "year")),
       by = "student_id", all = T
       ) %>%
        mutate(year = as.factor(year)) %>% 
        ggplot(aes(y = total)) +
        geom_boxplot() +
        theme_classic(base_family = "serif",
                      base_size = 13) +
        theme(legend.position = "top",
              axis.text.x = element_blank()) + 
        xlab("") +
        ylab("Total score")


b <- merge(readxl::read_excel("/Users/minsikkim/Dropbox (Personal)/Inha/5_Lectures/Advanced biostatistics/2023F/Midterm_result/BTE3207_2023F_midterm_results.xlsx"),
        readxl::read_excel("/Users/minsikkim/Dropbox (Personal)/Inha/5_Lectures/Advanced biostatistics/2023F/BTE3207_2023F_students_list.xlsx") %>% rename(student_id = "학번",
       department = "학부(과)",
       major = "전공",
       year = "학년") %>% select(c("student_id", "department", "major", "year")),
       by = "student_id", all = T
       ) %>%
        mutate(year = as.factor(year)) %>% 
        ggplot(aes(x = total)) +
        geom_histogram(binwidth = 7) +
        theme_classic(base_family = "serif",
                      base_size = 13) +
        theme(legend.position = "top") + 
        xlab("Score") +
        ylab("Number of students")
        
ggpubr::ggarrange(a, b, align = "hv", widths = c(1,3))

```

# For fun, a statistical test

Is there a different between junior and senior?

# Total score - boxplot by year

```{r}

merge(readxl::read_excel("/Users/minsikkim/Dropbox (Personal)/Inha/5_Lectures/Advanced biostatistics/2023F/Midterm_result/BTE3207_2023F_midterm_results.xlsx"),
        readxl::read_excel("/Users/minsikkim/Dropbox (Personal)/Inha/5_Lectures/Advanced biostatistics/2023F/BTE3207_2023F_students_list.xlsx") %>% rename(student_id = "학번",
       department = "학부(과)",
       major = "전공",
       year = "학년") %>% select(c("student_id", "department", "major", "year")),
       by = "student_id", all = T
       ) %>%
        mutate(year = as.factor(year)) %>% 
        ggplot(aes(y = total, x = year, fill = year)) +
        geom_boxplot() +
        #facet_wrap(~year, ncol = 1) +
        theme_classic(base_family = "serif",
                      base_size = 13) +
        ylab("Score") +
        xlab("Year of study at Inha")
        


```

# Total score - histogram by year

```{r}

merge(readxl::read_excel("/Users/minsikkim/Dropbox (Personal)/Inha/5_Lectures/Advanced biostatistics/2023F/Midterm_result/BTE3207_2023F_midterm_results.xlsx"),
        readxl::read_excel("/Users/minsikkim/Dropbox (Personal)/Inha/5_Lectures/Advanced biostatistics/2023F/BTE3207_2023F_students_list.xlsx") %>% rename(student_id = "학번",
       department = "학부(과)",
       major = "전공",
       year = "학년") %>% select(c("student_id", "department", "major", "year")),
       by = "student_id", all = T
       ) %>%
        mutate(year = as.factor(year)) %>% 
        ggplot(aes(x = total, fill = year)) +
        geom_histogram(aes(y = ..density..),
                       binwidth = 2, position = position_identity(), alpha = 0.5) +
        scale_fill_brewer(name = "Year of study at Inha",
                          type = "qual", 
                          palette = 6) +
        #facet_wrap(~year, ncol = 1) +
        theme_classic(base_family = "serif",
                      base_size = 13) +
        theme(legend.position = "top") + 
        xlab("Score") +
        ylab("Number of students")




```


t-test (unpaired, unequal variance)

H<sub>0</sub> (null hypothesis): true difference in means is 0, i.e., there is no difference between junior and senior
H<sub>a</sub> (alternative hypothesis): true difference in means is not equal to 0.



```{r}

data <- merge(readxl::read_excel("/Users/minsikkim/Dropbox (Personal)/Inha/5_Lectures/Advanced biostatistics/2023F/Midterm_result/BTE3207_2023F_midterm_results.xlsx"),
        readxl::read_excel("/Users/minsikkim/Dropbox (Personal)/Inha/5_Lectures/Advanced biostatistics/2023F/BTE3207_2023F_students_list.xlsx") %>% rename(student_id = "학번",
       department = "학부(과)",
       major = "전공",
       year = "학년") %>% select(c("student_id", "department", "major", "year")),
       by = "student_id", all = T
       )
        

t.test(subset(data, data$year == "4") %>% .$total,
       subset(data, data$year == "3") %>% .$total,
       paired = F, var.equal = F)



        
```

**p**-value is 0.48. Therefore, we accect the null hypothesis (there is no significant difference between mean of junior and senior).


Alternatively,

```{r}

merge(readxl::read_excel("/Users/minsikkim/Dropbox (Personal)/Inha/5_Lectures/Advanced biostatistics/2023F/Midterm_result/BTE3207_2023F_midterm_results.xlsx"),
        readxl::read_excel("/Users/minsikkim/Dropbox (Personal)/Inha/5_Lectures/Advanced biostatistics/2023F/BTE3207_2023F_students_list.xlsx") %>% rename(student_id = "학번",
       department = "학부(과)",
       major = "전공",
       year = "학년") %>% select(c("student_id", "department", "major", "year")),
       by = "student_id", all = T
       ) %>%
        mutate(year = as.factor(year)) %>%
        lm(total ~ year, data = .) %>% 
        summary


```


Correlation - % score by their orders

```{r}

midterm <- midterm %>%
        mutate(problem_full_score = case_when(problemset == 1 ~ 6,
                                      problemset == 2 ~ 5,
                                      problemset == 3 ~ 5,
                                      problemset == 4 ~ 5,
                                      problemset == 5 ~ 5,
                                      problemset == 6 ~ 5,
                                      problemset == 7 ~ 9,
                                      problemset == 8 ~ 5,
                                      problemset == 9 ~ 6,
                                      problemset == 10 ~ 6,
                                      problemset == 11 ~ 9,
                                      problemset == 12 ~ 8,
                                      .default = 0),
               perc_score = score/problem_full_score * 100)


midterm %>% 
        ggplot(aes(x = as.numeric(problemset), y = perc_score, color = as.numeric(problemset))) +
                geom_jitter() +
                theme_classic(base_family = "serif", base_size = 20) +
                viridis::scale_color_viridis() + 
                xlab("Problem set order") +
                ylab("% right answer") #+ 
                #geom_smooth(method = "lm") 

midterm %>% 
        ggplot(aes(x = as.numeric(problemset), y = perc_score, color = as.numeric(problemset))) +
                geom_jitter() +
                theme_classic(base_family = "serif", base_size = 20) +
                viridis::scale_color_viridis() + 
                xlab("Problem set order") +
                ylab("% right answer") + 
                geom_smooth(method = "lm") +
                geom_label(x = 7.5, y = 100,
                  label = lm_eqn(lm(data = midterm,perc_score ~ as.numeric(problemset))),
                  parse = TRUE,
                  family='serif',
                  size = 8, label.size = 0, color = "blue")


```


```{r}

lm(data = midterm, perc_score ~ as.numeric(problemset)) %>% summary

```

We are going to learn more about the linear model, at the later half of the semester!

# Bibliography

```{r warning=FALSE, message=FALSE, echo=FALSE}
#===============================================================================
#BTC.LineZero.Footer.1.1.0
#===============================================================================
#R markdown citation generator.
#===============================================================================
#RLB.Dependencies:
#   magrittr, pacman, stringr
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#BTC.Dependencies:
#   LineZero.Header
#===============================================================================
#Generates citations for each explicitly loaded library.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
str_libraries <- c("r", str_libraries)
for (str_libraries in str_libraries) {
        str_libraries |>
                pacman::p_citation() |>
                print(bibtex = FALSE) |>
                capture.output() %>%
                .[-1:-3] %>% .[. != ""] |>
                stringr::str_squish() |>
                stringr::str_replace("_", "") |>
                cat()
        cat("\n")
}
#===============================================================================
```
