---
title: "COD_week1_2_MGK_BTE3207"
author: "Minsik Kim"
date: "2024-09-02"
output:
  rmdformats::downcute:
    downcute_theme: "chaos"
  code_folding: show
  fig_width: 6
  fig_height: 6
  df_print: paged
  editor_options:
    chunk_output_type: inline
  markdown:
    wrap: 72
---

<style>
  /* Default light mode styles */
  .reactable {
    background-color: #ffffff !important; /* Light background */
    color: #000000 !important;            /* Dark text */
    border-color: #cccccc !important;     /* Light border */
  }

</style>


```{r warning=FALSE, message=FALSE, echo=FALSE, results='hide', setup}
#===============================================================================
#BTC.LineZero.Header.1.1.0
#===============================================================================
#R Markdown environment setup and reporting utility.
#===============================================================================
#RLB.Dependencies:
#   knitr, magrittr, pacman, rio, rmarkdown, rmdformats, tibble, yaml
#===============================================================================
#Input for document parameters, libraries, file paths, and options.
#=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=
knitr::opts_chunk$set(message=FALSE, warning = FALSE)


path_working <- 
        ifelse(sessionInfo()[1]$R.version$platform == "x86_64-pc-linux-gnu",
               "/mnt/4T_samsung/Dropbox/",
               ifelse(sessionInfo()[1]$R.version$platform == "aarch64-apple-darwin20",
                      "/Volumes/macdrive/Dropbox/", 
                      "/Users/minsikkim/Dropbox (Personal)/"))
path_library <- 
        ifelse(sessionInfo()[1]$R.version$platform == "x86_64-pc-linux-gnu",
               "/home/bagel/R_lib/",
               "/Library/Frameworks/R.framework/Resources/library/")


str_libraries <- c("tidyverse", "pacman", "yaml", "reactable")


YAML_header <-
        '---
title: "BTE3207 week 1-2"
author: "Minsik Kim"
date: "2024.09.02"
output:
    rmdformats::downcute:
        downcute_theme: "chaos"
        code_folding: hide
        fig_width: 6
        fig_height: 6
---'
seed <- "20240902"

#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Loads libraries, file paths, and other document options.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Boot <- function() {
        .libPaths(path_library)
        
        require(pacman)
        pacman::p_load(c("knitr", "rmarkdown", "rmdformats", "yaml"))
        
        knitr::opts_knit$set(root.dir = path_working)
        
        str_libraries |> unique() |> sort() -> str_libraries
        pacman::p_load(char = str_libraries)
        
        set.seed(seed)
}
FUN.LineZero.Boot()
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Outputs R environment report.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Report <- function() {
        cat("Line Zero Environment:\n\n")
        paste("R:", pacman::p_version(), "\n") |> cat()
        cat("Libraries:\n")
        for (str_libraries in str_libraries) {
                paste(
                        "    ", str_libraries, ": ", pacman::p_version(package = str_libraries),
                        "\n", sep = ""
                ) |> cat()
        }
        paste("\nOperating System:", pacman::p_detectOS(), "\n") |> cat()
        paste("    Library Path:", path_library, "\n") |> cat()
        paste("    Working Path:", path_working, "\n") |> cat()
        paste("Seed:", seed, "\n\n") |> cat()
        cat("YAML Header:\n")
        cat(YAML_header)
}
FUN.LineZero.Report()


```

# Working with data in R

There are different type of data in R. The smallest data is called `data element`, and when data element is gathered (with orders), that chunk of data is called `vector`.

Here are some examples.

## Data element

```{r}
alpha <- 1 # this is data element

alpha  # If we call alpha, the computer will display the value assigned to alpha.

```

## Vecetor

To assign vector in R, we use `c()`, with comma `,` as separators. For example:

```{r}

beta <- c(1, 2) # this is a vector

beta

```


```{r}
gamma <- c(alpha, beta)  # This is also a vector

gamma

```

Remember vectors are **ordered**.

```{r}

delta <- c(beta, alpha) # it is still a vector

delta

```



*When you call* `beta` *in R, it will retrieve data associated with *`beta`. *So, how do we assign characters (letters or categorical variable)?* **R recognizes character values with **`""`. For instance:

```{r}

delta_2 <- c("beta", "alpha") 

delta_2

```

With quotation marks, the R will recognize inputs as a new character data, not the data we assigned previously.


# Data frame

So, what should we do with larger data sets, such as Excel spreadsheets?

Vectors in R can be combined into a data frame, **as long as their lengths are the same.**

This is called data frame in R, and it can be created using the `data.frame()` function.


```{r}

epsilon <- data.frame(gamma, delta)  # This is a data frame
# Note that the lengths of both gamma and delta are the same.

epsilon

```


## Selecting variables (vectors) in data frame

A data frame has columns, where each column represents one variable (vector).

We use the dollar sign $ to select a specific variable from a data frame.


```{r}

epsilon$gamma

epsilon$delta


```

`epsilon$gamma` returns the exact same vector as just `gamma`, which was used for constructing epsilon.

Now, how can we generate a larger dataset for analysis?


## Data frame example

We can make data by adding multiple data elements to vectors, and we can bind those vectors to make one dataframe.


```{r}

subject <- c("Joe", "Trump", "Obama", "George")  # Assigning multiple data elements as a vector
height <- c(183, 190, 187, 182)
IsTall <- c("short", "tall", "tall", "short")

example_dataframe <- data.frame(subject, height, IsTall)  # Combining 3 vectors into one data frame

example_dataframe  # Displaying the example_data frame

```



## Loading Data

However, we won’t be typing all the data manually every time it’s needed. To save time, we will directly load data frames from our computer (most commonly `.csv`, but we can also import `.xlsx`) using the `read.csv()` function.

Here is an example using systolic blood pressure (SBP) data from [NHISS](https://nhiss.nhis.or.kr/bd/ab/bdabf003cv.do).

```{r}

dataset_sbp <- read.csv(file = "Inha/5_Lectures/2024/Advanced biostatistics/scripts/BTE3207_Advanced_Biostatistics/dataset/sbp_dataset_korea_2013-2014.csv")

dataset_sbp %>%
        reactable::reactable(sortable = T)

```

Here, the `file = "Inha/5_Lectures/Advanced_biostatistics/datasets/sbp_dataset/sbp_dataset_korea_2013-2014.csv"` need to be the path that you downloaded the file.

When working with data in R, it's often necessary to load data from your computer instead of typing everything manually. The most common file formats for data are `.csv` (Comma-Separated Values) and `.xlsx` (Excel files). 

To load a `.csv` file, we use the `read.csv()` function. However, to load the file correctly, you need to know the **file path**, which tells R where to find the file on your computer.

### Understanding File Paths

A file path is like a map that tells your computer where to find a file. There are two types of file paths:

1. **Absolute Path**: This is the full path from the root directory (e.g., `C:/` on Windows or `/` on Mac/Linux) to the file.
2. **Relative Path**: This is the path relative to your current working directory in R.

### Navigating File Paths in R

- **Mac/Linux Users**: File paths use forward slashes `/` to separate directories.
  - Example: `/Users/username/Documents/data.csv`
- **Windows Users**: File paths use backslashes `\`, but in R, you should use forward slashes `/` or double backslashes `\\`.
  - Example: `C:/Users/username/Documents/data.csv` or `C:\\Users\\username\\Documents\\data.csv`

### Using the Tab Key to Navigate

RStudio provides a convenient feature to help you navigate file paths. When you start typing a file path within quotes (e.g., `"`), you can press the `Tab` key to see a list of files and folders in the current directory. This feature is very helpful for avoiding mistakes in typing paths.

### Setting the Working Directory

Before loading a file, it’s useful to set the working directory to the folder where your data is stored. This way, you can use relative paths instead of absolute paths.

```{r}
#setwd("/Users/username/Documents")  # For Mac/Linux - this is example file path
#setwd("C:/Users/username/Documents")  # For Windows - this is example file path
```

You can check your current working directory by using:

```{r}
getwd()
```

### Example of Loading a CSV File

Here’s how you can load a CSV file into R:

```{r}
# Example for Mac/Linux
#dataset_sbp <- read.csv(file = "/Users/username/Documents/sbp_dataset.csv")

# Example for Windows
#dataset_sbp <- read.csv(file = "C:/Users/username/Documents/sbp_dataset.csv")
```

### Using the `file.choose()` Function

If you're unsure about the file path, R provides a `file.choose()` function that opens a file dialog to help you select the file interactively:

Try using the function after removing the `number sign (#)` of the below code

```{r}
#dataset_sbp <- read.csv(file = file.choose())

```

This method is platform-independent and can be a helpful alternative if you're having trouble with file paths.

The data is loaded as shown below. However, since it contains 1 million lines of data, displaying the entire dataset would make this document too long. Instead, we can use the `head()` or `tail()` functions to view only the first 10 or last 10 rows, respectively.

## head() and tail() Functions {.tabset}

### head()

```{r}
head(dataset_sbp)
```

### tail()

```{r}
tail(dataset_sbp)
```

### Full Dataset

```{r}
dataset_sbp
```

### {-}

This shows (subsets of) gender, age group, SBP, DBP, FBS, DIS, and BMI data for 1,000,000 subjects.

For now, we will use the dollar sign to select one variable from the data frame. Again, I will use the `head()` function to check the first 10 rows of data this time.

```{r}
head(dataset_sbp$SBP)
```

That’s a lot of data!

# Basic R Functions for Statistics

Now, let's use R to calculate summary statistics for this large dataset.

## Mean Calculation

The function for calculating the mean is `mean()`.

```{r}
mean(dataset_sbp$SBP)
```

This is the mean value of all the SBP data (1,000,000 rows).

## SD Calculation

The function for calculating the standard deviation is `sd()`.

```{r}
sd(dataset_sbp$SBP)
```

This is the standard deviation.

## Median

The function for calculating the median is `median()`.

```{r}
median(dataset_sbp$SBP)
```

## Function for Calculating Quantiles: `quantile()`

The `quantile()` function can be used to calculate multiple percentiles at once. Without specifying any percentile, it will return the 0%, 25%, 50%, 75%, and 100% percentiles.

```{r}
quantile(dataset_sbp$SBP)
```

Since the output of `quantile()` contains multiple data elements, the result is a **vector**. Conveniently, this output also includes names. To check the structure of this output, we can use the `str()` function.

```{r}
str(quantile(dataset_sbp$SBP))
```

The `str()` function indicates that it is a `named numeric variable`.

## Quantile() - Continued

But what if we want to find the value at a specific percentile? In that case, we need to provide more information to the function.

These pieces of information are called **arguments**.

Since functions in R can accept multiple inputs, we separate them with commas (`,`).

For example, to calculate the 10th percentile of the SBP dataset:

```{r}
quantile(x = dataset_sbp$SBP, probs = 0.1)
```

# Using the Question Mark to Explore a Function

As these functions can have multiple arguments, and there are many `packages` that can be installed to add new functions to R, developers provide detailed instructions. You can access this information by adding a `?` before a function name.

```{r}
?quantile
```

For some popular functions, details about the `arguments` will usually appear in pop-ups. You can use the `Tab` button to see what the function can do.


# Bibliography

```{r warning=FALSE, message=FALSE, echo=FALSE}
#===============================================================================
#BTC.LineZero.Footer.1.1.0
#===============================================================================
#R markdown citation generator.
#===============================================================================
#RLB.Dependencies:
#   magrittr, pacman, stringr
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#BTC.Dependencies:
#   LineZero.Header
#===============================================================================
#Generates citations for each explicitly loaded library.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
str_libraries <- c("r", str_libraries)
for (str_libraries in str_libraries) {
        str_libraries |>
                pacman::p_citation() |>
                print(bibtex = FALSE) |>
                capture.output() %>%
                .[-1:-3] %>% .[. != ""] |>
                stringr::str_squish() |>
                stringr::str_replace("_", "") |>
                cat()
        cat("\n")
}
#===============================================================================
```
