---
title: "COD_week12_2_MGK_BTE3207"
author: "Minsik Kim"
date: "2024-11-21"
output:
        rmdformats::downcute:
        downcute_theme: "chaos"
code_folding: show
fig_width: 6
fig_height: 6
df_print: paged
editor_options: 
        chunk_output_type: inline
markdown: 
        wrap: 72
---
        
        
        
```{r warning=FALSE, message=FALSE, echo=FALSE, results='hide', setup}
#===============================================================================
#BTC.LineZero.Header.1.1.0
#===============================================================================
#R Markdown environment setup and reporting utility.
#===============================================================================
#RLB.Dependencies:
#   knitr, magrittr, pacman, rio, rmarkdown, rmdformats, tibble, yaml
#===============================================================================
#Input for document parameters, libraries, file paths, and options.
#=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=
knitr::opts_chunk$set(message=FALSE, warning = FALSE)

path_working <- 
        ifelse(sessionInfo()[1]$R.version$platform == "x86_64-pc-linux-gnu",
               "/mnt/4T_samsung/Dropbox/",
               ifelse(sessionInfo()[1]$R.version$platform == "aarch64-apple-darwin20",
                      "/Volumes/macdrive/Dropbox/", 
                      "/Users/minsikkim/Dropbox (Personal)/"))
path_library <- 
        ifelse(sessionInfo()[1]$R.version$platform == "x86_64-pc-linux-gnu",
               "/home/bagel/R_lib/",
               "/Library/Frameworks/R.framework/Resources/library/")


str_libraries <- c("tidyverse", "pacman", "yaml", "reactable")



YAML_header <-
        '---
title: "BTE3207 week 11-2"
author: "Minsik Kim"
date: "2024.11.18"
output:
    rmdformats::downcute:
        downcute_theme: "chaos"
        code_folding: hide
        fig_width: 6
        fig_height: 6
---'
seed <- "20241118"

#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Loads libraries, file paths, and other document options.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Boot <- function() {
        .libPaths(path_library)
        
        require(pacman)
        pacman::p_load(c("knitr", "rmarkdown", "rmdformats", "yaml", "lmerTest"))
        
        knitr::opts_knit$set(root.dir = path_working)
        
        str_libraries |> unique() |> sort() -> str_libraries
        pacman::p_load(char = str_libraries)
        
        set.seed(seed)
}
FUN.LineZero.Boot()
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Outputs R environment report.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Report <- function() {
        cat("Line Zero Environment:\n\n")
        paste("R:", pacman::p_version(), "\n") |> cat()
        cat("Libraries:\n")
        for (str_libraries in str_libraries) {
                paste(
                        "    ", str_libraries, ": ", pacman::p_version(package = str_libraries),
                        "\n", sep = ""
                ) |> cat()
        }
        paste("\nOperating System:", pacman::p_detectOS(), "\n") |> cat()
        paste("    Library Path:", path_library, "\n") |> cat()
        paste("    Working Path:", path_working, "\n") |> cat()
        paste("Seed:", seed, "\n\n") |> cat()
        cat("YAML Header:\n")
        cat(YAML_header)
}
FUN.LineZero.Report()




```


# Analysis of variance

## Before begin..

Let's load the SBP dataset.

```{r}

dataset_sbp <- read.csv(file = "Inha/5_Lectures/2024/Advanced biostatistics/scripts/BTE3207_Advanced_Biostatistics/dataset/sbp_dataset_korea_2013-2014.csv")



head(dataset_sbp)

```


Making a new variable `hypertension`

```{r}

dataset_sbp$hypertension <- ifelse(dataset_sbp$SBP > 130 |
                                           dataset_sbp$DBP > 80,
                                   T,
                                   F)


dataset_sbp$history_of_hypertension <- ifelse(dataset_sbp$DIS == 1 |
                                           dataset_sbp$DIS == 2,
                                   T,
                                   F)

dataset_sbp$DIS <- factor(dataset_sbp$DIS, 
                          levels = c(1,2,3,4),
                          labels = c("Hypertension and diabetes",
                                     "Hypertension",
                                     "Diabetes",
                                     "No history"))

set.seed(1)

dataset_sbp_small <- subset(dataset_sbp, row.names(dataset_sbp) %in% sample(x = 1:1000000, size = 1000))




```


## DIS groups


```{r}

ggplot(dataset_sbp, aes(y = SBP, x = DIS)) +
        geom_boxplot() +
        xlab("History of hypertension or diabetes") +
        ylab("SBP (mmHg)") +
        theme_classic(base_size = 12)
        


```

# T-test?

DIS group 1 vs 4

```{r}

t.test(dataset_sbp %>% filter(as.numeric(DIS) == 1) %>% .$SBP,
       dataset_sbp %>% filter(as.numeric(DIS) == 4) %>% .$SBP,
       var.equal = F)
```


DIS group 2 vs 4

```{r}
t.test(dataset_sbp %>% filter(as.numeric(DIS) == 2) %>% .$SBP,
       dataset_sbp %>% filter(as.numeric(DIS) == 4) %>% .$SBP,
       var.equal = F)
```


DIS group 3 vs 4

```{r}

t.test(dataset_sbp %>% filter(as.numeric(DIS) == 3) %>% .$SBP,
       dataset_sbp %>% filter(as.numeric(DIS) == 4) %>% .$SBP,
       var.equal = F)
```


DIS group 1 vs 2

```{r}

t.test(dataset_sbp %>% filter(as.numeric(DIS) == 1) %>% .$SBP,
       dataset_sbp %>% filter(as.numeric(DIS) == 2) %>% .$SBP,
       var.equal = F)
```

....

Do you think it is reasonable?

# Linear model?

```{r}

lm(data = dataset_sbp, SBP ~ DIS) %>% 
        anova


aov(data = dataset_sbp, SBP ~ DIS) %>% summary


```

It shows how different whether the group difference estimates with beta values (as well as CIs and p-values).

But can we assess wheter the `DIS` is a significant factor affecting SBP?

Here comes the analysis of variance.

# Analysis of variance

Let's say we are comparing multipel groups

Making a new dataset 

```{r}

midterm <- data.frame(midterm = c(82, 83, 97, 83, 78, 68, 38, 59, 55),
           major = rep(c("BE", "MBE", "CHEM"), times=1, each=3))


```

```{r}
mean_sim <- 10
std_sim <- 5

lcb <- ((mean_sim - (3 * std_sim)) - 5)
ucb <- (((2 * mean_sim) + (3 * (2 * std_sim))) + 5)

ggplot(data = data.frame(u = c(lcb, ucb)),
       mapping = aes(x = u)) +
  stat_function(mapping = aes(colour = "Group 1"),
                fun = dnorm,
                args = list(mean = 20,
                            sd = std_sim)) +
  stat_function(mapping = aes(colour = "Group 2"),
                fun = dnorm,
                args = list(mean = 40,
                            sd = (std_sim))) +
        stat_function(mapping = aes(colour = "Group 3"),
                fun = dnorm,
                args = list(mean = 15,
                            sd = (std_sim))) +
  scale_colour_manual(values = c("red", "blue", "green")) +
  labs(x = "Values",
       y = "Densities",
       col = "Groups") +
        theme_classic(base_size = 12)




```
Variance of *group means*
 
```{r}

mean_sim <- 10
std_sim <- 5

lcb <- ((mean_sim - (3 * std_sim)) - 5)
ucb <- (((2 * mean_sim) + (3 * (2 * std_sim))) + 5)

ggplot(data = data.frame(u = c(lcb, ucb)),
       mapping = aes(x = u)) +
  stat_function(mapping = aes(colour = "Group 1"),
                fun = dnorm,
                args = list(mean = 20,
                            sd = std_sim)) +
  stat_function(mapping = aes(colour = "Group 2"),
                fun = dnorm,
                args = list(mean = 40,
                            sd = (std_sim))) +
        stat_function(mapping = aes(colour = "Group 3"),
                fun = dnorm,
                args = list(mean = 15,
                            sd = (std_sim))) +
        geom_vline(xintercept = c(20, 40, 15), color= c("red", "blue", "green")) +
        scale_colour_manual(values = c("grey", "grey", "grey")) +
        geom_vline(xintercept = mean(c(15, 40, 20)), col = "Black", size = 2) +
  labs(x = "Values",
       y = "Densities",
       col = "Groups") +
        theme_classic(base_size = 12)



```
# Uncertainty - variance within groups

## with smaller variance within group

```{r}


mean_sim <- 10
std_sim <- 5

lcb <- ((mean_sim - (3 * std_sim)) - 5)
ucb <- (((2 * mean_sim) + (3 * (2 * std_sim))) + 5)

ggplot(data = data.frame(u = c(lcb, ucb)),
       mapping = aes(x = u)) +
  stat_function(mapping = aes(colour = "Group 1"),
                fun = dnorm,
                args = list(mean = 20,
                            sd = 0.4*std_sim)) +
  stat_function(mapping = aes(colour = "Group 2"),
                fun = dnorm,
                args = list(mean = 40,
                            sd = 0.4*(std_sim))) +
        stat_function(mapping = aes(colour = "Group 3"),
                fun = dnorm,
                args = list(mean = 15,
                            sd = 0.4*(std_sim))) +
        geom_vline(xintercept = c(20, 40, 15), color= c("red", "blue", "green")) +
        scale_colour_manual(values = c("grey", "grey", "grey")) +
        geom_vline(xintercept = mean(c(15, 40, 20)), col = "Black", size = 2) +
  labs(x = "Values",
       y = "Densities",
       col = "Groups") +
        theme_classic(base_size = 12)



```

## Extremely smalle variance

```{r}


mean_sim <- 10
std_sim <- 5

lcb <- ((mean_sim - (3 * std_sim)) - 5)
ucb <- (((2 * mean_sim) + (3 * (2 * std_sim))) + 5)

ggplot(data = data.frame(u = c(lcb, ucb)),
       mapping = aes(x = u)) +
  stat_function(mapping = aes(colour = "Group 1"),
                fun = dnorm,
                args = list(mean = 20,
                            sd = 0.01*std_sim)) +
  stat_function(mapping = aes(colour = "Group 2"),
                fun = dnorm,
                args = list(mean = 40,
                            sd = 0.01*(std_sim))) +
        stat_function(mapping = aes(colour = "Group 3"),
                fun = dnorm,
                args = list(mean = 15,
                            sd = 0.01*(std_sim))) +
        geom_vline(xintercept = c(20, 40, 15), color= c("red", "blue", "green")) +
        scale_colour_manual(values = c("grey", "grey", "grey")) +
        geom_vline(xintercept = mean(c(15, 40, 20)), col = "Black", size = 2) +
  labs(x = "Values",
       y = "Densities",
       col = "Groups") +
        theme_classic(base_size = 12)



```
With extremly high sd

```{r}


mean_sim <- 10
std_sim <- 5

lcb <- ((mean_sim - (3 * std_sim)) - 5)
ucb <- (((2 * mean_sim) + (3 * (2 * std_sim))) + 5)

ggplot(data = data.frame(u = c(lcb, ucb)),
       mapping = aes(x = u)) +
  stat_function(mapping = aes(colour = "Group 1"),
                fun = dnorm,
                args = list(mean = 20,
                            sd = 10*std_sim)) +
  stat_function(mapping = aes(colour = "Group 2"),
                fun = dnorm,
                args = list(mean = 40,
                            sd = 10*(std_sim))) +
        stat_function(mapping = aes(colour = "Group 3"),
                fun = dnorm,
                args = list(mean = 15,
                            sd = 10*(std_sim))) +
        geom_vline(xintercept = c(20, 40, 15), color= c("red", "blue", "green")) +
        scale_colour_manual(values = c("grey", "grey", "grey")) +
        geom_vline(xintercept = mean(c(15, 40, 20)), col = "Black", size = 2) +
        ylim(c(0,0.01)) +
  labs(x = "Values",
       y = "Densities",
       col = "Groups") +
        theme_classic(base_size = 12)



```

Looking at extremecases clearifies what analysis of variance is testing.

## Midterm dataset

```{r}


ggplot(data = midterm) +
  stat_function(mapping = aes(colour = "BE"),
                fun = dnorm,
                args = list(mean = midterm %>% filter(major == "BE") %>% .$midterm %>% mean,
                            sd = midterm %>% filter(major == "BE") %>% .$midterm %>% sd)) +
  stat_function(mapping = aes(colour = "MBE"),
                fun = dnorm,
                args = list(mean = midterm %>% filter(major == "MBE") %>% .$midterm %>% mean,
                            sd = midterm %>% filter(major == "MBE") %>% .$midterm %>% sd)) +
        stat_function(mapping = aes(colour = "CHEM"),
                fun = dnorm,
                args = list(mean = midterm %>% filter(major == "CHEM") %>% .$midterm %>% mean,
                            sd = midterm %>% filter(major == "CHEM") %>% .$midterm %>% sd)) +
  scale_color_brewer(type = "qual", palette = 2) +
        xlim(c(0, 100)) +
  labs(x = "Midterm score",
       y = "Densities",
       col = "Groups") +
        theme_classic(base_size = 12)




```


## SBP dataset


In histogram, the data can be shown as

```{r}

ggplot(data = dataset_sbp) +
  stat_function(mapping = aes(colour = "History of hypertension and diabetes"),
                fun = dnorm,
                args = list(mean = dataset_sbp %>% filter(as.numeric(DIS) == 1) %>% .$SBP %>% mean,
                            sd = dataset_sbp %>% filter(as.numeric(DIS) == 1) %>% .$SBP %>% sd)) +
  stat_function(mapping = aes(colour = "History of hypertension"),
                fun = dnorm,
                args = list(mean = dataset_sbp %>% filter(as.numeric(DIS) == 2) %>% .$SBP %>% mean,
                            sd = dataset_sbp %>% filter(as.numeric(DIS) == 2) %>% .$SBP %>% sd)) +
        stat_function(mapping = aes(colour = "History of diabetes"),
                fun = dnorm,
                args = list(mean = dataset_sbp %>% filter(as.numeric(DIS) == 3) %>% .$SBP %>% mean,
                            sd = dataset_sbp %>% filter(as.numeric(DIS) == 3) %>% .$SBP %>% sd)) +
        stat_function(mapping = aes(colour = "No history"),
                fun = dnorm,
                args = list(mean = dataset_sbp %>% filter(as.numeric(DIS) == 4) %>% .$SBP %>% mean,
                            sd = dataset_sbp %>% filter(as.numeric(DIS) == 4) %>% .$SBP %>% sd)) +
  scale_color_brewer(type = "qual", palette = 2) +
        xlim(c(100,150)) +
  labs(x = "SBP (mmHg)",
       y = "Densities",
       col = "Groups") +
        theme_classic(base_size = 12)




```

# ANOVA in R

`aov()` function can calculate the result of analysis of variance

```{r}


aov(midterm ~ major, data = midterm) %>%
        summary



aov(SBP ~ DIS, data = dataset_sbp_small) %>%
        summary



```

Else, we can implement anova into effect modifications


```{r}

aov(SBP ~ SEX + BTH_G * BMI + FBS + DIS * DBP, data = dataset_sbp_small) %>%
        summary



lm(SBP ~ SEX + BTH_G * BMI + FBS + DIS * DBP, data = dataset_sbp_small) %>%
        summary

```


```{r}

aov(SBP ~  BTH_G * SEX * DIS * DBP, data = dataset_sbp_small) %>%
        summary


```

`anova()` function accepts linear model outputs


```{r}


lm(SBP ~ BTH_G * SEX * DIS * DBP, data = dataset_sbp_small) %>% 
        summary


```
https://stats.idre.ucla.edu/

https://rpubs.com/rslbliss/r_mlm_ws

https://methods101.com.au/docs/soci832_11_3_multilevel_models_week_11/


Dataset is students from 10 handpicked schools, representing a subset of students and schools from a US survey of eight-grade students at 1000 schools (800 public 200 private).

There are quite a lot of variables in the dataset, but we are going start by using just three:

# Bibliography

```{r warning=FALSE, message=FALSE, echo=FALSE}
#===============================================================================
#BTC.LineZero.Footer.1.1.0
#===============================================================================
#R markdown citation generator.
#===============================================================================
#RLB.Dependencies:
#   magrittr, pacman, stringr
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#BTC.Dependencies:
#   LineZero.Header
#===============================================================================
#Generates citations for each explicitly loaded library.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
str_libraries <- c("r", str_libraries)
for (str_libraries in str_libraries) {
        str_libraries |>
                pacman::p_citation() |>
                print(bibtex = FALSE) |>
                capture.output() %>%
                .[-1:-3] %>% .[. != ""] |>
                stringr::str_squish() |>
                stringr::str_replace("_", "") |>
                cat()
        cat("\n")
}
#===============================================================================
```
