---
title: "COD_week5_1_MGK_BTE3207"
author: "Minsik Kim"
date: "2024-09-30"
output:
        rmdformats::downcute:
        downcute_theme: "chaos"
code_folding: show
fig_width: 6
fig_height: 6
df_print: paged
editor_options: 
        chunk_output_type: inline
markdown: 
        wrap: 72
---
        
```{r warning=FALSE, message=FALSE, echo=FALSE, results='hide', setup}
#===============================================================================
#BTC.LineZero.Header.1.1.0
#===============================================================================
#R Markdown environment setup and reporting utility.
#===============================================================================
#RLB.Dependencies:
#   knitr, magrittr, pacman, rio, rmarkdown, rmdformats, tibble, yaml
#===============================================================================
#Input for document parameters, libraries, file paths, and options.
#=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=
knitr::opts_chunk$set(message=FALSE, warning = FALSE)


path_working <- 
        ifelse(sessionInfo()[1]$R.version$platform == "x86_64-pc-linux-gnu",
               "/mnt/4T_samsung/Dropbox/",
               ifelse(sessionInfo()[1]$R.version$platform == "aarch64-apple-darwin20",
                      "/Volumes/macdrive/Dropbox/", 
                      "/Users/minsikkim/Dropbox (Personal)/"))
path_library <- 
        ifelse(sessionInfo()[1]$R.version$platform == "x86_64-pc-linux-gnu",
               "/home/bagel/R_lib/",
               "/Library/Frameworks/R.framework/Resources/library")


str_libraries <- c("tidyverse", "reactable", "yaml", "pacman")


YAML_header <-
        '---
title: "BTE3207 week 5-1"
author: "Minsik Kim"
date: "2024-09-30"
output:
    rmdformats::downcute:
        downcute_theme: "chaos"
        code_folding: hide
        fig_width: 6
        fig_height: 6
---'
#seed <- "20230920"

#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Loads libraries, file paths, and other document options.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Boot <- function() {
        .libPaths(path_library)
        
        require(pacman)
        pacman::p_load(c("knitr", "rmarkdown", "rmdformats", "yaml"))
        
        knitr::opts_knit$set(root.dir = path_working)
        
        str_libraries |> unique() |> sort() -> str_libraries
        pacman::p_load(char = str_libraries)
        
        #set.seed(seed)
}
FUN.LineZero.Boot()
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Outputs R environment report.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Report <- function() {
        cat("Line Zero Environment:\n\n")
        paste("R:", pacman::p_version(), "\n") |> cat()
        cat("Libraries:\n")
        for (str_libraries in str_libraries) {
                paste(
                        "    ", str_libraries, ": ", pacman::p_version(package = str_libraries),
                        "\n", sep = ""
                ) |> cat()
        }
        paste("\nOperating System:", pacman::p_detectOS(), "\n") |> cat()
        paste("    Library Path:", path_library, "\n") |> cat()
        paste("    Working Path:", path_working, "\n") |> cat()
        #paste("Seed:", seed, "\n\n") |> cat()
        cat("YAML Header:\n")
        cat(YAML_header)
}
FUN.LineZero.Report()

```

# Before begin..

Let's load the SBP dataset.

```{r}
dataset_sbp <- read.csv(file = "Inha/5_Lectures/2024/Advanced biostatistics/scripts/BTE3207_Advanced_Biostatistics/dataset/sbp_dataset_korea_2013-2014.csv")

head(dataset_sbp)

```

# Degree of freedom

With smaller sampling number, the variance gets smaller

```{r}

set.seed(1)
replicate(5000, sample(dataset_sbp$SBP, 150)) %>%
        data.frame() %>% 
        colMeans() %>%
        hist(100,
             main = "Distribution of 5000 sample mean p-hat of hypertation from 1M subjects.\nEach sample N = 150") +
        stat_function(fun = dnorm,
                      args = list(mean = mean(dataset_sbp$SBP),
                                  sd = sd(dataset_sbp$SBP)))

dataset_sbp$SBP %>%
        data.frame() %>% 
        colMeans() %>%
        hist(100,
             main = "Distribution of total sample mean p-hat of hypertation from 1M subjects.\nEach sample N = 150") +
        stat_function(fun = dnorm,
                      args = list(mean = mean(dataset_sbp$SBP),
                                  sd = sd(dataset_sbp$SBP)))




set.seed(1)
ggplot(data = dataset_sbp) +
        geom_histogram(aes(x = SBP,
                           y = ..density..),
                       #binwidth = 0.01,
                       fill = "white",
                       col = "black") +
        ggtitle("Histogram of \nSBP of 1M Koreans") +
        ylab("Count") +
        xlim(c(75, 180)) +
        theme_classic(base_size = 20, base_family = "serif") +
        xlab("Normalized SBP (mmHg)") +
        stat_function(fun = dnorm, args = list(mean = mean(dataset_sbp$SBP),
                                               sd = sd(dataset_sbp$SBP)))

set.seed(1)

ggplot(data = dataset_sbp %>% sample_n(50)) +
        geom_histogram(aes(x = SBP,
                           y = ..density..),
                       #binwidth = 0.01,
                       fill = "white",
                       col = "black") +
        ggtitle("Histogram of \nSBP of random 50 Koreans") +
        ylab("Count") +
        xlim(c(75, 180)) +
        theme_classic(base_size = 20, base_family = "serif") +
        xlab("Normalized SBP (mmHg)") +
        stat_function(fun = dnorm, args = list(mean = mean(dataset_sbp$SBP),
                                               sd = sd(dataset_sbp$SBP)))



set.seed(1)

ggplot(data = dataset_sbp %>% sample_n(30)) +
        geom_histogram(aes(x = SBP,
                           y = ..density..),
                       #binwidth = 0.01,
                       fill = "white",
                       col = "black") +
        ggtitle("Histogram of \nSBP of random 30 Koreans") +
        ylab("Count") +
        xlim(c(75, 180)) +
        theme_classic(base_size = 20, base_family = "serif") +
        xlab("Normalized SBP (mmHg)") +
        stat_function(fun = dnorm, args = list(mean = mean(dataset_sbp$SBP),
                                               sd = sd(dataset_sbp$SBP)))



set.seed(1)

ggplot(data = dataset_sbp %>% sample_n(20)) +
        geom_histogram(aes(x = SBP,
                           y = ..density..),
                       #binwidth = 0.01,
                       fill = "white",
                       col = "black") +
        ggtitle("Histogram of \nSBP of random 20 Koreans") +
        ylab("Count") +
        xlim(c(75, 180)) +
        theme_classic(base_size = 20, base_family = "serif") +
        xlab("Normalized SBP (mmHg)") +
        stat_function(fun = dnorm, args = list(mean = mean(dataset_sbp$SBP),
                                               sd = sd(dataset_sbp$SBP)))



set.seed(1)

ggplot(data = dataset_sbp %>% sample_n(10)) +
        geom_histogram(aes(x = SBP,
                           y = ..density..),
                       #binwidth = 0.01,
                       fill = "white",
                       col = "black") +
        ggtitle("Histogram of \nSBP of random 10 Koreans") +
        ylab("Count") +
        xlim(c(75, 180)) +
        theme_classic(base_size = 20, base_family = "serif") +
        xlab("Normalized SBP (mmHg)") +
        stat_function(fun = dnorm, args = list(mean = mean(dataset_sbp$SBP),
                                               sd = sd(dataset_sbp$SBP)))



set.seed(1)

ggplot(data = dataset_sbp %>% sample_n(5)) +
        geom_histogram(aes(x = SBP,
                           y = ..density..),
                       #binwidth = 0.01,
                       fill = "white",
                       col = "black") +
        ggtitle("Histogram of \nSBP of random 5 Koreans") +
        ylab("Count") +
        xlim(c(75, 180)) +
        theme_classic(base_size = 20, base_family = "serif") +
        xlab("Normalized SBP (mmHg)") +
        stat_function(fun = dnorm, args = list(mean = mean(dataset_sbp$SBP),
                                               sd = sd(dataset_sbp$SBP)))



set.seed(1)
ggplot(data = dataset_sbp %>% sample_n(2)) +
        geom_histogram(aes(x = SBP,
                           y = ..density..),
                       #binwidth = 0.01,
                       fill = "white",
                       col = "black") +
        ggtitle("Histogram of \nSBP of random 2 Koreans") +
        ylab("Count") +
        xlim(c(75, 180)) +
        theme_classic(base_size = 20, base_family = "serif") +
        xlab("Normalized SBP (mmHg)") +
        stat_function(fun = dnorm, args = list(mean = mean(dataset_sbp$SBP),
                                               sd = sd(dataset_sbp$SBP)))



```


# Comparing data

To compare two different sample groups, we can caculate the difference between samples and look at CIs of **difference between them**. The schematic diagram of how this process works is listed in below figures. **Note: this is not a exact statistical test but drwan to help understanding the basic concept of it**

```{r}
dataset_sbp$hypertension <- ifelse(dataset_sbp$SBP > 130 | dataset_sbp$DBP > 80,
                                   1,
                                   0)
set.seed(1)
ggplot(data = dataset_sbp %>% sample_n(1000)) +
        geom_histogram(aes(x = SBP,
                           y = ..density..),
                       #binwidth = 0.01,
                       #fill = "white",
                       #col = "black",
                       alpha = 0.8) +
        ggtitle("Histogram of SBP") +
        ylab("Proportion") +
        xlim(c(75, 180)) +
        theme_classic(base_size = 15, base_family = "serif") +
        guides(fill = guide_legend(title = "Hypertension")) +
        xlab("SBP (mmHg)")
set.seed(1)
ggplot(data = dataset_sbp %>% sample_n(1000)) +
        geom_histogram(aes(x = SBP,
                           y = ..density..,
                           fill=factor(hypertension)),
                       #binwidth = 0.01,
                       #fill = "white",
                       #col = "black",
                       alpha = 0.4) +
        ggtitle("Histogram of SBP by two group") +
        ylab("Proportion") +
        xlim(c(75, 180)) +
        theme_classic(base_size = 15, base_family = "serif") +
        guides(fill = guide_legend(title = "Hypertension")) +
        xlab("SBP (mmHg)") +
set.seed(1)
ggplot(data = dataset_sbp %>% sample_n(1000)) +
        geom_histogram(aes(x = SBP,
                           y = ..density..,
                           fill=factor(hypertension)),
                       #binwidth = 0.01,
                       #fill = "white",
                       #col = "black",
                       alpha = 0.2) +
        ggtitle("Histogram of SBP by two group with normal curve") +
        ylab("Proportion") +
        xlim(c(75, 180)) +
        theme_classic(base_size = 15, base_family = "serif") +
        guides(fill = guide_legend(title = "Hypertension")) +
        xlab("SBP (mmHg)") +
        stat_function(fun = dnorm,
                          args = list(mean = mean(subset(dataset_sbp, hypertension == 1)$SBP),
                                      sd = sd(subset(dataset_sbp, hypertension == 1)$SBP)),
                      col = "blue") +
                stat_function(fun = dnorm,
                          args = list(mean = mean(subset(dataset_sbp, hypertension == 0)$SBP),
                                      sd = sd(subset(dataset_sbp, hypertension == 0)$SBP)),
                          col = "red")


set.seed(1)
high <- dataset_sbp %>% sample_n(1000) %>% subset(., hypertension == 1) %>% sample_n(100)
set.seed(1)
low <- dataset_sbp %>% sample_n(1000) %>% subset(., hypertension == 0) %>% sample_n(100)

ggplot() +
        geom_histogram(aes(x = high$SBP-low$SBP,
                           y = ..density..),
                       #binwidth = 0.01,
                       #fill = "white",
                       #col = "black",
                       alpha = 0.2) +
        ggtitle("Histogram of SBP **difference** by two group with normal curve") +
        ylab("Proportion") +
        #xlim(c(75, 180)) +
        theme_classic(base_size = 15, base_family = "serif") +
        theme(plot.title = ggtext::element_markdown()) +
        guides(fill = guide_legend(title = "Hypertension")) +
        xlab("SBP (mmHg)")
        



ggplot() +
        geom_histogram(aes(x = high$SBP-low$SBP,
                           y = ..density..),
                       #binwidth = 0.01,
                       #fill = "white",
                       #col = "black",
                       alpha = 0.2) +
        ggtitle("Histogram of SBP **difference** by two group with normal curve") +
        ylab("Proportion") +
        #xlim(c(75, 180)) +
        theme_classic(base_size = 15, base_family = "serif") +
        theme(plot.title = ggtext::element_markdown()) +
        guides(fill = guide_legend(title = "Hypertension")) +
        xlab("SBP (mmHg)") +
        stat_function(fun = dnorm,
                      args = list(mean = mean(high$SBP-low$SBP),
                                      sd = sd(high$SBP-low$SBP))) 


ggplot() +
        geom_histogram(aes(x = high$SBP-low$SBP,
                           y = ..density..),
                       #binwidth = 0.01,
                       #fill = "white",
                       #col = "black",
                       alpha = 0.2) +
        ggtitle("Histogram of SBP **difference** by two group with normal curve") +
        ylab("Proportion") +
        #xlim(c(75, 180)) +
        theme_classic(base_size = 15, base_family = "serif") +
        theme(plot.title = ggtext::element_markdown()) +
        guides(fill = guide_legend(title = "Hypertension")) +
        xlab("SBP (mmHg)") +
        stat_function(fun = dnorm,
                      args = list(mean = mean(high$SBP-low$SBP),
                                      sd = sd(high$SBP-low$SBP))) +
        geom_vline(aes(xintercept = 0), color = "red", linetype = "dashed", size = 1)

t.test(high$SBP, low$SBP, paired = F)

```


# CIs of sample proportion

The Confidence interval of sample proportion need to employ **exact test**. 

Here is an example how to calculate the CIs for p-hat in R.

```{r}



install.packages("interpretCI")
library(interpretCI)

#Normal approximation (2 sd)
x=propCI(n = 20, p = 0.15, alpha = 0.05)
x

#Normal approximation 
x=propCI(n = 20, p = 0.15, alpha = 0.05, alternative =  "wald")
x

#exact method - Clopper-Pearson interval.
binom.test(x = 20 * 0.15, n = 20, conf.level = 0.95)

```



# Bibliography

```{r warning=FALSE, message=FALSE, echo=FALSE}
#===============================================================================
#BTC.LineZero.Footer.1.1.0
#===============================================================================
#R markdown citation generator.
#===============================================================================
#RLB.Dependencies:
#   magrittr, pacman, stringr
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#BTC.Dependencies:
#   LineZero.Header
#===============================================================================
#Generates citations for each explicitly loaded library.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
str_libraries <- c("r", str_libraries)
for (str_libraries in str_libraries) {
        str_libraries |>
                pacman::p_citation() |>
                print(bibtex = FALSE) |>
                capture.output() %>%
                .[-1:-3] %>% .[. != ""] |>
                stringr::str_squish() |>
                stringr::str_replace("_", "") |>
                cat()
        cat("\n")
}
#===============================================================================
```
