---
title: "COD_week10_2_MGK_BTE3207"
author: "Minsik Kim"
date: "2024-11-06"
output:
        rmdformats::downcute:
        downcute_theme: "chaos"
code_folding: show
fig_width: 6
fig_height: 6
df_print: paged
editor_options: 
        chunk_output_type: inline
markdown: 
        wrap: 72
---
        

<!-- <style> -->
<!--   /* Default light mode styles */ -->
<!--   .reactable { -->
<!--     background-color: #ffffff !important; /* Light background */ -->
<!--     color: #000000 !important;            /* Dark text */ -->
<!--     border-color: #cccccc !important;     /* Light border */ -->
<!--   } -->

<!-- </style> -->
        
```{r warning=FALSE, message=FALSE, echo=FALSE, results='hide', setup}

#===============================================================================
#BTC.LineZero.Header.1.1.0
#===============================================================================
#R Markdown environment setup and reporting utility.
#===============================================================================
#RLB.Dependencies:
#   knitr, magrittr, pacman, rio, rmarkdown, rmdformats, tibble, yaml
#===============================================================================
#Input for document parameters, libraries, file paths, and options.
#=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=
knitr::opts_chunk$set(message=FALSE, warning = FALSE)

path_working <- 
        ifelse(sessionInfo()[1]$R.version$platform == "x86_64-pc-linux-gnu",
               "/mnt/4T_samsung/Dropbox/",
               ifelse(sessionInfo()[1]$R.version$platform == "aarch64-apple-darwin20",
                      "/Volumes/macdrive/Dropbox/", 
                      "/Users/minsikkim/Dropbox (Personal)/"))
path_library <- 
        ifelse(sessionInfo()[1]$R.version$platform == "x86_64-pc-linux-gnu",
               "/home/bagel/R_lib/",
               "/Library/Frameworks/R.framework/Resources/library/")


str_libraries <- c("tidyverse", "pacman", "yaml", "reactable")



YAML_header <-
        '---
title: "BTE3207 week 10-2"
author: "Minsik Kim"
date: "2024.11.06"
output:
    rmdformats::downcute:
        downcute_theme: "chaos"
        code_folding: hide
        fig_width: 6
        fig_height: 6
---'
seed <- "20241106"

#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Loads libraries, file paths, and other document options.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Boot <- function() {
        .libPaths(path_library)
        
        require(pacman)
        pacman::p_load(c("knitr", "rmarkdown", "rmdformats", "yaml"))
        
        knitr::opts_knit$set(root.dir = path_working)
        
        str_libraries |> unique() |> sort() -> str_libraries
        pacman::p_load(char = str_libraries)
        
        set.seed(seed)
}
FUN.LineZero.Boot()
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Outputs R environment report.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Report <- function() {
        cat("Line Zero Environment:\n\n")
        paste("R:", pacman::p_version(), "\n") |> cat()
        cat("Libraries:\n")
        for (str_libraries in str_libraries) {
                paste(
                        "    ", str_libraries, ": ", pacman::p_version(package = str_libraries),
                        "\n", sep = ""
                ) |> cat()
        }
        paste("\nOperating System:", pacman::p_detectOS(), "\n") |> cat()
        paste("    Library Path:", path_library, "\n") |> cat()
        paste("    Working Path:", path_working, "\n") |> cat()
        paste("Seed:", seed, "\n\n") |> cat()
        cat("YAML Header:\n")
        cat(YAML_header)
}
FUN.LineZero.Report()


```


# Multiple linear regression

## Before begin..

Let's load the SBP dataset.

```{r}

dataset_sbp <- read.csv(file = "Inha/5_Lectures/2024/Advanced biostatistics/scripts/BTE3207_Advanced_Biostatistics/dataset/sbp_dataset_korea_2013-2014.csv")



head(dataset_sbp)

```


Making a new variable `hypertension`

```{r}

dataset_sbp$hypertension <- ifelse(dataset_sbp$SBP > 130 |
                                           dataset_sbp$DBP > 80,
                                   T,
                                   F)


dataset_sbp$history_of_hypertension <- ifelse(dataset_sbp$DIS == 1 |
                                           dataset_sbp$DIS == 2,
                                   T,
                                   F)

set.seed(1)

dataset_sbp_small <- subset(dataset_sbp, row.names(dataset_sbp) %in% sample(x = 1:1000000, size = 1000))

dataset_sbp_small$DIS <- as.factor(dataset_sbp_small$DIS)


```

## Simple linear regression

With `lm()` function, we can calculate the linear model of DBP~SBP with the least squares algorithm.

```{r}

#A function that returns equation

lm_eqn <- function(m){
    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2, 
         list(a = format(unname(coef(m)[1]), digits = 2),
              b = format(unname(coef(m)[2]), digits = 2),
             r2 = format(summary(m)$r.squared, digits = 3)))
    as.character(as.expression(eq));
}




ggplot(data = dataset_sbp_small, aes(y = SBP, x = DBP)) +
        geom_point() +
        theme_classic(base_family = "serif", base_size = 20) +
        ylab("SBP (mmHg)") +
        xlab("DBP (mmHg)") + 
        geom_smooth(method = "lm") +
        geom_label(y = 170, x = 68,
                  label = lm_eqn(lm(data = dataset_sbp_small, SBP ~ DBP)),
                  parse = TRUE,
                  family='serif',
                  size = 8, label.size = 0, color = "blue")


lm(data = dataset_sbp_small, SBP ~ DBP) %>% confint()

#rnorm(100, mean = 100, sd = 1)

```

## Multiple regression


## Basic concept {.tabset}


### Deviding samples

In fact, there are two SEXs in the sample. 

```{r}

ggplot(dataset_sbp_small, aes(y = SBP, x = DBP)) +
        #geom_smooth(method = "lm", se = FALSE, color = "blue") +  # Plot regression slope
        #geom_segment(aes(xend = SBP, yend = predicted), alpha = .2) +  # alpha to fade lines
        geom_point(aes(col = history_of_hypertension)) +
        ylab("SBP (mmHg)") +
        xlab("DBP (mmHg)") + 
        theme_classic( base_family = "serif", base_size = 20) +
        theme(legend.position = "top") +
        guides(col=guide_legend(title = "History of hyper tension"))
        
  
```

### Deviding samples

In fact, there are two SEXs in the sample. 

```{r}

ggplot(dataset_sbp_small, aes(y = SBP, x = DBP)) +
        #geom_smooth(method = "lm", se = FALSE, color = "blue") +  # Plot regression slope
        #geom_segment(aes(xend = SBP, yend = predicted), alpha = .2) +  # alpha to fade lines
        geom_point(aes(col = history_of_hypertension)) +
        ylab("SBP (mmHg)") +
        xlab("DBP (mmHg)") +
        facet_wrap(~history_of_hypertension) +
        theme_classic( base_family = "serif", base_size = 20) +
        theme(legend.position = "top",
              plot.title = element_text(color = "Blue", hjust = 0.7, vjust = -23)) +
        guides(col=guide_legend(title = "History of hyper tension")) +
        geom_abline(intercept = 37, slope = 1.1, col = "blue") +
        ggtitle("y = 37 + 1.1 x")
        
  
```
# Multiple linear regression


```{r}

lm(data = dataset_sbp_small, SBP ~ DBP) %>% summary

lm(data = dataset_sbp_small, SBP ~ DBP + history_of_hypertension) %>% summary


```



```{r}
palette()


ggplot(dataset_sbp_small, aes(y = SBP, x = DBP)) +
        #geom_smooth(method = "lm", se = FALSE, color = "blue") +  # Plot regression slope
        #geom_segment(aes(xend = SBP, yend = predicted), alpha = .2) +  # alpha to fade lines
        geom_point(aes(col = history_of_hypertension)) +
        ylab("SBP (mmHg)") +
        xlab("DBP (mmHg)") + 
        theme_classic( base_family = "serif", base_size = 20) +
        theme(legend.position = "top") +
        guides(col=guide_legend(title = "History of hyper tension")) +
        geom_abline(intercept = 40, slope = 1.04, col = "#F8766D") +
        annotate(geom="text", x=120, y=140, label="LM of HT-False",
              color="#F8766D", family = "serif", size = 8) + 
        geom_abline(intercept = 46.42, slope = 1.04, col = "#00BFC4") +
        annotate(geom="text", x=100, y=170, label="LM of HT-True",
              color="#00BFC4", family = "serif", size = 8)

```


### {-}

## Continuous variable as covariate

### Deviding samples



```{r}

ggplot(dataset_sbp_small, aes(y = SBP, x = DBP)) +
        geom_smooth(method = "lm", se = FALSE, color = "blue") +  # Plot regression slope
        #geom_segment(aes(xend = SBP, yend = predicted), alpha = .2) +  # alpha to fade lines
        geom_point() +
        ylab("SBP (mmHg)") +
        xlab("DBP (mmHg)") +
        theme_classic( base_family = "serif", base_size = 20) +
        theme(legend.position = "top",
              plot.title = element_text(color = "Blue", hjust = 0.7, vjust = -23)) +
        guides(col=guide_legend(title = "History of hyper tension")) + 
        geom_label(y = 170, x = 68,
                  label = lm_eqn(lm(data = dataset_sbp_small, SBP ~ DBP)),
                  parse = TRUE,
                  family='serif',
                  size = 8, label.size = 0, color = "blue")


        
  
```


```{r}

ggplot(dataset_sbp_small, aes(y = SBP, x = BTH_G)) +
        geom_smooth(method = "lm", se = FALSE, color = "blue") +  # Plot regression slope
        #geom_segment(aes(xend = SBP, yend = predicted), alpha = .2) +  # alpha to fade lines
        geom_point() +
        ylab("SBP (mmHg)") +
        xlab("Age group") +
        theme_classic( base_family = "serif", base_size = 20) +
        theme(legend.position = "top",
              plot.title = element_text(color = "Blue", hjust = 0.7, vjust = -23)) +
        guides(col=guide_legend(title = "History of hyper tension")) + 
        geom_label(y = 170, x = 18,
                  label = lm_eqn(lm(data = dataset_sbp_small, SBP ~ BTH_G)),
                  parse = TRUE,
                  family='serif',
                  size = 8, label.size = 0, color = "blue")

  
```


```{r}

lm(data = dataset_sbp, formula = SBP ~ DBP)


model_result <- lm(data = dataset_sbp, formula = SBP ~ DBP)

dataset_sbp$DIS <- as.factor(dataset_sbp$DIS)
dataset_sbp %>% summary

dataset_sbp %>% 
        lm(formula = SBP ~ DBP + DIS + DBP * DIS) %>%
        summary

``` 



```{r}


dataset_sbp_small %>%
        lm(SBP ~ DBP + BTH_G, data = .) %>% 
        summary

dataset_sbp_small %>%
        lm(SBP ~ DBP + SEX, data = .) %>% 
        summary

car::avPlots(dataset_sbp_small %>%
        lm(SBP ~ DBP + history_of_hypertension, data = .))

car::avPlots(dataset_sbp_small %>%
        lm(SBP ~ DBP + history_of_hypertension, data = .))


```

```{r}

dataset_sbp_small %>%
        lm(SBP ~ SEX, data = .) %>% 
        summary


dataset_sbp_small %>%
        lm(SBP ~ history_of_hypertension, data = .) %>% 
        summary

        
dataset_sbp_small %>%
        lm(SBP ~ DBP + history_of_hypertension, data = .) %>% 
        summary

dataset_sbp_small %>%
        lm(SBP ~ DBP + history_of_hypertension, data = .) %>%
        car::avPlots()
        
dataset_sbp_small %>%
        ggplot(., aes(y = SBP, x = history_of_hypertension)) +
        geom_jitter()
  
```

# Multiple linear regression


```{r}

lm(data = dataset_sbp_small, SBP ~ DBP) %>% summary

lm(data = dataset_sbp_small, SBP ~ DBP + history_of_hypertension) %>% summary


```



```{r}
palette()

ggplot(dataset_sbp_small, aes(y = SBP, x = DBP)) +
        #geom_smooth(method = "lm", se = FALSE, color = "blue") +  # Plot regression slope
        #geom_segment(aes(xend = SBP, yend = predicted), alpha = .2) +  # alpha to fade lines
        geom_point(aes(col = history_of_hypertension)) +
        ylab("SBP (mmHg)") +
        xlab("DBP (mmHg)") + 
        theme_classic( base_family = "serif", base_size = 20) +
        theme(legend.position = "top") +
        guides(col=guide_legend(title = "History of hyper tension")) +
        geom_abline(intercept = 40, slope = 1.04, col = "#F8766D") +
        annotate(geom="text", x=120, y=140, label="LM of HT-False",
              color="#F8766D", family = "serif", size = 8) + 
        geom_abline(intercept = 46.42, slope = 1.04, col = "#00BFC4") +
        annotate(geom="text", x=100, y=170, label="LM of HT-True",
              color="#00BFC4", family = "serif", size = 8)

```




### Simulation





# Bibliography

```{r warning=FALSE, message=FALSE, echo=FALSE}
#===============================================================================
#BTC.LineZero.Footer.1.1.0
#===============================================================================
#R markdown citation generator.
#===============================================================================
#RLB.Dependencies:
#   magrittr, pacman, stringr
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#BTC.Dependencies:
#   LineZero.Header
#===============================================================================
#Generates citations for each explicitly loaded library.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
str_libraries <- c("r", str_libraries)
for (str_libraries in str_libraries) {
        str_libraries |>
                pacman::p_citation() |>
                print(bibtex = FALSE) |>
                capture.output() %>%
                .[-1:-3] %>% .[. != ""] |>
                stringr::str_squish() |>
                stringr::str_replace("_", "") |>
                cat()
        cat("\n")
}
#===============================================================================
```
