---
title: "COD_week4_2_MGK_BTE3207"
author: "Minsik Kim"
date: "2024-09-27"
output:
rmdformats::downcute:
downcute_theme: "chaos"
code_folding: show
fig_width: 6
fig_height: 6
df_print: paged
editor_options: 
        chunk_output_type: inline
markdown: 
        wrap: 72
---
        
```{r warning=FALSE, message=FALSE, echo=FALSE, results='hide', setup}
#===============================================================================
#BTC.LineZero.Header.1.1.0
#===============================================================================
#R Markdown environment setup and reporting utility.
#===============================================================================
#RLB.Dependencies:
#   knitr, magrittr, pacman, rio, rmarkdown, rmdformats, tibble, yaml
#===============================================================================
#Input for document parameters, libraries, file paths, and options.
#=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=
knitr::opts_chunk$set(message=FALSE, warning = FALSE)


path_working <- 
        ifelse(sessionInfo()[1]$R.version$platform == "x86_64-pc-linux-gnu",
               "/mnt/4T_samsung/Dropbox/",
               ifelse(sessionInfo()[1]$R.version$platform == "aarch64-apple-darwin20",
                      "/Volumes/macdrive/Dropbox/", 
                      ""))
path_library <- 
        ifelse(sessionInfo()[1]$R.version$platform == "x86_64-pc-linux-gnu",
               "/home/bagel/R_lib/",
               "/Library/Frameworks/R.framework/Resources/library")



str_libraries <- c("tidyverse", "reactable", "yaml", "pacman")

YAML_header <-
        '---
title: "BTE3207 week 4-2"
author: "Minsik Kim"
date: "2024.09.27"
output:
    rmdformats::downcute:
        downcute_theme: "chaos"
        code_folding: hide
        fig_width: 6
        fig_height: 6
---'


#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Loads libraries, file paths, and other document options.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Boot <- function() {
        .libPaths(path_library)
        
        require(pacman)
        pacman::p_load(c("knitr", "rmarkdown", "rmdformats", "yaml"))
        
        knitr::opts_knit$set(root.dir = path_working)
        
        str_libraries |> unique() |> sort() -> str_libraries
        pacman::p_load(char = str_libraries)
        
        #set.seed(seed)
}
FUN.LineZero.Boot()
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Outputs R environment report.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Report <- function() {
        cat("Line Zero Environment:\n\n")
        paste("R:", pacman::p_version(), "\n") |> cat()
        cat("Libraries:\n")
        for (str_libraries in str_libraries) {
                paste(
                        "    ", str_libraries, ": ", pacman::p_version(package = str_libraries),
                        "\n", sep = ""
                ) |> cat()
        }
        paste("\nOperating System:", pacman::p_detectOS(), "\n") |> cat()
        paste("    Library Path:", path_library, "\n") |> cat()
        paste("    Working Path:", path_working, "\n") |> cat()
        #paste("Seed:", seed, "\n\n") |> cat()
        cat("YAML Header:\n")
        cat(YAML_header)
}
FUN.LineZero.Report()

```

# Before begin..

Let's load the SBP dataset.

```{r}
dataset_sbp <- read.csv(file = "Inha/5_Lectures/2024/Advanced biostatistics/scripts/BTE3207_Advanced_Biostatistics/dataset/sbp_dataset_korea_2013-2014.csv")

head(dataset_sbp)

```


# The normal distribution


```{r}
set.seed(1)
sim_100_5 <- rnorm(1000, mean = 100, sd = 5) #makes a data with 1000 elements having mean of 100 and sd of 5

hist(sim_100_5, 100, probability = T) #how does it look like?
curve(dnorm(x, mean=100, sd=5), 
      col="darkblue", lwd=2, add=TRUE, yaxt="n")

mean(sim_100_5) 
sd(sim_100_5)

```

# normal distribution

Let's compare some data collected from real people (random 100 Koreans) with the normal distribution

```{r}

set.seed(1)
sbp_10000 <- sample(dataset_sbp$SBP, 10000) %>%
        data.frame()

hist_sbp_10000 <- ggplot(data = sbp_10000) +
        geom_histogram(aes(x = .,
                           y =..density..),
                       bins = 15,
                       fill = "white",
                       col = "black") +
        ggtitle("Histogram of SBP of 10000 Koreans") +
        ylab("Probability") +
        xlab("SBP (mmHg)")

hist_sbp_10000

hist_sbp_10000 + 
        stat_function(fun = dnorm, args = list(mean = mean(sbp_10000$.), sd = sd(sbp_10000$.)))

sbp_10000$. %>% mean

sbp_10000$. %>% sd

sbp_10000$. %>% median

```

# z-score

How can we say the location of a person with 130 mmHg of SBP, in terms of distribution? We usually calculate z-score, to understand the location of one sample more easilty.

```{r}
z_sbp_10000 <- (131-mean(sbp_10000$.))/sd(sbp_10000$.)

z_sbp_10000
```

Here, the z-score of a person with SBP of 131 is 0.63.

Using this information, we cancalculate the percentile of him (assuming the SBP dataset is following the normal distribution)

```{r}

pnorm(z_sbp_10000)


```

It's percentile is 73.6%. 

In other words,

1. 29% of samples are having greater value than 130

2. 71% of samples are having smaller SBP value than 130

Meanwhile, the actural 73.6th percentile of our dataset (sbp_10000) was 130. Why?

```{r}
quantile(sbp_10000$., 0.736)
```

#Normalization

min-max normalization makes minimum vale to the 0 and maximum value to 1.

```{r}

dataset_sbp$norm_SBP <- (dataset_sbp$SBP - min(dataset_sbp$SBP))/
        (max(dataset_sbp$SBP) - min(dataset_sbp$SBP))


ggplot(data = dataset_sbp) +
        geom_histogram(aes(x = SBP,
                           y = ..count..),
                       binwidth = 0.01,
                       fill = "white",
                       col = "black") +
        ggtitle("Histogram of SBP of 10000 Koreans") +
        ylab("Count") +
        theme_classic(base_size = 20, base_family = "serif") +
        xlab("SBP (mmHg)")



ggplot(data = dataset_sbp) +
        geom_histogram(aes(x = norm_SBP,
                           y = ..count..),
                       binwidth = 0.01,
                       fill = "white",
                       col = "black") +
        ggtitle("Histogram of Min-Max Normalized\nSBP of 10000 Koreans") +
        ylab("Count") +
        theme_classic(base_size = 20, base_family = "serif") +
        xlab("Normalized SBP (mmHg)")
#        stat_function(fun = dnorm, args = list(mean = mean(dataset_sbp$norm_SBP),
#                                               sd = sd(dataset_sbp$norm_SBP)))


```

# Standardization

Standardization makes mean to 0 and SD into 1.

```{r}
dataset_sbp$std_SBP <- (dataset_sbp$SBP - mean(dataset_sbp$SBP))/
        sd(dataset_sbp$SBP)

ggplot(data = dataset_sbp) +
        geom_histogram(aes(x = std_SBP,
                           y = ..count..),
                       binwidth = 0.3,
                       fill = "white",
                       col = "black") +
        ggtitle("Histogram of standardized\nSBP of 10000 Koreans") +
        ylab("Probability") +
        xlab("Standardized SBP (mmHg)") +
        theme_classic(base_family = "serif", base_size = 20) 
        #stat_function(fun = dnorm, args = list(mean = mean(dataset_sbp$std_SBP), sd = sd(dataset_sbp$std_SBP)))

```


# Skewed data and z-score

We can calculate z-score of skewed data, but their calculation does not match with our assumption "the data is approximately normal", therefore it results in wrong predictions.


```{r}

dataset_sbp$FBS %>%
        hist(probability = T, main = "Histogram of blood sugar level (FBS)", 15)


dataset_sbp$FBS %>% mean


dataset_sbp$FBS %>% sd

dataset_sbp$FBS %>% median

```

So, if we use z-score to guess the 2.5% percentile's FBS level,

```{r}

(dataset_sbp$FBS %>% mean) - 2 * (dataset_sbp$FBS %>% sd)


```

However, in actual database, 

```{r}
quantile(dataset_sbp$FBS, 0.025)
```

Thee z-score cannot represent the actual dataset.

# Bibliography

```{r warning=FALSE, message=FALSE, echo=FALSE}
#===============================================================================
#BTC.LineZero.Footer.1.1.0
#===============================================================================
#R markdown citation generator.
#===============================================================================
#RLB.Dependencies:
#   magrittr, pacman, stringr
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#BTC.Dependencies:
#   LineZero.Header
#===============================================================================
#Generates citations for each explicitly loaded library.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
str_libraries <- c("r", str_libraries)
for (str_libraries in str_libraries) {
        str_libraries |>
                pacman::p_citation() |>
                print(bibtex = FALSE) |>
                capture.output() %>%
                .[-1:-3] %>% .[. != ""] |>
                stringr::str_squish() |>
                stringr::str_replace("_", "") |>
                cat()
        cat("\n")
}
#===============================================================================
```
