---
title: "COD_2025_week5_1_MGK_BTE3207"
author: "Minsik Kim"
date: "2025-09-28"
output:
        rmdformats::downcute:
        downcute_theme: "chaos"
code_folding: show
fig_width: 6
fig_height: 6
df_print: paged
editor_options: 
        chunk_output_type: inline
markdown: 
        wrap: 72
---
        
```{r warning=FALSE, message=FALSE, echo=FALSE, results='hide', setup}
#===============================================================================
#BTC.LineZero.Header.1.1.0
#===============================================================================
#R Markdown environment setup and reporting utility.
#===============================================================================
#RLB.Dependencies:
#   knitr, magrittr, pacman, rio, rmarkdown, rmdformats, tibble, yaml
#===============================================================================
#Input for document parameters, libraries, file paths, and options.
#=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=


# --- working dir:
host <- Sys.info()[["nodename"]]

paths <- list(
  "bagel-S2600CO" = list(#Minsik's lab server
    work = "/mnt/4T_samsung/Dropbox/",
    lib  = "/home/bagel/R_lib/"
  ),
  "Minsiks-Mac-mini-2.local" = list(#Minsik's mac mini
    work = "/Volumes/macdrive/Dropbox/",
    lib  = "/Library/Frameworks/R.framework/Resources/library/"
  ),
  "Minsiks-MacBook-Pro.local|Minsiks-MBP.ht.home" = list(#Minsik's macbook pro
    work = "/Users/minsikkim/Dropbox (Personal)/",
    lib  = "/Library/Frameworks/R.framework/Resources/library/"
  ),
  "MGB030975" = list(#LaiLab mac mini
    work = "/Volumes/Kingston/Dropbox/",
    lib  = "/Library/Frameworks/R.framework/Resources/library/"
  )
)

pick_cfg <- function(host, map) {
  if (!is.null(map[[host]])) return(map[[host]])
  for (k in names(map)) if (grepl(k, host, ignore.case = TRUE)) return(map[[k]])
  NULL
}
cfg <- pick_cfg(host, paths)
if (is.null(cfg)) stop("No path for the host: ", host)

path_working <- cfg$work
path_library <- cfg$lib

# Adding libraries
knitr::opts_knit$set(root.dir = path_working)


str_libraries <- c("tidyverse", "pacman", "yaml", "ggplot2", "tidyverse", "swirl")


YAML_header <-
        '---
title: "BTE3207 week 5-1"
author: "Minsik Kim"
date: "2025.09.28"
output:
    rmdformats::downcute:
        downcute_theme: "chaos"
        code_folding: hide
        fig_width: 6
        fig_height: 6
---'
seed <- "20230904"

#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Loads libraries, file paths, and other document options.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Boot <- function() {
        .libPaths(path_library)
        
        require(pacman)
        pacman::p_load(c("knitr", "rmarkdown", "rmdformats", "yaml"))
        
        knitr::opts_knit$set(root.dir = path_working)
        
        str_libraries |> unique() |> sort() -> str_libraries
        pacman::p_load(char = str_libraries)
        
        set.seed(seed)
}
FUN.LineZero.Boot()
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Outputs R environment report.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Report <- function() {
        cat("Line Zero Environment:\n\n")
        paste("R:", pacman::p_version(), "\n") |> cat()
        cat("Libraries:\n")
        for (str_libraries in str_libraries) {
                paste(
                        "    ", str_libraries, ": ", pacman::p_version(package = str_libraries),
                        "\n", sep = ""
                ) |> cat()
        }
        paste("\nOperating System:", pacman::p_detectOS(), "\n") |> cat()
        paste("    Library Path:", path_library, "\n") |> cat()
        paste("    Working Path:", path_working, "\n") |> cat()
        paste("Seed:", seed, "\n\n") |> cat()
        cat("YAML Header:\n")
        cat(YAML_header)
}
FUN.LineZero.Report()

```

# Before begin..

Let's load the SBP dataset.

```{r}
dataset_sbp <- read.csv(file = "Inha/5_Lectures/1_BTE3207_Biostats/2025F/scripts/BTE3207_Advanced_Biostatistics/dataset/sbp_dataset_korea_2013-2014.csv") 

head(dataset_sbp)

```


# Sampling distribution and variability

Example 1

Let's see this example. If we sample a small `subset` of data, the distribution will look different from larger set of samples.

*Original data's histogram*

```{r}

hist(dataset_sbp$SBP, main = "Histogram of SBP")


```

# Sample variance

*Let's say we sampled 2 samples from the original dataset.*

```{r}

set.seed(1)

print("Two randomly selecte samples")

dataset_sbp$SBP %>%
             sample(2)

print("Two randomly selecte samples -2 ")

set.seed(2)


dataset_sbp$SBP %>%
             sample(2)


```

Two different random samplings showed difference in mean values.

We call this sample variance.


# Sampling distribution

Let's see how they looks like in histogram. 

This time, the sample number is slightly higher. 

*Histogram of 10 samples from the dataset (run 1)*

```{r}

dataset_sbp$SBP %>%
             sample(10) %>%
        hist(main = "Histogram of SBP, N = 10", 10)

```


*Histogram of 10 samples from the dataset (run 2)*

```{r}

dataset_sbp$SBP %>%
             sample(10) %>%
        hist(main = "Histogram of SBP, N = 10")

```

We **randomly** *sampled* 10 dataset from the `dataset_SBP` and looked at their distribution. Do you think they look the same?

No, unfortunately they are not.

Since the sampling takes samples randomly, their mean, distribution is different.


Of course, since this sampling was done by computer, we can make the sampling process ***reproducible*** by a `set.seed()` function in R.


```{r}

set.seed(1)

dataset_sbp$SBP %>%
             sample(10) %>%
        hist(main = "Histogram of SBP, N = 10")

set.seed(1)

dataset_sbp$SBP %>%
             sample(10) %>%
        hist(main = "Histogram of SBP, N = 10")


```
      
If we set the seed, the sample sampling procedure will use the `seed number` (in this case it was `1`) for it's algorithm. When the seed number is the same, the computer will generate the same data every time.


*However*, can we set seed for actual sampling processes in the real world? As it is impossible, the samples are always different from the population, and this phenomena is called `sampling variability`.

That means, in other words, the samples and their summary statistics actually cannot represent the population.

Then, how can we estimate the population?


Example data set for presentation material

```{r}

set.seed(1)

dataset_sbp$SBP %>%
             sample(20) %>%
        hist(main = "Histogram of SBP, N = 20")

dataset_sbp$SBP %>%
        sample(20) %>% 
        data.frame %>% 
        summarise(mean = mean(.),
                  sd = sd(.))


set.seed(1)

dataset_sbp$SBP %>%
             sample(50) %>%
        hist(main = "Histogram of SBP, N = 50")

dataset_sbp$SBP %>%
        sample(50) %>% 
        data.frame %>% 
        summarise(mean = mean(.),
                  sd = sd(.))



set.seed(1)

dataset_sbp$SBP %>%
             sample(100) %>%
        hist(main = "Histogram of SBP, N = 100")

dataset_sbp$SBP %>%
        sample(100) %>% 
        data.frame %>% 
        summarise(mean = mean(.),
                  sd = sd(.))


set.seed(1)

dataset_sbp$SBP %>%
             sample(1000) %>%
        hist(main = "Histogram of SBP, N = 1000")

dataset_sbp$SBP %>%
        sample(1000) %>% 
        data.frame %>% 
        summarise(mean = mean(.),
                  sd = sd(.))


set.seed(1)

dataset_sbp$SBP %>%
             sample(10000) %>%
        hist(main = "Histogram of SBP, N = 10000")

dataset_sbp$SBP %>%
        sample(10000) %>% 
        data.frame %>% 
        summarise(mean = mean(.),
                  sd = sd(.))



set.seed(1)

dataset_sbp$SBP %>%
             sample(100000) %>%
        hist(main = "Histogram of SBP, N = 100000")

dataset_sbp$SBP %>%
        sample(100000) %>% 
        data.frame %>% 
        summarise(mean = mean(.),
                  sd = sd(.))



```
## Repeated, n = 50


```{r}
set.seed(1)
df <- lapply(1:5, function(i) {
  data.frame(Sample = i,
             SBP = sample(dataset_sbp$SBP, 50))
}) %>% bind_rows()

ggplot(df, aes(x=SBP)) +
  geom_histogram(bins=15, fill="grey", color="black") +
  facet_wrap(~Sample) +
  theme_minimal() +
  theme(strip.text = element_text(size=8),
        axis.text.x = element_text(angle = 45)) +
  labs(title="Histograms of 5 random samples (N=50 each)")

```


```{r}


set.seed(1)

df <- lapply(1:72, function(i) {
  data.frame(Sample = i,
             SBP = sample(dataset_sbp$SBP, 50))
}) %>% bind_rows()


df_means <- df %>% 
  group_by(Sample) %>% 
  summarise(mean_SBP = mean(SBP))

ggplot(df, aes(x=SBP)) +
  geom_histogram(bins=15, fill="grey", color="black") +
  facet_wrap(~Sample) +
  theme_minimal() +
  theme(strip.text = element_text(size=8)) +
  labs(title="Histograms of 72 random samples (N=50 each)",
       subtitle="Red dashed line = sample mean")

ggplot(df, aes(x=SBP)) +
  geom_histogram(bins=15, fill="grey", color="black") +
  geom_vline(data=df_means, aes(xintercept=mean_SBP),
             color="red", linetype="dashed", size=0.7) +
  facet_wrap(~Sample) +
  theme_minimal() +
  theme(strip.text = element_text(size=8)) +
  labs(title="Histograms of 72 random samples (N=50 each)",
       subtitle="Red dashed line = sample mean")

```

# Sample mean

This time, let's do some other calculations.

We are going to sample 20 samples (randomly), 5000 times. And then, let's see its histogram

```{r}
set.seed(1)
replicate(50, sample(dataset_sbp$SBP, 20)) %>%
        data.frame() %>% 
        colMeans() %>%
        hist(100,
             main = "Distribution of 50 sample mean SBP from 1M subjects.\nEach sample N = 20",
             xlim = c(105, 135))


```

```{r}
set.seed(1)
replicate(500, sample(dataset_sbp$SBP, 20)) %>%
        data.frame() %>% 
        colMeans() %>%
        hist(100,
             main = "Distribution of 500 sample mean SBP from 1M subjects.\nEach sample N = 20",
             xlim = c(105, 135))


```


```{r}
set.seed(1)
replicate(5000, sample(dataset_sbp$SBP, 20)) %>%
        data.frame() %>% 
        colMeans() %>%
        hist(100,
             main = "Distribution of 5000 sample mean SBP from 1M subjects.\nEach sample N = 20",
             xlim = c(105, 135))


```

What if we increase the number of random sampling?

```{r}

set.seed(1)
replicate(5000, sample(dataset_sbp$SBP, 20)) %>%
        data.frame() %>% 
        colMeans() %>%
        hist(100,
             main = "Distribution of 5000 sample mean SBP from 1M subjects.\nEach sample N = 20",
             xlim = c(105, 135))


set.seed(1)
replicate(5000, sample(dataset_sbp$SBP, 50)) %>%
        data.frame() %>% 
        colMeans() %>%
        hist(100,
             main = "Distribution of 5000 sample mean SBP from 1M subjects.\nEach sample N = 50",
             xlim = c(105, 135))



set.seed(1)
replicate(5000, sample(dataset_sbp$SBP, 150)) %>%
        data.frame() %>% 
        colMeans() %>%
        hist(100,
             main = "Distribution of 5000 sample mean SBP from 1M subjects.\nEach sample N = 150",
             xlim = c(105, 135))


```


How do they look?

Let's see their variation in boxplots



```{r}

set.seed(1)
dataset_pbs_r5000_n50_250_400 <- data.frame(`n = 50` = (replicate(5000, sample(dataset_sbp$FBS, 50)) %>%
                               data.frame %>%
                               colMeans()),
           `n = 250` = (replicate(5000, sample(dataset_sbp$FBS, 250)) %>%
                                data.frame %>%
                                colMeans()),
           `n = 400` = (replicate(5000, sample(dataset_sbp$FBS, 400)) %>%
                                data.frame %>%
                                colMeans()),
           check.names = F
           ) %>%
        gather(condition, FBS, `n = 50`:`n = 400`, factor_key=TRUE) 

dataset_pbs_r5000_n50_250_400 %>% 
        boxplot(data = .,
                FBS ~ condition, 
                main = "Estimated sampling distribution, \nsample mean of 5000 random samples of \nn = 50, n = 250, and n = 400",
                xlab = "Sample N of each ramdom sampling")

```


Summary table

```{r}

dataset_pbs_r5000_n50_250_400 %>% 
        group_by(condition) %>%
        summarise(mean = mean(FBS),
                  sd = sd(FBS))

dataset_sbp %>% 
        summarise(mean = mean(FBS),
                  sd = sd(FBS))
```


Note that the change in variation is dependent on N, not the number of repeat

```{r}


set.seed(1)

dataset_sbp_r500_n20_50_150 <- data.frame(`n = 20` = (replicate(500, sample(dataset_sbp$SBP, 20)) %>%
                               data.frame %>%
                               colMeans()),
           `n = 100` = (replicate(500, sample(dataset_sbp$SBP, 100)) %>%
                                data.frame %>%
                                colMeans()),
           `n = 150` = (replicate(500, sample(dataset_sbp$SBP, 150)) %>%
                                data.frame %>%
                                colMeans()),
           check.names = F
           ) %>%
        gather(condition, SBP, `n = 20`:`n = 150`, factor_key=TRUE) 

set.seed(1)

dataset_sbp_r50_n20_50_150 <- data.frame(`n = 20` = (replicate(50, sample(dataset_sbp$SBP, 20)) %>%
                               data.frame %>%
                               colMeans()),
           `n = 100` = (replicate(50, sample(dataset_sbp$SBP, 100)) %>%
                                data.frame %>%
                                colMeans()),
           `n = 150` = (replicate(50, sample(dataset_sbp$SBP, 150)) %>%
                                data.frame %>%
                                colMeans()),
           check.names = F
           ) %>%
        gather(condition, SBP, `n = 20`:`n = 150`, factor_key=TRUE) 

set.seed(1)

dataset_sbp_r5000_n20_50_150 <- data.frame(`n = 20` = (replicate(5000, sample(dataset_sbp$SBP, 20)) %>%
                               data.frame %>%
                               colMeans()),
           `n = 100` = (replicate(50, sample(dataset_sbp$SBP, 100)) %>%
                                data.frame %>%
                                colMeans()),
           `n = 150` = (replicate(50, sample(dataset_sbp$SBP, 150)) %>%
                                data.frame %>%
                                colMeans()),
           check.names = F
           ) %>%
        gather(condition, SBP, `n = 20`:`n = 150`, factor_key=TRUE) 

dataset_sbp_r50_n20_50_150$`Number of repeat` <- 50
dataset_sbp_r500_n20_50_150$`Number of repeat` <- 500
dataset_sbp_r5000_n20_50_150$`Number of repeat` <- 5000


rbind(dataset_sbp_r50_n20_50_150,
      dataset_sbp_r500_n20_50_150,
      dataset_sbp_r5000_n20_50_150) %>%
        mutate(`Number of repeat` = as.factor(`Number of repeat`)) %>% 
        ggplot(data = ., aes(x = condition, y = SBP)) +
        geom_boxplot(aes(fill = `Number of repeat`)) +
        ggtitle("Estimated sampling distribution, \nsample mean of some random samples of \nn = 20, n = 50, and n = 150") +
        xlab("Sample N of each ramdom sampling")

?geom_boxplot()

```

# Sample mean and sample distribution of skewed data

As same as we observed the tendency of sample means with bell-shaped dataset, we are going to investigate skewed data

*Original data's histogram*

```{r}

hist(dataset_sbp$FBS, main = "Histogram of FBS", 10)

dataset_sbp %>% summarise(mean = mean(FBS),
                              sd = sd(FBS))

```

## Sample variance

*Let's say we sampled 2 samples from the original dataset.*

```{r}

set.seed(1)

print("Two randomly selected samples")

dataset_sbp$FBS %>%
             sample(2)

print("Two randomly selected samples -2 ")

set.seed(2)


dataset_sbp$FBS %>%
             sample(2)


```

Two different random samplings showed difference in mean values.

We call this sample variance.


## Sampling distribution

Let's see how they looks like in histogram. 

This time, the sample number is slightly higher. 

*Histogram of 10 samples from the dataset (run 1)*

```{r}
set.seed(3)
dataset_sbp$FBS %>%
             sample(10) %>%
        hist(main = "Histogram of FBS, N = 10", 10)

set.seed(3)
dataset_sbp$FBS %>%
        sample(10) %>%
        data.frame() %>%
        summarise(mean = mean(.),
                              sd = sd(.))




```


Histograms of multiple random sampling and their sample mean


```{r}

set.seed(1)
replicate(5000, sample(dataset_sbp$FBS, 50)) %>%
        data.frame() %>% 
        colMeans() %>%
        hist(100,
             main = "Distribution of 5000 sample mean FBS from 1M subjects.\nEach sample N = 50",
             xlim = c(85, 120))


set.seed(1)
replicate(5000, sample(dataset_sbp$FBS, 250)) %>%
        data.frame() %>% 
        colMeans() %>%
        hist(100,
             main = "Distribution of 5000 sample mean FBS from 1M subjects.\nEach sample N = 250",
             xlim = c(85, 120))



set.seed(1)
replicate(5000, sample(dataset_sbp$FBS, 400)) %>%
        data.frame() %>% 
        colMeans() %>%
        hist(100,
             main = "Distribution of 5000 sample mean FBS from 1M subjects.\nEach sample N = 400",
             xlim = c(85, 120))


```


How do they look?

Let's see their variation in boxplots



```{r}

set.seed(1)
dataset_sbp_r5000_n20_50_150 <- data.frame(`n = 20` = (replicate(5000, sample(dataset_sbp$SBP, 20)) %>%
                               data.frame %>%
                               colMeans()),
           `n = 100` = (replicate(5000, sample(dataset_sbp$SBP, 100)) %>%
                                data.frame %>%
                                colMeans()),
           `n = 150` = (replicate(5000, sample(dataset_sbp$SBP, 150)) %>%
                                data.frame %>%
                                colMeans()),
           check.names = F
           ) %>%
        gather(condition, SBP, `n = 20`:`n = 150`, factor_key=TRUE) 

dataset_sbp_r5000_n20_50_150 %>% 
        boxplot(data = .,
                SBP ~ condition, 
                main = "Estimated sampling distribution, \nsample mean of 5000 random samples of \nn = 20, n = 50, and n = 150",
                xlab = "Sample N of each ramdom sampling")

```


Summary table

```{r}

dataset_sbp_r5000_n20_50_150 %>% 
        group_by(condition) %>%
        summarise(mean = mean(SBP),
                  sd = sd(SBP))

dataset_sbp %>% 
        summarise(mean = mean(SBP),
                  sd = sd(SBP))
```


# Binary variables

Summary table

The criteria for diagnosing hyper tension (high blood pressure) is as follows. 

Normal: SBP less than 120 and DBP less than 80 mm Hg; Elevated: SBP 120 to 129 and DBP less than 80 mm Hg; Stage 1 hypertension: SBP 130 to 139 or DBP 80 to 89 mm Hg; Stage 2 hypertension: SBP greater than or equal to 140 mm Hg or DBP greater than or equal to 90 mm Hg.

Using that categorization, we are going to create a new variable called `hypertension`.

```{r}

dataset_sbp$hypertension <- ifelse(dataset_sbp$SBP > 130 | dataset_sbp$DBP > 80,
                                   1,
                                   0)

```

And the histogram of hypertension will look like this.

```{r}

dataset_sbp$hypertension %>% hist(main = "Histogram of hypertension")

dataset_sbp$hypertension %>% mean

```

## Sample mean of binary variable

If we take random 50 samples, due to sampling variation, the sample mean will show different results.

```{r}
set.seed(1)
dataset_sbp$hypertension %>% sample(50) %>% hist(main = "Histogram of hypertension of \n 50 random samples")

set.seed(1)
dataset_sbp$hypertension %>% sample(50) %>% mean
```

```{r}

dataset_sbp$hypertension %>% hist(main = "Histogram of hypertension")

dataset_sbp$hypertension %>% mean

```

```{r}

set.seed(1)

df <- lapply(1:72, function(i) {
  data.frame(Sample = i,
             SBP = sample(dataset_sbp$hypertension, 50))
}) %>% bind_rows()


df_means <- df %>% 
  group_by(Sample) %>% 
  summarise(mean_SBP = mean(SBP))

ggplot(df, aes(x=SBP)) +
  geom_histogram(bins=15, fill="grey", color="black") +
  facet_wrap(~Sample) +
  theme_minimal() +
  theme(strip.text = element_text(size=8)) +
  labs(title="Histograms of 72 random samples (N=50 each)",
       subtitle="Red dashed line = sample mean")

ggplot(df, aes(x=SBP)) +
  geom_histogram(bins=15, fill="grey", color="black") +
  geom_vline(data=df_means, aes(xintercept=mean_SBP),
             color="red", linetype="dashed", size=0.7) +
  facet_wrap(~Sample) +
  theme_minimal() +
  theme(strip.text = element_text(size=8)) +
  labs(title="Histograms of 72 random samples (N=50 each)",
       subtitle="Red dashed line = sample mean")

```


## Binomial distribution

```{r}
set.seed(123)

p_hat <- mean(dataset_sbp$hypertension)

n <- 3
sim_counts <- replicate(10000, {
  sample(dataset_sbp$hypertension, n, replace=TRUE) %>% sum()
})
hist(sim_counts, breaks=seq(-0.5, n+0.5, 1),
     freq=FALSE, 
     main=paste(n, "random sampling of hypertention status (n=",n,", p≈",round(p_hat,2),")"),
     xlab="Number of hypertensive subjects")
x <- 0:n
probs <- dbinom(x, size=n, prob=p_hat)
lines(x, probs, col="red", lwd=2) 

n <- 5
sim_counts <- replicate(10000, {
  sample(dataset_sbp$hypertension, n, replace=TRUE) %>% sum()
})

hist(sim_counts, breaks=seq(-0.5, n+0.5, 1),
     freq=FALSE, 
     main=paste(n, "random sampling of hypertention status (n=",n,", p≈",round(p_hat,2),")"),
     xlab="Number of hypertensive subjects")
x <- 0:n
probs <- dbinom(x, size=n, prob=p_hat)
lines(x, probs, col="red", lwd=2) 


n <- 10
sim_counts <- replicate(10000, {
  sample(dataset_sbp$hypertension, n, replace=TRUE) %>% sum()
})
hist(sim_counts, breaks=seq(-0.5, n+0.5, 1),
     freq=FALSE, 
     main=paste(n, "random sampling of hypertention status (n=",n,", p≈",round(p_hat,2),")"),
     xlab="Number of hypertensive subjects")
x <- 0:n
probs <- dbinom(x, size=n, prob=p_hat)
lines(x, probs, col="red", lwd=2) 

n <- 20
sim_counts <- replicate(10000, {
  sample(dataset_sbp$hypertension, n, replace=TRUE) %>% sum()
})

hist(sim_counts, breaks=seq(-0.5, n+0.5, 1),
     freq=FALSE, 
     main=paste(n, "random sampling of hypertention status (n=",n,", p≈",round(p_hat,2),")"),
     xlab="Number of hypertensive subjects")
x <- 0:n
probs <- dbinom(x, size=n, prob=p_hat)
lines(x, probs, col="red", lwd=2) 



n <- 1000
sim_counts <- replicate(10000, {
  sample(dataset_sbp$hypertension, n, replace=TRUE) %>% sum()
})

hist(sim_counts, breaks=seq(-0.5, n+0.5, 1),
     freq=FALSE, 
     main=paste(n, "random sampling of hypertention status (n=",n,", p≈",round(p_hat,2),")"),
     xlab="Number of hypertensive subjects")
x <- 0:n
probs <- dbinom(x, size=n, prob=p_hat)
lines(x, probs, col="red", lwd=2) 

```


## Sampling distribution of binary variable

```{r}
set.seed(1)
h_r_5000_n50 <- replicate(5000, sample(dataset_sbp$hypertension, 50)) 
h_r_5000_n50 %>% data.frame() %>% colMeans()
replicate(5000, sample(dataset_sbp$hypertension, 50)) %>%
        data.frame() %>% 
        colMeans() %>%
        hist(100,
             main = "Distribution of 5000 sample mean p-hat of hypertation from 1M subjects.\nEach sample N = 50",
             xlim = c(0, 0.6))


set.seed(1)
replicate(5000, sample(dataset_sbp$hypertension, 150)) %>%
        data.frame() %>% 
        colMeans() %>%
        hist(100,
             main = "Distribution of 5000 sample mean p-hat of hypertation from 1M subjects.\nEach sample N = 150",
             xlim = c(0, 0.6))



set.seed(1)
replicate(5000, sample(dataset_sbp$hypertension, 500)) %>%
        data.frame() %>% 
        colMeans() %>%
        hist(100,
             main = "Distribution of 5000 sample mean p-hat of hypertation from 1M subjects.\nEach sample N = 500",
xlim = c(0, 0.6))


```

## box plot 

```{r}
set.seed(1)
dataset_hypertension_r5000_n_50_150_500 <- data.frame(`n = 50` = (replicate(5000, sample(dataset_sbp$hypertension, 50)) %>%
                               data.frame %>%
                               colMeans()),
           `n = 150` = (replicate(5000, sample(dataset_sbp$hypertension, 150)) %>%
                                data.frame %>%
                                colMeans()),
           `n = 500` = (replicate(5000, sample(dataset_sbp$hypertension, 500)) %>%
                                data.frame %>%
                                colMeans()),
           check.names = F
           ) %>%
        gather(condition, hypertension, `n = 50`:`n = 500`, factor_key=TRUE) 

dataset_hypertension_r5000_n_50_150_500 %>% 
        boxplot(data = .,
                hypertension ~ condition, 
                main = "Estimated sampling distribution, \nsample mean of 5000 random samples of \nn = 50, n = 150, and n = 500",
                xlab = "Sample N of each ramdom sampling",
                ylab = "P-hat of mean of 5000 random sample sets")

```
## Summary table

```{r}
dataset_hypertension_r5000_n_50_150_500 %>%
        group_by(condition) %>%
        summarise(mean = mean(hypertension),
                  sd = sd(hypertension))

dataset_sbp %>% summarise(mean = mean(hypertension),
                  sd = sd(hypertension))
```


# Standard error (SE)

## Standard error of and it's theoretical value for it's subsets and sampling distribution

```{r}

dataset_sbp %>% summarise(mean = mean(SBP),
                              sd = sd(SBP))

sd(dataset_sbp$SBP)/sqrt(20)
sd(dataset_sbp$SBP)/sqrt(100)
sd(dataset_sbp$SBP)/sqrt(150)
```


## Standard error of sample proportions

```{r}

dataset_sbp %>% summarise(mean = mean(hypertension),
                              sd = sd(hypertension))

sd(dataset_sbp$hypertension)/sqrt(50)
sd(dataset_sbp$hypertension)/sqrt(150)
sd(dataset_sbp$hypertension)/sqrt(500)

sd(dataset_sbp$hypertension)

set.seed(8)
dataset_sbp$hypertension %>% sample(100) %>% sd
set.seed(1)
dataset_sbp$hypertension %>% sample(100) %>% sd 
set.seed(5)
dataset_sbp$hypertension %>% sample(100) %>% sd 




```


# Bibliography

```{r warning=FALSE, message=FALSE, echo=FALSE}
#===============================================================================
#BTC.LineZero.Footer.1.1.0
#===============================================================================
#R markdown citation generator.
#===============================================================================
#RLB.Dependencies:
#   magrittr, pacman, stringr
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#BTC.Dependencies:
#   LineZero.Header
#===============================================================================
#Generates citations for each explicitly loaded library.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
str_libraries <- c("r", str_libraries)
for (str_libraries in str_libraries) {
        str_libraries |>
                pacman::p_citation() |>
                print(bibtex = FALSE) |>
                capture.output() %>%
                .[-1:-3] %>% .[. != ""] |>
                stringr::str_squish() |>
                stringr::str_replace("_", "") |>
                cat()
        cat("\n")
}
#===============================================================================
```
