---
title: "COD_week7_1_MGK_BTE3207"
author: "Minsik Kim"
date: "2023-10-09"
output:
        rmdformats::downcute:
        downcute_theme: "chaos"
code_folding: show
fig_width: 6
fig_height: 6
df_print: paged
editor_options: 
        chunk_output_type: inline
markdown: 
        wrap: 72
---
        

        
```{r warning=FALSE, message=FALSE, echo=FALSE, results='hide', setup}
#===============================================================================
#BTC.LineZero.Header.1.1.0
#===============================================================================
#R Markdown environment setup and reporting utility.
#===============================================================================
#RLB.Dependencies:
#   knitr, magrittr, pacman, rio, rmarkdown, rmdformats, tibble, yaml
#===============================================================================
#Input for document parameters, libraries, file paths, and options.
#=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=
knitr::opts_chunk$set(message=FALSE, warning = FALSE)

path_working <- "/Users/minsikkim/Dropbox (Personal)/Inha/5_Lectures/Advanced biostatistics/scripts/BTE3207_Advanced_Biostatistics/"
path_library <- "/Library/Frameworks/R.framework/Resources/library"
str_libraries <- c("tidyverse", "pacman", "yaml")



YAML_header <-
        '---
title: "BTE3207 week 7-1"
author: "Minsik Kim"
date: "2032.10.10"
output:
    rmdformats::downcute:
        downcute_theme: "chaos"
        code_folding: hide
        fig_width: 6
        fig_height: 6
---'
seed <- "20230920"

#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Loads libraries, file paths, and other document options.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Boot <- function() {
        .libPaths(path_library)
        
        require(pacman)
        pacman::p_load(c("knitr", "rmarkdown", "rmdformats", "yaml"))
        
        knitr::opts_knit$set(root.dir = path_working)
        
        str_libraries |> unique() |> sort() -> str_libraries
        pacman::p_load(char = str_libraries)
        
        set.seed(seed)
}
FUN.LineZero.Boot()
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Outputs R environment report.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Report <- function() {
        cat("Line Zero Environment:\n\n")
        paste("R:", pacman::p_version(), "\n") |> cat()
        cat("Libraries:\n")
        for (str_libraries in str_libraries) {
                paste(
                        "    ", str_libraries, ": ", pacman::p_version(package = str_libraries),
                        "\n", sep = ""
                ) |> cat()
        }
        paste("\nOperating System:", pacman::p_detectOS(), "\n") |> cat()
        paste("    Library Path:", path_library, "\n") |> cat()
        paste("    Working Path:", path_working, "\n") |> cat()
        paste("Seed:", seed, "\n\n") |> cat()
        cat("YAML Header:\n")
        cat(YAML_header)
}
FUN.LineZero.Report()


```


# Before begin..

Let's load the SBP dataset.

```{r}

read.csv(file = "/Users/minsikkim/Dropbox (Personal)/Inha/5_Lectures/Advanced biostatistics/scripts/BTE3207_Advanced_Biostatistics/dataset/sbp_dataset_korea_2013-2014.csv")

dataset_sbp <- read.csv(file = "/Users/minsikkim/Dropbox (Personal)/Inha/5_Lectures/Advanced biostatistics/scripts/BTE3207_Advanced_Biostatistics/dataset/sbp_dataset_korea_2013-2014.csv")

head(dataset_sbp)

```

Making a new variable `hypertension`

```{r}

dataset_sbp$hypertension <- ifelse(dataset_sbp$SBP > 130 |
                                           dataset_sbp$DBP > 80,
                                   T,
                                   F)


```

# norm functions

`rnorm()` - random sampling

```{r}

rnorm(5)

#rnorm(100, mean = 100, sd = 1)

```

`pnorm()` - probability of given z-score (z-score table results)

```{r}

pnorm(2)

pnorm(0)

pnorm(1)

pnorm(3)

```


`qnorm()` - gets z-score of given probability

```{r}
qnorm(0.97724985)

qnorm(0.5)
```

`pnorm()` and `qnorm()` are reverse!

```{r}


pnorm(qnorm(0))

qnorm(pnorm(2))

```

`dnorm()` - *probability density function*, the height of normal curve.

```{r}

dnorm(0)

dnorm(1)

dnorm(-1)

```
Ploting of dnorm result

```{r}
seq(-4, 4, by = 0.5)

plot(dnorm(seq(-4, 4, by = 0.5))~seq(-4, 4, by = 0.5), 
     ylab = "dnorm() result",
     xlab = "dnorm() input",)

```


Ploting of normal curve

```{r}
#Create a sequence of 100 equally spaced numbers between -4 and 4
x <- seq(-4, 4, length=100)

#create a vector of values that shows the height of the probability distribution
#for each value in x
y <- dnorm(x)

#plot x and y as a scatterplot with connected lines (type = "l") and add
#an x-axis with custom labels
plot(x,y, type = "l", lwd = 2, axes = FALSE, xlab = "", ylab = "") +
        axis(1, at = -3:3, labels = c("-3s", "-2s", "-1s", "mean", "1s", "2s", "3s")) +
        abline(v = 2, col = "red") +
        abline(v = -2, col = "red")
        
```

# Statistical tests


## t-test

getting confidence interval of some data

```{r}
t.test(c(1, 2, 3))


result <- t.test(dataset_sbp$SBP) # Extract the confidence interval 

result$p.value


result$conf.int


```

comparing groups with t-test

```{r}

result <- t.test(SBP ~ SEX, data = dataset_sbp) # Extract the confidence interval 


result <- t.test(SBP ~ hypertension, data = dataset_sbp) # Extract the confidence interval 
result
```

## z.test()

one-sample z-test

```{r}

install.packages("BSDA")

library(BSDA)



result <- z.test(dataset_sbp$hypertension,
                 sigma.x = mean(dataset_sbp$hypertension) *
                         (1-mean(dataset_sbp$hypertension))
                 )


result

```

Hypertention table

```{r}
table(dataset_sbp$SEX)
table(dataset_sbp$DIS)

table(dataset_sbp$SEX, dataset_sbp$hypertension)


```

two-sample z-test

```{r}
#Maknig subsets of data

library(tidyverse)

dataset_male <- dataset_sbp %>% subset(., .$SEX == 1)
dataset_female <-  dataset_sbp %>% subset(., .$SEX == 2)

result <- z.test(x = dataset_male$hypertension,
                 y = dataset_female$hypertension,
                 sigma.x = mean(dataset_male$hypertension) * (1-mean(dataset_male$hypertension)),
                 sigma.y = mean(dataset_female$hypertension) * (1-mean(dataset_female$hypertension))
                 )


result

```


## Chi-square test

Chi-square test for two groups

```{r}

result <- chisq.test(x = dataset_sbp$hypertension,
                     y = dataset_sbp$SEX)

result

```

**p**-value is the same as z-score test

Chi-square test for multiple groups

```{r}

table(dataset_sbp$SEX, dataset_sbp$DIS)

result <- chisq.test(y = dataset_sbp$DIS,
           x = dataset_sbp$SEX)

result

```

## Exact test

```{r}

#Author DataFlair

data_frame <- read.csv("https://goo.gl/j6lRXD")

#Reading CSV
data <- table(data_frame$treatment, data_frame$improvement)


fisher.test(data)



```

```{r}

#Author DataFlair
table(dataset_sbp$SEX,
                  dataset_sbp$hypertension)

fisher.test(table(dataset_sbp$SEX,
                  dataset_sbp$hypertension))

table(dataset_sbp$SEX,
                  dataset_sbp$hypertension) %>%
        fisher.test()

```

As sample number is high, Fisher's exact test results are having the same result as Chi-square / z-test.


## Confidence interval calculations, statistical tests

`epitools` pacakge automates all the calculations and tests!


```{r}
install.packages("epitools")

library(epitools)


#Confidence interval of RR
riskratio.wald(data)

#Confidence interval of RR
oddsratio.wald(data)

```


# Bibliography

```{r warning=FALSE, message=FALSE, echo=FALSE}
#===============================================================================
#BTC.LineZero.Footer.1.1.0
#===============================================================================
#R markdown citation generator.
#===============================================================================
#RLB.Dependencies:
#   magrittr, pacman, stringr
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#BTC.Dependencies:
#   LineZero.Header
#===============================================================================
#Generates citations for each explicitly loaded library.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
str_libraries <- c("r", str_libraries)
for (str_libraries in str_libraries) {
        str_libraries |>
                pacman::p_citation() |>
                print(bibtex = FALSE) |>
                capture.output() %>%
                .[-1:-3] %>% .[. != ""] |>
                stringr::str_squish() |>
                stringr::str_replace("_", "") |>
                cat()
        cat("\n")
}
#===============================================================================
```
