---
title: "COD_week2_2_MGK_BTE3207"
author: "Minsik Kim"
date: "2023-09-05"
output:
        rmdformats::downcute:
        downcute_theme: "chaos"
code_folding: show
fig_width: 6
fig_height: 6
df_print: paged
editor_options: 
        chunk_output_type: inline
markdown: 
        wrap: 72
---
        
```{r warning=FALSE, message=FALSE, echo=FALSE, results='hide', setup}
#===============================================================================
#BTC.LineZero.Header.1.1.0
#===============================================================================
#R Markdown environment setup and reporting utility.
#===============================================================================
#RLB.Dependencies:
#   knitr, magrittr, pacman, rio, rmarkdown, rmdformats, tibble, yaml
#===============================================================================
#Input for document parameters, libraries, file paths, and options.
#=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=
knitr::opts_chunk$set(message=FALSE, warning = FALSE)

path_working <- "/Users/minsikkim/Dropbox (Personal)/Inha/5_Lectures/Advanced biostatistics/scripts/BTE3207_Advanced_Biostatistics/"
path_library <- "/Library/Frameworks/R.framework/Resources/library"
str_libraries <- c("tidyverse", "pacman", "yaml", "ggplot2", "tidyverse", "swirl")


YAML_header <-
        '---
title: "BTE3207 week 2-1"
author: "Minsik Kim"
date: "2032.09.04"
output:
    rmdformats::downcute:
        downcute_theme: "chaos"
        code_folding: hide
        fig_width: 6
        fig_height: 6
---'
seed <- "20230904"

#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Loads libraries, file paths, and other document options.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Boot <- function() {
        .libPaths(path_library)
        
        require(pacman)
        pacman::p_load(c("knitr", "rmarkdown", "rmdformats", "yaml"))
        
        knitr::opts_knit$set(root.dir = path_working)
        
        str_libraries |> unique() |> sort() -> str_libraries
        pacman::p_load(char = str_libraries)
        
        set.seed(seed)
}
FUN.LineZero.Boot()
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Outputs R environment report.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Report <- function() {
        cat("Line Zero Environment:\n\n")
        paste("R:", pacman::p_version(), "\n") |> cat()
        cat("Libraries:\n")
        for (str_libraries in str_libraries) {
                paste(
                        "    ", str_libraries, ": ", pacman::p_version(package = str_libraries),
                        "\n", sep = ""
                ) |> cat()
        }
        paste("\nOperating System:", pacman::p_detectOS(), "\n") |> cat()
        paste("    Library Path:", path_library, "\n") |> cat()
        paste("    Working Path:", path_working, "\n") |> cat()
        paste("Seed:", seed, "\n\n") |> cat()
        cat("YAML Header:\n")
        cat(YAML_header)
}
FUN.LineZero.Report()


```

# Before begin..

*Basics of R*

## Get current pathway of R working environment

```{r}
getwd()
```


## Listing files in current pathway

```{r}
list.files()
```

## Changing directory

Use `tab` for navigating folders!

```{r}
setwd("dataset/")

getwd()
```

Now, you are in dataset folder!

## Going back to original folder

To go back to the upper path,

```{r}
setwd("..")

getwd()
```

Now you came back to the original working directory.


# Logical values...having logical values.

```{r}
a = 1

a

```

`=` does the same thing as `<-`. 

To test if the thing are same, R uses `==`.


```{r}
a == 1

```

as we inserted `a <- 1` in the previous code chunk, this test results in `TRUE`

```{r}
"a" == 1

```
This test will test whether a character, `"a"`, is the same with a numeric value `1`. As they are not the same, it returns `FALSE`

Here, `TRUE` and `FALSE` results are , and they are one type of **binary variable**. It works for longer vectors or variables as well.

```{r}

c(1, 2, 3, 4, 5) == c(1, 2, 2, 4, 5)

c(1, 2, 3, 4, 5) == 1

```

And it results in a vector of all the logical tests. Using this, we can filter data easily!


# Howe to select values

To select values from vector, we use `[]` (square brackets).

```{r}
a <- c(1, 2, 3)

a[1]

```

`a[1]` will result in the first data element in this vector.

It is slightly different with some data with names.

```{r}

names(a) <- c("first", "second", "third")

str(a)

```
now the vector `a` is named numeric vector.

In this case,

```{r}
a[1] 

```

The results will be the name and the value!

This sometimes causes some problem when calculating data.

In that case, we need to use double brackets`[[]]`

```{r}
a[[1]]
```


By the way, selecting multiple numbers can be done with colons`:`.

```{r}
1:10

5:20
```

As it has output of vector (multiple elements), both below codes will work the same way

```{r}

a[c(1, 2, 3)]

a[1:3]

```

For selecting data in data frames, it works the same way but it separates the rows and columns using comma. Here is one example.

```{r}
dataframe_example <- data.frame(Joe = c(1:1000),
                                Trump = sample(1:1000, 100),
                                Obama = sample(1:1000, 100),
                                George = sample(1:1000, 100)
                                )

head(dataframe_example)
```

This is a data frame, meaning nothing (it just has 100 numbers randomly selected from a numbers between 1 to 1000), besides Biden. It has ordered numbers from 1 to 100.


To select some data, we use numbers or columns again. But we separate inputs with a comma `,`.

Selecting 1st row and 1st cloumn

```{r}

dataframe_example[1,1]



```

Selecting multiple rows in column 1

```{r}

dataframe_example[1:10, 1]



```


Selecting multiple rows and columns

```{r}
dataframe_example[3:5, 1:2]
```

# How to install more functions in R

We use `install.packages()` function to install The CRAN (Comprehensive R Archive Network) server.


```{r}
install.packages("tidyverse")
```

However, the package you just installed, is on your computer (somewhere in a folder called libraries), but they are not `loaded` to R program (simply saying, you did not open them). To use the installed packages, you need to use function `library()`

```{r}
library(tidyverse)
```

Now you can use `tidyverse` package!

`tidyverse` package is the one helps you writing code / summarizing results.

When you learn how to install new packages, you can navigate functions in that package using two colons (`::`).

```{r}
tidyverse::tidyverse_conflicts()
```

As we installed a new package, there could be a `conflits` in functions. `tidyverse_conflicts()` shows the list of those conflicts. 

As developers are doing things by their own, and CRAN does not have a system controlling the names of the newly developed functions. That means, a new function from a package that you installed, can be overlapped with other functions from other packages!


```{r}
dataset_sbp <- read.csv(file = "/Users/minsikkim/Dropbox (Personal)/Inha/5_Lectures/Advanced biostatistics/scripts/BTE3207_Advanced_Biostatistics/dataset/sbp_dataset_korea_2013-2014.csv") 

head(dplyr::filter(dataset_sbp, SEX == 1))

head(stats::filter(dataset_sbp$SEX, rep(1,3)))
```
`dplyr::filter()` will filter out the data based on given condition, `SEX == 1`.

(doing the same thing as `subset()`)

However, `stats::filter()` (which is the basic package comes with R) does different thing. It applies linear filtering to a univariate time series or to each series separately of a multivariate time series.

The best practice is to note all the function names with `::`. But generally, you don't have to do it as it is not that common problem.

# Basic tidyverse

Tidy verse helps you writing code efficiently. But how?


Let's see this example. We want to filter sample, based on some condition. And we have multiple conditions.

```{r}

head(filter(dataset_sbp, SEX == 1 & SBP > 120))

```

This function filtered based on multiple conditions, when `SEX == 1` and `SBP > 120`. But how are we going to do some imputation, and then filter out based on some conditions?


```{r}

head(filter(filter(filter(dataset_sbp, SEX == 1), SBP > 120), FBS > 110))

```

This function filtered based on multiple conditions, when `SEX == 1` and `SBP > 120`. Plus, it has `head` function outside again. 


It can be done with this code and it does the same thing.

```{r}

dataset_sbp %>%
        filter(SEX == 1) %>%
        filter(SBP > 120) %>%
        filter(FBS > 110) %>%
        head()

```


But how are we going to do some imputation, and then filter out based on some conditions?


Let's see this example again.

We can try adding multiple lines of code to do this. Let's say we are interested in the difference between SBP and DBP. And then we want to categorize them with genders. And then, we want to filter out the data based on their quantile.

```{r}

dataset_sbp$Diff_SBP_DBP <- dataset_sbp$SBP - dataset_sbp$DBP

dataset_sbp_male <- filter(dataset_sbp, SEX == 1)
dataset_sbp_female <- filter(dataset_sbp, SEX == 2)

avg_male <- mean(dataset_sbp_male$Diff_SBP_DBP)

avg_female <- mean(dataset_sbp_female$Diff_SBP_DBP)

sd_male <- sd(dataset_sbp_male$Diff_SBP_DBP)

sd_female <- sd(dataset_sbp_female$Diff_SBP_DBP)

data.frame(SEX = c(1, 2),
           aberage_by_group = c(avg_male, avg_female),
           sd_by_group = c(sd_male, sd_female))

```

We did it! However, the codes are quite nasty, and we have generated unnecessary intermediate data frames as well. Isn't there a smarter way?

# Piping

The good news is, `tidyverse::` package has a great feature called *piping*. In basic R, if we do not assign values with `<-`, the computer will just show the result and it won't store the output. 

Piping helps employing that output temprarilly, using `%>%`

```{r}
dataset_sbp %>% head()

```

Selection of piped data in tidyverse can be done with dot `.`.

```{r}
dataset_sbp %>% .$SEX %>% head()

```


The data will be moved the the next function, and will be employed for calculation.

```{r}
dataset_sbp %>% mutate(Diff_SBP_DBP = SBP - DBP) %>% head()
```

See? Here, `mutate()` is a function for calculating new variable in `tidyverse`.

Let's do the same thing with tidyverse.

```{r}

dataset_sbp %>% 
        mutate(Diff_SBP_DBP = SBP - DBP) %>% 
        group_by(SEX) %>%
        summarise(avereage_by_group = mean(Diff_SBP_DBP),
                  sd_by_group = sd(Diff_SBP_DBP))


```
Try to use piping as many as you can. 
And use lots of lines and indents.
It will make your code look tidier.

These application is also possible (I didn't make a bad example as it would take too much of my personal time)

```{r}


dataset_sbp %>% 
        group_by(SEX, DIS) %>%
        summarise(avereage = mean(Diff_SBP_DBP),
                  sd = sd(Diff_SBP_DBP))



```


# How to learn basic R (optional)

`swirl()`

swirl teaches you R programming and data science interactively, at your own pace, and right in the R console!


```{r}
install.packages("swirl")
library(swirl)
```

Don't go too further,,, it will do almost the half of my job, teaching (bio)stats.

# Bibliography

```{r warning=FALSE, message=FALSE, echo=FALSE}
#===============================================================================
#BTC.LineZero.Footer.1.1.0
#===============================================================================
#R markdown citation generator.
#===============================================================================
#RLB.Dependencies:
#   magrittr, pacman, stringr
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#BTC.Dependencies:
#   LineZero.Header
#===============================================================================
#Generates citations for each explicitly loaded library.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
str_libraries <- c("r", str_libraries)
for (str_libraries in str_libraries) {
        str_libraries |>
                pacman::p_citation() |>
                print(bibtex = FALSE) |>
                capture.output() %>%
                .[-1:-3] %>% .[. != ""] |>
                stringr::str_squish() |>
                stringr::str_replace("_", "") |>
                cat()
        cat("\n")
}
#===============================================================================
```
