---
title: "COD_week2_1_MGK_BTE3207"
author: "Minsik Kim"
date: "2023-09-04"
output:
rmdformats::downcute:
downcute_theme: "chaos"
code_folding: show
fig_width: 6
fig_height: 6
df_print: paged
editor_options: 
        chunk_output_type: inline
markdown: 
        wrap: 72
---
        
```{r warning=FALSE, message=FALSE, echo=FALSE, results='hide', setup}
#===============================================================================
#BTC.LineZero.Header.1.1.0
#===============================================================================
#R Markdown environment setup and reporting utility.
#===============================================================================
#RLB.Dependencies:
#   knitr, magrittr, pacman, rio, rmarkdown, rmdformats, tibble, yaml
#===============================================================================
#Input for document parameters, libraries, file paths, and options.
#=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=
knitr::opts_chunk$set(message=FALSE, warning = FALSE)

path_working <- "/Users/minsikkim/Dropbox (Personal)/Inha/5_Lectures/Advanced biostatistics/scripts/BTE3207_Advanced_Biostatistics/"
path_library <- "/Library/Frameworks/R.framework/Resources/library"
str_libraries <- c("tidyverse", "pacman", "yaml")


YAML_header <-
        '---
title: "BTE3207 week 2-1"
author: "Minsik Kim"
date: "2032.09.04"
output:
    rmdformats::downcute:
        downcute_theme: "chaos"
        code_folding: hide
        fig_width: 6
        fig_height: 6
---'
seed <- "20230904"

#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Loads libraries, file paths, and other document options.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Boot <- function() {
        .libPaths(path_library)
        
        require(pacman)
        pacman::p_load(c("knitr", "rmarkdown", "rmdformats", "yaml"))
        
        knitr::opts_knit$set(root.dir = path_working)
        
        str_libraries |> unique() |> sort() -> str_libraries
        pacman::p_load(char = str_libraries)
        
        set.seed(seed)
}
FUN.LineZero.Boot()
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Outputs R environment report.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Report <- function() {
        cat("Line Zero Environment:\n\n")
        paste("R:", pacman::p_version(), "\n") |> cat()
        cat("Libraries:\n")
        for (str_libraries in str_libraries) {
                paste(
                        "    ", str_libraries, ": ", pacman::p_version(package = str_libraries),
                        "\n", sep = ""
                ) |> cat()
        }
        paste("\nOperating System:", pacman::p_detectOS(), "\n") |> cat()
        paste("    Library Path:", path_library, "\n") |> cat()
        paste("    Working Path:", path_working, "\n") |> cat()
        paste("Seed:", seed, "\n\n") |> cat()
        cat("YAML Header:\n")
        cat(YAML_header)
}
FUN.LineZero.Report()


```

# Basic visualization of data

We are going to visualize this example dataset, SBP (systolic blood pressure) data from

https://nhiss.nhis.or.kr/bd/ab/bdabf003cv.do.

```{r}

dataset_sbp <- read.csv(file = "/Users/minsikkim/Dropbox (Personal)/Inha/5_Lectures/Advanced biostatistics/scripts/BTE3207_Advanced_Biostatistics/dataset/sbp_dataset_korea_2013-2014.csv") 

```

## Summary statistics of data

We can use `mean()`, `sd()`, `median()` functions to calculate summary statistics.

```{r}
cat("#Mean of SBP of 1M subject\n",
      mean(dataset_sbp$SBP),
      "\n\n#Standard deviation of SBP of 1M subject\n",
      sd(dataset_sbp$SBP),
      "\n\n#Median of SBP of 1M subject\n",
      median(dataset_sbp$SBP)
      )
```

`cat()` function prints out the character. `\n` changes the line of the console.

Alternatively, `print()` can be used but the output of `print()` function is recognized as data in R. `cat()` just adds lines in the console.


## summary()

Instead, R has a convenient function called `summary()`.

```{r}

summary(dataset_sbp$SBP)

```

Voila! Now we can roughly see how the data looks like. However, it will be more straight forward if we can see the data in a form of figure.

## hist()

`hist()` function creates histogram in R. It as multiple arguments to make more informative histogram as output. For example,

```{r}

hist(dataset_sbp$SBP)

```

## hist() - breaks

`hist()` as argument called `breaks = `, which can manually set the number of bars in the histogram.

```{r}

hist(dataset_sbp$SBP,
     breaks = 5)

```

```{r}

hist(dataset_sbp$SBP,
     breaks = 10)

```
We can change number of bars by setting `breaks = ` argument for `hist()` function.

## hist() - continued

By assigning `main = ` (title of histogram), `xlab = ` (x-axis label) and `ylab = ` (y-axis label), we can create a histogram with more detailed information.


```{r}

hist(dataset_sbp$SBP, 
     breaks = 10,
     main = "Systolic Blood Pressure (SBP)\nof 1M Koreans in 2013-2014",
     xlab = "SBP (mmHg) of 1M Koreans",
     ylab = "Number of measurements"
     )

```

## hist() - percentage

By imputing `density` variable of histogram, we can create a histogram with y axis that is having percentage.


```{r}

h <- hist(dataset_sbp$SBP,
         breaks = 10, 
         plot = F)

h$density <- h$counts/sum(h$counts)*100

plot(h,
     freq = FALSE,
     main = "Systolic Blood Pressure (SBP)\nof 1M Koreans in 2013-2014",
     xlab = "SBP (mmHg) of 1M Koreans",
     ylab = "Percentage of subject"
     )

```


# Boxplot

However, what should we do if we want to some summary statistic results as figures? Statisticians simply use boxplots. Boxplots can be generated by `boxplot()` command with information of what will be the x axis or colors.

## boxplot() - simple

We can directly put variable(vector) of our interest as `x`.

```{r}

boxplot(x = dataset_sbp$SBP,
        main = "Systolic Blood Pressure (SBP)\nof 1M Koreans in 2013-2014",
        ylab = "SBP (mmHg)"
        )

```

## boxplot() - with x axis

Boxplot is useful in comparing data, by adding more information along x-axis.
To make more redundant boxplot, use `formula = y ~ x`. Here, `y` will be the variable of y-axis and x will be the x-axis.

**Remember x should be categorical (discrete) data.**

```{r}

boxplot(formula = SBP ~ SEX,
        data = dataset_sbp,
        main = "Systolic Blood Pressure (SBP)\nof 1M Koreans in 2013-2014",
        ylab = "SBP (mmHg)",
        xlab = "Gender (male: 1, female: 2)"
        )

```

## boxplot() - x-axis label

we can also change x-axis texts with `names = ` argument.

```{r}

boxplot(SBP ~ SEX,
        data = dataset_sbp,
        main = "Systolic Blood Pressure (SBP)\nof 1M Koreans in 2013-2014",
        ylab = "SBP (mmHg)",
        xlab = "Gender",
        names = c("Male",
                  "Female")
        )

```

# Question

It seems like the histogram of SBP is somewhat having multiple peaks in the data. Can you tell why?

```{r}

hist(dataset_sbp$SBP,
     breaks = 500, 
     main = "Histogram of Systolic Blood Pressure (SBP)\nof 1M Koreans in 2013-2014 with 500 breaks",
     xlab = "SBP (mmHg) of 1M Koreans",
     ylab = "Number of measurements"
     )

```

# Common distributions

## Symmetric & bell-shaped

```{r}

hist(dataset_sbp$SBP,
     breaks = 10, 
     main = "Histogram of Systolic Blood Pressure (SBP)\nof 1M Koreans in 2013-2014 with 500 breaks",
     xlab = "SBP (mmHg) of 1M Koreans",
     ylab = "Number of measurements"
     )

```

## Right (positively) skewed

```{r}

hist(dataset_sbp$FBS,
     breaks = 10, 
     main = "Fasting Blood Sugar (FBS) levels\nof 1M Koreans in 2013-2014",
     xlab = "FBS (mg/L) of 1M Koreans",
     ylab = "Number of measurements"
     )

```
## Left (negatively skewed)


```{r}

hist(100-dataset_sbp$FBS,
     breaks = 10, 
     main = "100 - FBS",
     xlab = "100 - FBS",
     ylab = "Number of measurements"
     )

```

## Uniform distributions

```{r}
hist(dataset_sbp$BTH_G,
     breaks = 30, 
     main = "Age of cohort",
     xlab = "Age (years old)",
     ylab = "Number of measurements"
     )

```

# Bibliography

```{r warning=FALSE, message=FALSE, echo=FALSE}
#===============================================================================
#BTC.LineZero.Footer.1.1.0
#===============================================================================
#R markdown citation generator.
#===============================================================================
#RLB.Dependencies:
#   magrittr, pacman, stringr
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#BTC.Dependencies:
#   LineZero.Header
#===============================================================================
#Generates citations for each explicitly loaded library.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
str_libraries <- c("r", str_libraries)
for (str_libraries in str_libraries) {
        str_libraries |>
                pacman::p_citation() |>
                print(bibtex = FALSE) |>
                capture.output() %>%
                .[-1:-3] %>% .[. != ""] |>
                stringr::str_squish() |>
                stringr::str_replace("_", "") |>
                cat()
        cat("\n")
}
#===============================================================================
```
