---
title: "COD_week14_1_MGK_BTE3207"
author: "Minsik Kim"
date: "2023-11-28"
output:
        rmdformats::downcute:
        downcute_theme: "chaos"
code_folding: show
fig_width: 6
fig_height: 6
df_print: paged
editor_options: 
        chunk_output_type: inline
markdown: 
        wrap: 72
---
        
        
        
```{r warning=FALSE, message=FALSE, echo=FALSE, results='hide', setup}
#===============================================================================
#BTC.LineZero.Header.1.1.0
#===============================================================================
#R Markdown environment setup and reporting utility.
#===============================================================================
#RLB.Dependencies:
#   knitr, magrittr, pacman, rio, rmarkdown, rmdformats, tibble, yaml
#===============================================================================
#Input for document parameters, libraries, file paths, and options.
#=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=
knitr::opts_chunk$set(message=FALSE, warning = FALSE)

path_working <- 
        ifelse(sessionInfo()[1]$R.version$platform == "aarch64-apple-darwin20",
               "/Volumes/MacMini Drive/Dropbox (Personal)/Inha/5_Lectures/Advanced biostatistics/scripts/BTE3207_Advanced_Biostatistics/", 
               "/Users/minsikkim/Dropbox (Personal)/Inha/5_Lectures/Advanced biostatistics/scripts/BTE3207_Advanced_Biostatistics/")
path_library <- 
        "/Library/Frameworks/R.framework/Resources/library"
str_libraries <- c("tidyverse", "pacman", "yaml")



YAML_header <-
        '---
title: "BTE3207 week 13-2"
author: "Minsik Kim"
date: "2032.11.24"
output:
    rmdformats::downcute:
        downcute_theme: "chaos"
        code_folding: hide
        fig_width: 6
        fig_height: 6
---'
seed <- "20230920"

#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Loads libraries, file paths, and other document options.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Boot <- function() {
        .libPaths(path_library)
        
        require(pacman)
        pacman::p_load(c("knitr", "rmarkdown", "rmdformats", "yaml", "lmerTest", "vegan", "phyloseq", "ggplot"))
        
        knitr::opts_knit$set(root.dir = path_working)
        
        str_libraries |> unique() |> sort() -> str_libraries
        pacman::p_load(char = str_libraries)
        
        set.seed(seed)
}
FUN.LineZero.Boot()
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Outputs R environment report.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Report <- function() {
        cat("Line Zero Environment:\n\n")
        paste("R:", pacman::p_version(), "\n") |> cat()
        cat("Libraries:\n")
        for (str_libraries in str_libraries) {
                paste(
                        "    ", str_libraries, ": ", pacman::p_version(package = str_libraries),
                        "\n", sep = ""
                ) |> cat()
        }
        paste("\nOperating System:", pacman::p_detectOS(), "\n") |> cat()
        paste("    Library Path:", path_library, "\n") |> cat()
        paste("    Working Path:", path_working, "\n") |> cat()
        paste("Seed:", seed, "\n\n") |> cat()
        cat("YAML Header:\n")
        cat(YAML_header)
}
FUN.LineZero.Report()


```


# Agenda

This lecture will practice running 

1. running regerssion analysis to binary outcomes (logistic regression)

and

2. Multiple hypothesis testing

# Before begin..

Let's load the data

```{r}

download.file("dataset/sbp_dataset_korea_2013-2014.csv","https://github.com/minsiksudo/BTE3207_Advanced_Biostatistics/blob/main/dataset/sbp_dataset_korea_2013-2014.csv")

dataset_sbp <- read.csv("dataset/sbp_dataset_korea_2013-2014.csv")

```


```{r}

dataset_sbp$BTH_G %>% hist

```
Where, each level of BTH_G means
1: 20 to 24 years old
2: 25 to 26 years old
...
26: is 73 to 74
27: 75 years old or older.

# Data manipulation

from this dataset, we can calulate 

## Hypertension

```{r}
dataset_sbp$hypertension <- ifelse(dataset_sbp$SBP > 130 |
                                           dataset_sbp$DBP > 80,
                                   1,
                                   0) %>%
        as.factor

```



## Obesity

```{r}
dataset_sbp$obesity <- ifelse(dataset_sbp$BMI > 25,
                                   1,
                                   0) %>% 
        as.factor()
```





## Female (re-level)

```{r}
dataset_sbp$Female <- ifelse(dataset_sbp$SEX == 2, 
                          1,
                          0) %>%
        as.factor
```




## DIS (re-level)

```{r}
dataset_sbp$DIS <- factor(dataset_sbp$DIS, 
                          levels = c(1,2,3,4),
                          labels = c("Hypertension and diabetes",
                                     "Hypertension",
                                     "Diabetes",
                                     "No history"))
```


```{r}
set.seed(1)

dataset_sbp_small <- subset(dataset_sbp,
                        row.names(dataset_sbp)%in% sample(x = 1:1000000, size = 1000))

```

# Generating table

```{r}
table(dataset_sbp$obesity,dataset_sbp$Female)


```

```{r}

ggplot(dataset_sbp_small, aes(x = Female, y = obesity)) +
        geom_point() +
        ylab("Fasting blood sugar (mg/dL)") +
        theme_classic()

```

```{r}

ggplot(dataset_sbp_small, aes(x = FBS, y = obesity)) +
        geom_point() +
        ylab("Fasting blood sugar (mg/dL)") +
        theme_classic()

```

```{r}
male_oo <- table(dataset_sbp$obesity,dataset_sbp$SEX)[2,1]/table(dataset_sbp$obesity,dataset_sbp$SEX)[1,1]
female_oo <- table(dataset_sbp$obesity,dataset_sbp$SEX)[2,2]/table(dataset_sbp$obesity,dataset_sbp$SEX)[1,2]

odds_ratio_female_to_male <- female_oo/male_oo
odds_ratio_female_to_male


```

# Logistic regression with binary input (unadjusted)


```{r}
glm_obesity_gender <- glm(data = dataset_sbp, obesity ~ Female,family = "binomial")
summary(glm_obesity_gender)

```

```{r}
glm_obesity_gender$coefficients["Female1"][[1]]
```


```{r}
glm_obesity_gender$coefficients["Female1"][[1]] %>% exp()
```

The odds ratio of obesity from male to female is 0.5788211.

## confidence interval


```{r}
confint(glm_obesity_gender)
```

After exponemtial transformation...

```{r}
confint(glm_obesity_gender) %>% exp
```


# With continuous variable

```{r}
glm_obesity_fbs <- glm(data = dataset_sbp, obesity ~ FBS,family = "binomial")
summary(glm_obesity_fbs)

```

```{r}

confint(glm_obesity_fbs) %>% exp

```

# Multiple logistic regression

In case the Gender and FBS as correlation, they may adjust the effect of obese

## Data visualiation

```{r}

ggplot(dataset_sbp_small, aes(x = Female, y = FBS)) +
        geom_boxplot() +
        ylab("Fasting blood sugar (mg/dL)") +
        theme_classic()

```

Does the FBS differ by Gender?

```{r}

lm(dat = dataset_sbp, FBS ~ Female) %>%
        summary

```
It looks like so.

Let's run multiple logistic regression 


```{r}

glm(data = dataset_sbp, obesity ~ Female,family = "binomial") %>% confint() %>% exp




glm(data = dataset_sbp, obesity ~ FBS,family = "binomial") %>% confint() %>% exp


glm(data = dataset_sbp, obesity ~ DIS,family = "binomial") %>% confint() %>% exp





glm_obesity_gender_fbs <- glm(data = dataset_sbp, obesity ~ Female + FBS,family = "binomial")
glm_obesity_gender_fbs %>% confint() %>% exp

glm(data = dataset_sbp, obesity ~ Female + FBS + DIS,family = "binomial") %>% confint() %>% exp



```

## Confidence intervals 

```{r}

confint(glm_obesity_gender_fbs)

```
```{r}

confint(glm_obesity_gender_fbs) %>% exp

```


# Effect modification

To see if there is a effect modication, anova can be used.

```{r}


glm_obesity_gender_fbs_int <- glm(data = dataset_sbp, obesity ~ Female * FBS,family = "binomial")
anova(glm_obesity_gender_fbs_int, test = "F")

```


```{r}

summary(glm_obesity_gender_fbs_int)

```

# Binomial distribution

How can we caclulate the confidence interval?



# Data visualization of logistic regression


## For categorical varaibles

We can visualize the Odds ratio and its confidence interval using `forest plot`

```{r}
#install.packages("sjPlot")
library(sjPlot)

glm(data = dataset_sbp_small, obesity ~ DIS, family = "binomial") %>% 
        plot_model() +
        theme_classic()

```

```{r}
plot_models(
        glm(data = dataset_sbp_small, hypertension ~ DIS, family = "binomial"),
        glm(data = dataset_sbp_small, Female ~ DIS, family = "binomial"),
        glm(data = dataset_sbp_small, obesity ~ DIS, family = "binomial"),
        grid = TRUE
        ) +
        theme_classic()

```
Logistic plot for binary predictors



## Continuous variables




```{r}
dataset_sbp_small$FBS %>% summary

pred1 <- data.frame(FBS = seq(from = 60,
                                    to = 352, by = 2))

pred1$predict <- predict(glm_obesity_fbs, newdata = pred1, type = "response")

predict(glm_obesity_fbs, newdata = pred1, type = "response")
predict(glm_obesity_fbs, data.frame(FBS = 0), type = "response")


plot(dataset_sbp_small$FBS, as.numeric(dataset_sbp_small$obesity == 1),
     ylab = "Obesity",
     xlab = "FBS (mg/dL)") +
lines(pred1$FBS, pred1$predict)


```
```{R}

ggplot(dataset_sbp_small, aes(x = FBS, y = as.numeric(obesity) - 1)) +
    geom_point() +
    labs(y = "Probability of obesity", x = "FBS (mg/dL)") +
        theme_classic()

```




## Binomial distribution

Using binomial distribution, we can breifly know what the data is telling.

Let's assume that we are flipping a coin.

```{r}
#install.packages("tidydice")
library(tidydice)

flip_coin(times = 1)

```

We can flip the coin 1 times, and the result will be


```{r}

flip_coin(times = 2) %>% .$success %>% (mean)
flip_coin(times = 2) %>% .$success %>% (mean)

```



Sometimes we get 5 heads by 10, but many of trials will have results of 4 heads/3heads/6heads...etc. 



We can simulate 1000 round of 10-flips as below.

```{r}
flip_coin(times = 10, rounds = 1000) %>% head()


```


```{r}


flip_coin(times = 2, rounds = 10) %>%
        group_by(round) %>%
        summarise(probability = mean(success))

c(1.0, 0.5, 0.5, 0.0, 0.5, 0.0, 0.5, 0.0, 0.5, 1.0	) %>% hist(main = "Probability of getting heads when flipiing a coin\n(2 times X 10 rounds)", probability = T)


```



```{r}
flip_coin(times = 10, rounds = 100000) %>%
        group_by(round) %>%
        summarise(probability = mean(success)) %>%
        .$probability %>%
        hist(., main = "Probability of getting heads when flipiing a coin\n(10 times X 100000 rounds)",
             breaks = 10, xlim = c(0,1), probability = T)
```


```{r}
flip_coin(times = 50, rounds = 100000) %>%
        group_by(round) %>%
        summarise(probability = mean(success)) %>%
        .$probability %>%
        hist(., main = "Probability of getting heads when flipiing a coin\n(50 times X 100000 rounds)", 
             breaks = 25, probability = T, xlim = c(0,1))
```


```{r}
flip_coin(times = 100, rounds = 10000) %>%
        group_by(round) %>%
        summarise(probability = mean(success)) %>%
        .$probability %>%
        hist(., main = "Probability of getting heads when flipiing a coin\n(100 times X 10000 rounds)", 
             breaks = 50, probability = T, xlim = c(0,1))
```
```{r}
flip_coin(times = 1000, rounds = 10000) %>%
        group_by(round) %>%
        summarise(probability = mean(success)) %>%
        .$probability %>%
        hist(., main = "Probability of getting heads when flipiing a coin\n(1000 times X 10000 rounds)", 
             breaks = 20, xlim = c(0,1),
             probability = T)
```


As continuous variable is So, we usually "Predict" the proportion of the sample.


```{r}

glm_obesity_fbs_small <- glm(data = dataset_sbp, obesity ~ FBS, family = "binomial")


ilink <- family(glm_obesity_fbs_small)$linkinv

pred1 <- with(dataset_sbp_small,
           data.frame(FBS = seq(min(FBS), max(FBS),
                                       length = 100)))

pred1 <- cbind(pred1, predict(glm_obesity_fbs_small, pred1, type = "link", se.fit = TRUE)[1:2])

pred1 <- transform(pred1, Fitted = ilink(fit), Upper = ilink(fit + (2 * se.fit)),
                Lower = ilink(fit - (2 * se.fit)))


ggplot(dataset_sbp_small, aes(x = FBS, y = as.numeric(obesity) - 1)) +
    geom_point()
        theme_classic() +
    labs(y = "Probability of obesity", x = "FBS (mg/dL)")

ggplot(dataset_sbp_small, aes(x = FBS, y = as.numeric(obesity) - 1)) +
    geom_point() +
    geom_line(data = pred1, aes(y = Fitted, x = FBS)) + 
    geom_ribbon(data = pred1, aes(ymin = Lower, ymax = Upper, x = FBS),
    fill = "steelblue2", alpha = 0.2, inherit.aes = FALSE) +
        theme_classic() +
    labs(y = "Probability of obesity", x = "FBS (mg/dL)")

ggplot(dataset_sbp_small, aes(x = FBS, y = as.numeric(obesity) - 1)) +
    #geom_point() +
    geom_line(data = pred1, aes(y = Fitted, x = FBS)) + 
    geom_ribbon(data = pred1, aes(ymin = Lower, ymax = Upper, x = FBS),
    fill = "steelblue2", alpha = 0.2, inherit.aes = FALSE) +
        theme_classic() +
    labs(y = "Probability of obesity", x = "FBS (mg/dL)")

```




```{r}

glm_hypertension_sbp <- glm(data = dataset_sbp_small, hypertension ~ SBP, family = "binomial")


ilink <- family(glm_hypertension_sbp)$linkinv

pred1 <- with(dataset_sbp_small,
           data.frame(SBP = seq(min(SBP), max(SBP),
                                       length = 100)))

pred1 <- cbind(pred1, predict(glm_hypertension_sbp, pred1, type = "link", se.fit = TRUE)[1:2])

pred1 <- transform(pred1, Fitted = ilink(fit), Upper = ilink(fit + (2 * se.fit)),
                Lower = ilink(fit - (2 * se.fit)))


ggplot(dataset_sbp_small, aes(x = SBP, y = as.numeric(hypertension)-1)) +
    geom_point() +
        theme_classic() +
    labs(y = "Hypretension", x = "SBP (mmHg)")

ggplot(dataset_sbp_small, aes(x = SBP, y = as.numeric(hypertension)-1)) +
    geom_point() +
    geom_line(data = pred1, aes(y = Fitted, x = SBP)) + 
    geom_ribbon(data = pred1, aes(ymin = Lower, ymax = Upper, x = SBP),
    fill = "steelblue2", alpha = 0.2, inherit.aes = FALSE) +
        theme_classic() +
    labs(y = "Hypretension", x = "SBP (mmHg)")

ggplot(dataset_sbp_small, aes(x = SBP, y = as.numeric(hypertension)-1)) +
    #geom_point() +
    geom_line(data = pred1, aes(y = Fitted, x = SBP)) + 
    geom_ribbon(data = pred1, aes(ymin = Lower, ymax = Upper, x = SBP),
    fill = "steelblue2", alpha = 0.2, inherit.aes = FALSE) +
        theme_classic() +
    labs(y = "Hypretension", x = "SBP (mmHg)")

```



### Predicted porportion by groups

```{r}

dataset_sbp_small_dis1 <- dataset_sbp_small %>%
                                    subset(., as.numeric(.$DIS) == 1)
glm_hypertension_sbp <- glm(data = dataset_sbp_small_dis1, 
                            hypertension ~ SBP, family = "binomial")

ilink <- family(glm_hypertension_sbp)$linkinv
pred1 <- with(dataset_sbp_small_dis1,
           data.frame(SBP = seq(min(SBP), max(SBP),
                                       length = 100)))
pred1 <- cbind(pred1, predict(glm_hypertension_sbp, pred1, type = "link", se.fit = TRUE)[1:2])
pred1 <- transform(pred1, Fitted = ilink(fit), Upper = ilink(fit + (2 * se.fit)),
                Lower = ilink(fit - (2 * se.fit)))

dataset_sbp_small_dis2 <- dataset_sbp_small %>%
                                    subset(., as.numeric(.$DIS) == 2)
glm_hypertension_sbp <- glm(data = dataset_sbp_small_dis2, 
                            hypertension ~ SBP, family = "binomial")

ilink <- family(glm_hypertension_sbp)$linkinv
pred2 <- with(dataset_sbp_small_dis2,
           data.frame(SBP = seq(min(SBP), max(SBP),
                                       length = 100)))
pred2 <- cbind(pred2, predict(glm_hypertension_sbp, pred2, type = "link", se.fit = TRUE)[1:2])
pred2 <- transform(pred2, Fitted = ilink(fit), Upper = ilink(fit + (2 * se.fit)),
                Lower = ilink(fit - (2 * se.fit)))


dataset_sbp_small_dis3 <- dataset_sbp_small %>%
                                    subset(., as.numeric(.$DIS) == 3)
glm_hypertension_sbp <- glm(data = dataset_sbp_small_dis3, 
                            hypertension ~ SBP, family = "binomial")

ilink <- family(glm_hypertension_sbp)$linkinv
pred3 <- with(dataset_sbp_small_dis3,
           data.frame(SBP = seq(min(SBP), max(SBP),
                                       length = 100)))
pred3 <- cbind(pred3, predict(glm_hypertension_sbp, pred3, type = "link", se.fit = TRUE)[1:2])
pred3 <- transform(pred3, Fitted = ilink(fit), Upper = ilink(fit + (2 * se.fit)),
                Lower = ilink(fit - (2 * se.fit)))




dataset_sbp_small_dis4 <- dataset_sbp_small %>%
                                    subset(., as.numeric(.$DIS) == 4)
glm_hypertension_sbp <- glm(data = dataset_sbp_small_dis4, 
                            hypertension ~ SBP, family = "binomial")

ilink <- family(glm_hypertension_sbp)$linkinv
pred4 <- with(dataset_sbp_small_dis4,
           data.frame(SBP = seq(min(SBP), max(SBP),
                                       length = 100)))
pred4 <- cbind(pred4, predict(glm_hypertension_sbp, pred4, type = "link", se.fit = TRUE)[1:2])
pred4 <- transform(pred4, Fitted = ilink(fit), Upper = ilink(fit + (2 * se.fit)),
                Lower = ilink(fit - (2 * se.fit)))



ggplot(dataset_sbp_small, aes(x = SBP, y = as.numeric(hypertension)-1)) +
    #geom_point() +
    geom_line(data = pred1, aes(y = Fitted, x = SBP), col = "blue") + 
    geom_ribbon(data = pred1, aes(ymin = Lower, ymax = Upper, x = SBP),
    fill = "steelblue2", alpha = 0.2, inherit.aes = FALSE) +
        theme_classic() +
        geom_line(data = pred2, aes(y = Fitted, x = SBP), col = "red") + 
    geom_ribbon(data = pred1, aes(ymin = Lower, ymax = Upper, x = SBP),
    fill = "pink", alpha = 0.2, inherit.aes = FALSE) +
        theme_classic() +
    
        geom_line(data = pred3, aes(y = Fitted, x = SBP), col = "green") + 
    geom_ribbon(data = pred3, aes(ymin = Lower, ymax = Upper, x = SBP),
    fill = "lightgreen", alpha = 0.2, inherit.aes = FALSE) +
        theme_classic() +
    
        geom_line(data = pred4, aes(y = Fitted, x = SBP), col = "brown") + 
    geom_ribbon(data = pred4, aes(ymin = Lower, ymax = Upper, x = SBP),
    fill = "yellow", alpha = 0.2, inherit.aes = FALSE) +
        theme_classic()  + 
    labs(y = "Hypretension", x = "SBP (mmHg)")

```


# How to use this logistic regression in prediction?

## Types of errors


```{R}

ggplot(dataset_sbp_small, aes(x = SBP)) +
    geom_histogram() +
    labs(y = "Hypertension", x = "SBP (mmHg)") +
    theme_classic() +
        facet_wrap(~hypertension, ncol = 1)

```

```{R}

dataset_sbp_small$pred_hypertension <- 
        ifelse(glm(data = dataset_sbp_small, hypertension ~ SBP, family = "binomial") %>% predict(., type = "link") > 0, 1, 0)

```


```{R}

table(dataset_sbp_small$hypertension, dataset_sbp_small$pred_hypertension)

```


```{R}

ggplot(dataset_sbp_small, aes(x = SBP, y = hypertension, col = as.factor(pred_hypertension))) +
               geom_point() +
        theme_classic() +
        guides(col=guide_legend(title="Prediction"))

```

```{R}
?roc
roc(dataset_sbp_small$hypertension,  dataset_sbp_small$pred_hypertensio, plot = TRUE,
               print.auc = TRUE) +
        xlim(c(1,0))

```



## Separating dataset into train and test sets

```{r}
set.seed(1)
default_idx = sample(nrow(dataset_sbp), 5000)
default_trn = dataset_sbp[default_idx, ]
default_tst = dataset_sbp[-default_idx, ]

```


```{R}
#install.packages("pROC")
library(pROC)

model_train <- glm(data = default_trn, hypertension ~ SBP, family = "binomial")

test_prob = predict(model_train,
                    newdata = default_tst,
                    type = "response")

test_roc = roc(default_tst$hypertension ~ test_prob,
               plot = TRUE,
               print.auc = TRUE) +
        xlim(c(1,0))

```


# Bibliography

```{r warning=FALSE, message=FALSE, echo=FALSE}
#===============================================================================
#BTC.LineZero.Footer.1.1.0
#===============================================================================
#R markdown citation generator.
#===============================================================================
#RLB.Dependencies:
#   magrittr, pacman, stringr
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#BTC.Dependencies:
#   LineZero.Header
#===============================================================================
#Generates citations for each explicitly loaded library.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
str_libraries <- c("r", str_libraries)
for (str_libraries in str_libraries) {
        str_libraries |>
                pacman::p_citation() |>
                print(bibtex = FALSE) |>
                capture.output() %>%
                .[-1:-3] %>% .[. != ""] |>
                stringr::str_squish() |>
                stringr::str_replace("_", "") |>
                cat()
        cat("\n")
}
#===============================================================================
```
