---
title: "COD_week16_1_MGK_BTE3207"
author: "Minsik Kim"
date: "2023-12-14"
output:
        rmdformats::downcute:
        downcute_theme: "chaos"
code_folding: show
fig_width: 6
fig_height: 6
df_print: paged
editor_options: 
        chunk_output_type: inline
markdown: 
        wrap: 72
---
        
        
        
```{r warning=FALSE, message=FALSE, echo=FALSE, results='hide', setup}
#===============================================================================
#BTC.LineZero.Header.1.1.0
#===============================================================================
#R Markdown environment setup and reporting utility.
#===============================================================================
#RLB.Dependencies:
#   knitr, magrittr, pacman, rio, rmarkdown, rmdformats, tibble, yaml
#===============================================================================
#Input for document parameters, libraries, file paths, and options.
#=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=
knitr::opts_chunk$set(message=FALSE, warning = FALSE)

path_working <- 
        ifelse(sessionInfo()[1]$R.version$platform == "aarch64-apple-darwin20",
               "/Volumes/MacMini Drive/Dropbox (Personal)/Inha/5_Lectures/Advanced biostatistics/scripts/BTE3207_Advanced_Biostatistics/", 
               "/Users/minsikkim/Dropbox (Personal)/Inha/5_Lectures/Advanced biostatistics/scripts/BTE3207_Advanced_Biostatistics/")
path_library <- 
        "/Library/Frameworks/R.framework/Resources/library"
str_libraries <- c("tidyverse", "pacman", "yaml")



YAML_header <-
        '---
title: "BTE3207 week 16-1"
author: "Minsik Kim"
date: "2032.12.14"
output:
    rmdformats::downcute:
        downcute_theme: "chaos"
        code_folding: hide
        fig_width: 6
        fig_height: 6
---'
seed <- "20230920"

#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Loads libraries, file paths, and other document options.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Boot <- function() {
        .libPaths(path_library)
        
        require(pacman)
        pacman::p_load(c("knitr", "rmarkdown", "rmdformats", "yaml", "lmerTest", "vegan", "phyloseq", "ggplot"))
        
        knitr::opts_knit$set(root.dir = path_working)
        
        str_libraries |> unique() |> sort() -> str_libraries
        pacman::p_load(char = str_libraries)
        
        set.seed(seed)
}
FUN.LineZero.Boot()
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Outputs R environment report.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Report <- function() {
        cat("Line Zero Environment:\n\n")
        paste("R:", pacman::p_version(), "\n") |> cat()
        cat("Libraries:\n")
        for (str_libraries in str_libraries) {
                paste(
                        "    ", str_libraries, ": ", pacman::p_version(package = str_libraries),
                        "\n", sep = ""
                ) |> cat()
        }
        paste("\nOperating System:", pacman::p_detectOS(), "\n") |> cat()
        paste("    Library Path:", path_library, "\n") |> cat()
        paste("    Working Path:", path_working, "\n") |> cat()
        paste("Seed:", seed, "\n\n") |> cat()
        cat("YAML Header:\n")
        cat(YAML_header)
}
FUN.LineZero.Report()


```


# Final statistics

```{r}
final_score <- readxl::read_excel("/Volumes/MacMini Drive/Dropbox (Personal)/Inha/5_Lectures/Advanced biostatistics/2023F/Scoreboard/BTE3207_2023F_final_results.xlsx") 

```

```{r}
final_score_by_student <-
        final_score %>% 
        gather(., Problemset, Score, `1`:`total`) %>%
        .[order(.$student_id), ] %>%
        mutate(Fullscore  = rep(c(5, 5, 8, 10, 5, 10, 10, 5, 58), 26)) %>%
        select(-c("rank"))
final_score_by_student %>% filter(name == "강준석")
final_score_by_student %>% filter(name == "권은선")
final_score_by_student %>% filter(name == "김나현")
final_score_by_student %>% filter(name == "김보송")
final_score_by_student %>% filter(name == "김수인")
final_score_by_student %>% filter(name == "김윤아")
final_score_by_student %>% filter(name == "김준우")
final_score_by_student %>% filter(name == "김지인")
final_score_by_student %>% filter(name == "김지현")
final_score_by_student %>% filter(name == "류진")
final_score_by_student %>% filter(name == "박준영")
final_score_by_student %>% filter(name == "박준하")
final_score_by_student %>% filter(name == "박지용")
final_score_by_student %>% filter(name == "서창호")
final_score_by_student %>% filter(name == "김준우")
final_score_by_student %>% filter(name == "성호정")
final_score_by_student %>% filter(name == "손지현")
final_score_by_student %>% filter(name == "오준혁")
final_score_by_student %>% filter(name == "옥재현")
final_score_by_student %>% filter(name == "유상민")
final_score_by_student %>% filter(name == "이규섭")
final_score_by_student %>% filter(name == "이민정")
final_score_by_student %>% filter(name == "이성준")
final_score_by_student %>% filter(name == "이예원")
final_score_by_student %>% filter(name == "이한영")
final_score_by_student %>% filter(name == "이효정")

final_score_by_student %>% filter(name == "황용일")


final_score_by_student %>% filter(name == "김준우")
final_score_by_student %>% filter(name == "김준우")
final_score_by_student %>% filter(name == "김준우")
final_score_by_student %>% filter(name == "김준우")
final_score_by_student %>% filter(name == "김준우")
final_score_by_student %>% filter(name == "김준우")
final_score_by_student %>% filter(name == "김준우")



```


```{r}
final_score %>% 
        ggplot(aes(x = total)) +
        geom_histogram(bins = 20) +
        theme_classic() 

final_score$total %>% summary        

```


```{r}
gather(final_score, problem, score, `1`:total, factor_key = T) %>%
        ggplot(aes(x = score, fill = problem)) +
        geom_histogram(bins = 8) +
        facet_wrap(~problem, scales = "free_x") +
        theme_classic() 


```

# Q-Q plot

Were they normally distributed?

```{r}

gather(final_score, problem, score, `1`:total, factor_key = T) %>%
        ggplot(aes(sample = score, col = problem)) +
        stat_qq() + 
        stat_qq_line() +
        facet_wrap(~problem, scales = "free_y") +
        theme_classic()


```

# Does the student with high midterm score has high final score?

```{r}


BTE3207_summary <- data.frame(student_id = final_score$student_id,
           final = final_score$total,
           midterm = readxl::read_excel("/Volumes/MacMini Drive/Dropbox (Personal)/Inha/5_Lectures/Advanced biostatistics/2023F/Scoreboard/BTE3207_2023F_midterm_results.xlsx") %>%
        .$total)


```

```{r}

BTE3207_summary %>%
        ggplot(aes(x = midterm, y = final, col = student_id)) +
        geom_point() +
        scale_color_viridis_c() +
        theme_classic() +
        theme(legend.position = "none") +
        xlab("Midterm score") +
        ylab("Final score") +
        geom_smooth(method = "lm", color = "gray", fill = "lightgray", linetype = "dashed")
                
        


lm(data = BTE3207_summary, final ~ midterm) %>% summary




```

# Agenda

This lecture will practice running 

Multiple hypothesis testing correction

# What is multiple hypothesis testing?

```{r}


random_data <- data.frame(random_1 = sample(1:100, 10),
                          random_2 = sample(1:100, 10))

random_data$random_1 %>% hist
random_data$random_2 %>% hist
```


```{r}
lm(data = random_data, random_1 ~ random_2)


```
Of course, two random data sets are not having significant correlation.

However, what if we have many many many random dataset?

```{r}
set.seed(10)
random_1000 <- replicate(1000, sample(1:100, 10)) %>%
        data.frame() %>%
        rownames_to_column("number")



head(random_1000)

```


```{r}

data_long <- gather(random_1000, sample, data, X1:X1000, factor_key=TRUE)

data_long

```



```{r}

test_lm <- lm(data = data_long, data ~ as.factor(sample)) 

test_lm %>% summary

```

```{r}
summary(test_lm) %>%
        .$coef %>%
        .[,4] %>%
        hist
        

        
        

```
Some random data got significant..?

THis trend gets worse with normally distributed datasets.

```{r}
set.seed(10)
random_1000 <- replicate(1000, rnorm(10, mean = 0, sd = 1)) %>%
        data.frame() %>%
        rownames_to_column("number")



head(random_1000)

```

```{r}

data_long <- gather(random_1000, sample, data, X1:X1000, factor_key=TRUE)

data_long

```

```{r}

test_lm <- lm(data = data_long, data ~ as.factor(sample)) 

test_lm %>% summary

```

```{r}
summary(test_lm) %>%
        .$coef %>%
        .[,4] %>%
        hist
        
```


This is the caveat of having big data. with random dataset, and having multiple `hypothesis testing`, you will find `something` that is statistically significant. 

# Multipel hypothesis testing correciton methods



## Benjamini-Hochberg

FDR (False discovery rate)-based adjusting.

Assuming that the FDR is 0.1 (10% of significnat association is false-positives).

```{r}

summary(test_lm) %>%
        .$coef %>%
        .[,4] %>%
        p.adjust(., method="BH") %>%
        hist

```



## Bonferroni

Just multiplying `p-values` with `number of tests`!

In this case, only p-value with lower than 0.00005 is statistically significant

```{r}

summary(test_lm) %>%
        .$coef %>%
        .[,4] %>%
        p.adjust(., method="bonferroni") %>%
        hist

```

# qvalue package

`qvalue` package uses similar approach to BH correction but with more difficult assumptions.......This package takes a list of p-values resulting from the simultaneous testing of many hypotheses and estimates their q-values and local FDR values. The q-value of a test measures the proportion of false positives incurred (called the false discovery rate) when that particular test is called significant. The local FDR measures the posterior probability the null hypothesis is true given the test's p-value. Various plots are automatically generated, allowing one to make sensible significance cut-offs.


```{r}

#if (!require("BiocManager", quietly = TRUE))
#    install.packages("BiocManager")

#BiocManager::install("qvalue")

summary(test_lm) %>%
        .$coef %>%
        .[,4] %>%
        qvalue::qvalue(.) %>% 
        hist



```
```{r}
summary(test_lm) %>%
        .$coef %>%
        .[,4] %>%
        qvalue::qvalue(.) %>% 
        .$qvalues %>% 
        hist

```

# Bibliography

```{r warning=FALSE, message=FALSE, echo=FALSE}
#===============================================================================
#BTC.LineZero.Footer.1.1.0
#===============================================================================
#R markdown citation generator.
#===============================================================================
#RLB.Dependencies:
#   magrittr, pacman, stringr
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#BTC.Dependencies:
#   LineZero.Header
#===============================================================================
#Generates citations for each explicitly loaded library.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
str_libraries <- c("r", str_libraries)
for (str_libraries in str_libraries) {
        str_libraries |>
                pacman::p_citation() |>
                print(bibtex = FALSE) |>
                capture.output() %>%
                .[-1:-3] %>% .[. != ""] |>
                stringr::str_squish() |>
                stringr::str_replace("_", "") |>
                cat()
        cat("\n")
}
#===============================================================================
```