---
title: "COD_week1_2_MGK_BTE3207"
author: "Minsik Kim"
date: "2023-09-04"
output:
        rmdformats::downcute:
        downcute_theme: "chaos"
code_folding: show
fig_width: 6
fig_height: 6
df_print: paged
editor_options: 
        chunk_output_type: inline
markdown: 
        wrap: 72
---
        
```{r warning=FALSE, message=FALSE, echo=FALSE, results='hide', setup}
#===============================================================================
#BTC.LineZero.Header.1.1.0
#===============================================================================
#R Markdown environment setup and reporting utility.
#===============================================================================
#RLB.Dependencies:
#   knitr, magrittr, pacman, rio, rmarkdown, rmdformats, tibble, yaml
#===============================================================================
#Input for document parameters, libraries, file paths, and options.
#=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=
knitr::opts_chunk$set(message=FALSE, warning = FALSE)

path_working <- "/Users/minsikkim/Dropbox (Personal)/Inha/5_Lectures/Advanced biostatistics/scripts/BTE3207_Advanced_Biostatistics/"
path_library <- "/Library/Frameworks/R.framework/Resources/library"
str_libraries <- c("tidyverse", "pacman", "yaml")


YAML_header <-
        '---
title: "BTE3207 week 1-2"
author: "Minsik Kim"
date: "2032.09.04"
output:
    rmdformats::downcute:
        downcute_theme: "chaos"
        code_folding: hide
        fig_width: 6
        fig_height: 6
---'
seed <- "20230904"

#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Loads libraries, file paths, and other document options.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Boot <- function() {
        .libPaths(path_library)
        
        require(pacman)
        pacman::p_load(c("knitr", "rmarkdown", "rmdformats", "yaml"))
        
        knitr::opts_knit$set(root.dir = path_working)
        
        str_libraries |> unique() |> sort() -> str_libraries
        pacman::p_load(char = str_libraries)
        
        set.seed(seed)
}
FUN.LineZero.Boot()
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Outputs R environment report.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Report <- function() {
        cat("Line Zero Environment:\n\n")
        paste("R:", pacman::p_version(), "\n") |> cat()
        cat("Libraries:\n")
        for (str_libraries in str_libraries) {
                paste(
                        "    ", str_libraries, ": ", pacman::p_version(package = str_libraries),
                        "\n", sep = ""
                ) |> cat()
        }
        paste("\nOperating System:", pacman::p_detectOS(), "\n") |> cat()
        paste("    Library Path:", path_library, "\n") |> cat()
        paste("    Working Path:", path_working, "\n") |> cat()
        paste("Seed:", seed, "\n\n") |> cat()
        cat("YAML Header:\n")
        cat(YAML_header)
}
FUN.LineZero.Report()


```

# Working with data in R

There are different type of data in R. The smallest data is called `data element`, and when data element is gathered (with orders), that chunk of data is called `vector`.

Here are some examples.

## Data element

```{r}
alpha <- 1 # this is data element

alpha # if we call alpha, computer will tell you what is assaigned as alpha.

```

## Vecetor

To assign vector in R, we use `c()`, with comma `,` as separators. For example, 

```{r}

beta <- c(1, 2) # this is vector

beta

```


```{r}
gamma <- c(alpha, beta) # it is still a vector

gamma

```

Remember vectors are **ordered**.

```{r}

delta <- c(beta, alpha) # it is still a vector

delta

```

*If you call* `beta` *in R, it will find data named as *`beta`. *Then, how do we assign characters (letters, categorical variable)?* **R recognizes character values with **`""`. For instance,

```{r}

delta_2 <- c("beta", "alpha") 

delta_2

```

With quotation marks, the R will recognize inputs as a new character data, not the data we assigned previously.


# Data frame

Then, what should we do with larger data, such as excel spread sheets?

Vectors in R can be chucked together, **when their legnth is the same**.

This is called data frame in R, and it can be assigned using `data.frame()` function.


```{r}

epsilon <- data.frame(gamma, delta) # this is data frame
# note that the length of both gamma and delta is the same.

epsilon

```


## Selecting variables (vectors) in data frame

Data frame has columns, where each column is one variable (vectors)

we use dollar sign to select one variable from data frame

```{r}

epsilon$gamma

epsilon$delta


```

`epsilon$gamma` returns exact the same vector as just `gamma`, which was used for constructing epsilon.

Then, how can we generate larger dataset to analyze something?

## Data frame example

We can make data by adding multiple data elements to vectors, and we can bind those vectors to make one dataframe.


```{r}
subject <- c("Joe", "Trump", "Obama", "George") #assigning multiple data as a vector
height <- c(183, 190, 187, 182)
IsTall <- c("short", "tall", "tall", "short")

example_dataframe <- data.frame(subject, height, IsTall) #binding 3 vectors as one data frame

example_dataframe #loading the example_data frame

```

## Loading data

However, we are not going to type all the data whenever they are needed. To save our previous time, we will directly load the data frame from our computer (most commonly `.csv`, we can also import `.xlsx`) using `read.csv()` function.

Here is the example of a SBP (systolic blood pressure) data from https://nhiss.nhis.or.kr/bd/ab/bdabf003cv.do.

```{r}

dataset_sbp <- read.csv(file = "/Users/minsikkim/Dropbox (Personal)/Inha/5_Lectures/Advanced biostatistics/datasets/sbp_dataset/sbp_dataset_korea_2013-2014.csv") 

```

The data is loaded as below. However, as it contains 1M lines of data, calling the data will make this document too long. Here, we can use `head()` or `tail()` functions to see only 10-first or 10-last dataset, respectively.

## head() and tail() functions {.tabset}

### head() 

```{r}

head(dataset_sbp)

```

### tail()

```{r}

tail(dataset_sbp)

```

### full dataset

```{r}

dataset_sbp

```

### {-}

This shows (subsets of) gender, age group, SBP, DBP, FBS, DIS, and BMI data of 1,000,000 subjects. 

For now, we will use dollar sign to select one variable from dataframe. Again, I will use head(function to check starting 10 data at this time)


```{r}

head(dataset_sbp$SBP)

```

Such a long data!

# Basic R functions for statistics

Now, let's use computer to calculate summary statistics of this large dataset.


## Mean calculation

Function for calculating mean is `mean()`

```{r}
mean(dataset_sbp$SBP)
```

This is the mean value of all the SBP data (1,000,000).

## SD calculation

Function for calculating standard deviation is `sd()`

```{r}

sd(dataset_sbp$SBP)

```

This is the SD.

## Median

function for calculating median: median()

Median can be calculated by using `median()`


```{r}

median(dataset_sbp$SBP)

```


## Function for calculating quantile: qunatile()

`quantile()` function has two functions. Without assigning any percentile value of our interest, it will list all the quantiles (0, 25%, 50%, 75% and 100% percentiles) at once.

```{r}

quantile(dataset_sbp$SBP)

```

As the output of `quantile()` shows multiple data element, the output is a **vector**. Kindly, this output also has names as well. To check what the data has output, we can test `str()`. 

```{r}

str(quantile(dataset_sbp$SBP))

```

The str() says it is `named numeric variable`.

## quantile() - continued

However, what should we do if we are curious about values at a specific percentile? In that case we need to give more information to functions.

These information handed over to function are called **arguments**.

As functions in R can have multiple inputs, we separate them with commas (`,`). 

for example, 

10% percentile of the SBP dataset can be calculated using

```{r}

quantile(x = dataset_sbp$SBP, 0.1)

```

# Use question mark to see what function is doing

As thous functions were constructed by multiple , and there are lots of `packages` that can be newly installed to employ new functions in R. Developers listed all the details and instructions when they made new functions, and they can be reviewed by adding `questionmark` to a function.

```{r}
?quantile(data)
```


For some popular functions, details for `arguments` are usullally embeded on pop-ups. Use `tab` button to see what function can do.


# Bibliography

```{r warning=FALSE, message=FALSE, echo=FALSE}
#===============================================================================
#BTC.LineZero.Footer.1.1.0
#===============================================================================
#R markdown citation generator.
#===============================================================================
#RLB.Dependencies:
#   magrittr, pacman, stringr
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#BTC.Dependencies:
#   LineZero.Header
#===============================================================================
#Generates citations for each explicitly loaded library.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
str_libraries <- c("r", str_libraries)
for (str_libraries in str_libraries) {
        str_libraries |>
                pacman::p_citation() |>
                print(bibtex = FALSE) |>
                capture.output() %>%
                .[-1:-3] %>% .[. != ""] |>
                stringr::str_squish() |>
                stringr::str_replace("_", "") |>
                cat()
        cat("\n")
}
#===============================================================================
```
