---
title: "COD_week11_1_MGK_BTE3207"
author: "Minsik Kim"
date: "2023-11-05"
output:
        rmdformats::downcute:
        downcute_theme: "chaos"
code_folding: show
fig_width: 6
fig_height: 6
df_print: paged
editor_options: 
        chunk_output_type: inline
markdown: 
        wrap: 72
---
        

        
```{r warning=FALSE, message=FALSE, echo=FALSE, results='hide', setup}
#===============================================================================
#BTC.LineZero.Header.1.1.0
#===============================================================================
#R Markdown environment setup and reporting utility.
#===============================================================================
#RLB.Dependencies:
#   knitr, magrittr, pacman, rio, rmarkdown, rmdformats, tibble, yaml
#===============================================================================
#Input for document parameters, libraries, file paths, and options.
#=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=
knitr::opts_chunk$set(message=FALSE, warning = FALSE)

path_working <- "/Users/minsikkim/Dropbox/Inha/5_Lectures/Advanced biostatistics/scripts/BTE3207_Advanced_Biostatistics/"
path_library <- "/Library/Frameworks/R.framework/Resources/library"
str_libraries <- c("tidyverse", "pacman", "yaml")



YAML_header <-
        '---
title: "BTE3207 week 11-1"
author: "Minsik Kim"
date: "2032.11.5"
output:
    rmdformats::downcute:
        downcute_theme: "chaos"
        code_folding: hide
        fig_width: 6
        fig_height: 6
---'
seed <- "20230920"

#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Loads libraries, file paths, and other document options.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Boot <- function() {
        .libPaths(path_library)
        
        require(pacman)
        pacman::p_load(c("knitr", "rmarkdown", "rmdformats", "yaml"))
        
        knitr::opts_knit$set(root.dir = path_working)
        
        str_libraries |> unique() |> sort() -> str_libraries
        pacman::p_load(char = str_libraries)
        
        set.seed(seed)
}
FUN.LineZero.Boot()
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Outputs R environment report.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Report <- function() {
        cat("Line Zero Environment:\n\n")
        paste("R:", pacman::p_version(), "\n") |> cat()
        cat("Libraries:\n")
        for (str_libraries in str_libraries) {
                paste(
                        "    ", str_libraries, ": ", pacman::p_version(package = str_libraries),
                        "\n", sep = ""
                ) |> cat()
        }
        paste("\nOperating System:", pacman::p_detectOS(), "\n") |> cat()
        paste("    Library Path:", path_library, "\n") |> cat()
        paste("    Working Path:", path_working, "\n") |> cat()
        paste("Seed:", seed, "\n\n") |> cat()
        cat("YAML Header:\n")
        cat(YAML_header)
}
FUN.LineZero.Report()


```

# MPG dataset

`mpg` dataset (miles per gallon, about fuel efficiency) is default-data installed in `ggplot2` package.

```{r}
library(ggplot2)

head(mpg)

```

We can se there are multiple types of car-classes.

```{r}

mpg$class %>% table

```

Is there a difference between 2-seater vs regular (midsize) cars in city mpg?

We can run a t-test!

```{r}
city_coup <- mpg %>% filter(class == "2seater") %>% .$cty
city_midsize <- mpg %>% filter(class == "midsize") %>% .$cty
city_coup
city_midsize
        
t.test(city_coup, city_midsize, paired = F, var.equal = F)        
```

p-value is really low! 

So, if a car have all the same feature, and if they are having different numbers of seat, does MPG goes down?


It does not make sense. But why dis happend in this dataset?

Lets see a scatter plot of them

```{r}

mpg %>% 
        ggplot(., aes(y = cty, x = displ)) +
        geom_point() +
        theme_classic() +
        ylab("MPG at urban environment") +
        xlab("Engine size")
        

```

You can see a negative correlation between `displ` (Engine size) and `cty` (MPG in urban area).

And if we overlay the information of  `class` on this plot,

```{r}

mpg %>% 
        ggplot(., aes(y = cty, x = displ, col = class)) +
        geom_point() +
        theme_classic() +
        ylab("MPG at urban environment") +
        scale_color_brewer(type = "qual") +
        xlab("Engine size")

```

```{r}

data_regular_2seat %>% 
        ggplot(., aes(y = cty, col = displ, x = class)) +
        geom_point() +
        theme_classic() +
        ylab("MPG at urban environment") +
        scale_color_continuous(type = "viridis") +
        xlab("Car type")

```



```{r}


data_regular_2seat %>% 
        ggplot(., aes(y = cty, col = displ, x = class)) +
        geom_point() +
        theme_classic() +
        ylab("MPG at urban environment") +
        #scale_color_brewer(type = "qual") +
        xlab("Engine size")


lm(data = data_regular_2seat, cty ~ displ + class) %>% summary
lm(data = data_regular_2seat, cty ~ displ * class) %>% summary



```


```{r}


mpg %>% 
        ggplot(., aes(y = cty, col = manufacturer, x = displ)) +
        geom_point() +
        theme_classic() +
        ylab("MPG at urban environment") +
        #scale_color_brewer(type = "qual") +
        xlab("Engine size")



mpg %>% 
        ggplot(., aes(y = cty, col = manufacturer, x = displ)) +
        geom_point() +
        theme_classic() +
        ylab("MPG at urban environment") +
        #facet_wrap(~manufacturer) + 
        geom_smooth(method = "lm", 
              aes(color = manufacturer), 
              se = FALSE) + 
        #scale_color_brewer(type = "qual") +
        xlab("Engine size")


```

```{r}

lm(cty ~ displ + manufacturer + displ * manufacturer,
   data = mpg) %>% 
        summary


```

Or we can stratifiy the analysis

```{r}
mpg %>% 
        ggplot(., aes(y = cty, col = manufacturer, x = displ)) +
        geom_point() +
        theme_classic() +
        ylab("MPG at urban environment") +
        facet_wrap(~manufacturer) + 
        geom_smooth(method = "lm", 
              aes(color = manufacturer), 
              se = FALSE) + 
        #scale_color_brewer(type = "qual") +
        xlab("Engine size")


```


```{r}

mpg %>% filter(manufacturer == "toyota") %>% 
        lm(cty ~ displ, data = .) %>% 
        summary

mpg %>% filter(manufacturer == "volkswagen") %>% 
        lm(cty ~ displ, data = .) %>% 
        summary


```


Generally, 2setaer was having higher engine size!

As 

# Before begin..

Let's load the SBP dataset.

```{r}

dataset_sbp <- read.csv(file = "/Users/minsikkim/Dropbox/Inha/5_Lectures/Advanced biostatistics/scripts/BTE3207_Advanced_Biostatistics/dataset/sbp_dataset_korea_2013-2014.csv")



head(dataset_sbp)

```


Making a new variable `hypertension`

```{r}

dataset_sbp$hypertension <- ifelse(dataset_sbp$SBP > 130 |
                                           dataset_sbp$DBP > 80,
                                   T,
                                   F)


```

# DBP ~ SBP

With a scatter plot, we can see there is a correlation between SBP and DBP.

```{r}

set.seed(1)
dataset_sbp_small <- subset(dataset_sbp, row.names(dataset_sbp) %in% sample(x = 1:1000000, size = 1000))

ggplot(data = dataset_sbp_small, aes(x = SBP, y = DBP)) +
        geom_point() +
        theme_classic( base_family = "serif", base_size = 20) +
        xlab("SBP (mmHg)") +
        ylab("DBP (mmHg)")
        



```

# Data with trendline

With `lm()` function, we can calculate the linear model of DBP~SBP with the least squares algorithm.

```{r}

#A function that returns equation

lm_eqn <- function(m){
    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2, 
         list(a = format(unname(coef(m)[1]), digits = 2),
              b = format(unname(coef(m)[2]), digits = 2),
             r2 = format(summary(m)$r.squared, digits = 3)))
    as.character(as.expression(eq));
}


set.seed(1)

dataset_sbp_small <- subset(dataset_sbp, row.names(dataset_sbp) %in% sample(x = 1:1000000, size = 1000))

ggplot(data = dataset_sbp_small, aes(x = SBP, y = DBP)) +
        geom_point() +
        theme_classic(base_family = "serif", base_size = 20) +
        xlab("SBP (mmHg)") +
        ylab("DBP (mmHg)") + 
        geom_smooth(method = "lm") +
        geom_label(x = 130, y = 115,
                  label = lm_eqn(lm(data = dataset_sbp_small, DBP ~ SBP)),
                  parse = TRUE,
                  family='serif',
                  size = 8, label.size = 0, color = "blue")




#rnorm(100, mean = 100, sd = 1)

```

# Categorized outcome

```{r}

dataset_sbp_small %>%
        ggplot(., aes(x = BTH_G, y = SBP)) +
        geom_jitter()

lm(SBP ~ BTH_G, data = dataset_sbp_small) %>% summary


lm(SBP ~ BTH_G + DIS, data = dataset_sbp_small) %>% summary

lm(SBP ~ BTH_G, data = dataset_sbp_small %>% subset(., .$DIS == 1)) %>% summary
lm(SBP ~ BTH_G, data = dataset_sbp_small %>% subset(., .$DIS == 2)) %>% summary
lm(SBP ~ BTH_G, data = dataset_sbp_small %>% subset(., .$DIS == 3)) %>% summary        
lm(SBP ~ BTH_G, data = dataset_sbp_small %>% subset(., .$DIS == 4)) %>% summary        

lm(SBP ~ BTH_G + BMI + FBS + SEX, data = dataset_sbp_small %>% subset(., .$DIS == 4)) %>% summary        

```

# Bibliography

```{r warning=FALSE, message=FALSE, echo=FALSE}
#===============================================================================
#BTC.LineZero.Footer.1.1.0
#===============================================================================
#R markdown citation generator.
#===============================================================================
#RLB.Dependencies:
#   magrittr, pacman, stringr
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#BTC.Dependencies:
#   LineZero.Header
#===============================================================================
#Generates citations for each explicitly loaded library.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
str_libraries <- c("r", str_libraries)
for (str_libraries in str_libraries) {
        str_libraries |>
                pacman::p_citation() |>
                print(bibtex = FALSE) |>
                capture.output() %>%
                .[-1:-3] %>% .[. != ""] |>
                stringr::str_squish() |>
                stringr::str_replace("_", "") |>
                cat()
        cat("\n")
}
#===============================================================================
```
