---
title: "COD_week13_2_MGK_BTE3207"
author: "Minsik Kim"
date: "2024-11-24"
output:
        rmdformats::downcute:
        downcute_theme: "chaos"
code_folding: show
fig_width: 6
fig_height: 6
df_print: paged
editor_options: 
        chunk_output_type: inline
markdown: 
        wrap: 72
---
        
        
        
```{r warning=FALSE, message=FALSE, echo=FALSE, results='hide', setup}
#===============================================================================
#BTC.LineZero.Header.1.1.0
#===============================================================================
#R Markdown environment setup and reporting utility.
#===============================================================================
#RLB.Dependencies:
#   knitr, magrittr, pacman, rio, rmarkdown, rmdformats, tibble, yaml
#===============================================================================
#Input for document parameters, libraries, file paths, and options.
#=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=
knitr::opts_chunk$set(message=FALSE, warning = FALSE)

path_working <- 
        ifelse(sessionInfo()[1]$R.version$platform == "x86_64-pc-linux-gnu",
               "/mnt/4T_samsung/Dropbox/",
               ifelse(sessionInfo()[1]$R.version$platform == "aarch64-apple-darwin20",
                      "/Volumes/macdrive/Dropbox/", 
                      "/Users/minsikkim/Dropbox (Personal)/"))
path_library <- 
        ifelse(sessionInfo()[1]$R.version$platform == "x86_64-pc-linux-gnu",
               "/home/bagel/R_lib/",
               "/Library/Frameworks/R.framework/Resources/library/")


str_libraries <- c("tidyverse", "pacman", "yaml", "reactable")



YAML_header <-
        '---
title: "BTE3207 week 13-2"
author: "Minsik Kim"
date: "2024.11.25"
output:
    rmdformats::downcute:
        downcute_theme: "chaos"
        code_folding: hide
        fig_width: 6
        fig_height: 6
---'
seed <- "20241125"

#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Loads libraries, file paths, and other document options.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Boot <- function() {
        .libPaths(path_library)
        
        require(pacman)
        pacman::p_load(c("knitr", "rmarkdown", "rmdformats", "yaml", "lmerTest"))
        
        knitr::opts_knit$set(root.dir = path_working)
        
        str_libraries |> unique() |> sort() -> str_libraries
        pacman::p_load(char = str_libraries)
        
        set.seed(seed)
}
FUN.LineZero.Boot()
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Outputs R environment report.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Report <- function() {
        cat("Line Zero Environment:\n\n")
        paste("R:", pacman::p_version(), "\n") |> cat()
        cat("Libraries:\n")
        for (str_libraries in str_libraries) {
                paste(
                        "    ", str_libraries, ": ", pacman::p_version(package = str_libraries),
                        "\n", sep = ""
                ) |> cat()
        }
        paste("\nOperating System:", pacman::p_detectOS(), "\n") |> cat()
        paste("    Library Path:", path_library, "\n") |> cat()
        paste("    Working Path:", path_working, "\n") |> cat()
        paste("Seed:", seed, "\n\n") |> cat()
        cat("YAML Header:\n")
        cat(YAML_header)
}
FUN.LineZero.Report()




```


# Agenda

This lecture will practice running 

1. Plackett-Burman Design (PBD)

and

2. Response surface method (RSM)


# Before begin..

Let's install packages.

```{r}
install.packages("daewr")

library(daewr)

```

`PBDes` function creates the experimental design of the PBD.
randomize = T will generate a matrix with mixed orders.


```{r}

daewr::PBDes(nfactors = 11, nruns = 12)

# Error examples
#  daewr::PBDes(nfactors = 12, nruns = 12) 


daewr::PBDes(nfactors = 10, nruns = 12)

#daewr::PBDes(nfactors = 15, nruns = 24)

```


After setting your `higher level` and `lower level` based on your intuition, you can run a experiment with 12 experimental groups, and get the result out of the data.

analyiss of PBD outcome

You can run a miultiple linear regression that is containing all the data!


```{r}
library(tidyverse)
#<<<<<<< HEAD

download.file("https://raw.githubusercontent.com/minsiksudo/BTE3207_Advanced_Biostatistics/refs/heads/main/PBD_example_data.csv", "PBD_example_data.csv")

pbd_data <- read.csv("PBD_example_data.csv")

```

```{r}
 lm(Biomass ~ Glucose + YE + KH2PO4 + K2HPO4 + MgSO4 + CaCl2 + FeSO4 + Fructose + NH4Cl + TM + Vitamin,
#<<<<<<< HEAD
    data = pbd_data) %>% summary

```

This is your Plackett-Burman result.


## RSM


```{r}
install.packages("rsm")
library(rsm)

```


using `ccd` function, you can generate a experimental design that can be used for RSM experiment and analysis

```{R}

#2 factor experiment
ccd (2)
#3 factor experiment
ccd (3)

SOdes2 <- ccd (2, n0 = c(4,6), alpha = "rotatable", inscribed = F)

SOdes2

```


With `coding` argument, you can set the levels of your data into actual values

```{r}
SOdes2 <- ccd (3, n0 = c(4,6), alpha = "rotatable", inscribed = F,
               coding = list (
                x1 ~ (Glucose - 50)/10, 
                x2 ~ (NaNO3 - 5)/2, 
                x3 ~ (K2HPO4 - 2)/1)
)

SOdes2
```

# Analysis of RSM


```{r}

CR1.rsm <- rsm(Yield ~ Block + SO(Time, Temp), data = ChemReact)

summary(CR1.rsm)

```

Here, besides the interaction term, all the term is significant. 

Anova here tests the "lack of fit" which tests the variance not explaned by model but by the error (from the repeated experiments). This term's p-values should be greater than effect of others!


To draw contour, use `contour` function

```{r}

contour(CR1.rsm, ~ Time + Temp, image = TRUE, at = summary(CR1.rsm)$canonical$xs)


```

To draw 3d plot, use `persp` function

```{r}

persp(CR1.rsm, 
      Time ~ Temp, #plot xs
      col=rainbow(50),
      xlabs = c(expression("Time (min)", "Temperature (°C)")), # axis labels
      theta=30, 
      phi=30, r = 120, d=120, 
      border = NULL, 
      ltheta = 0, 
      lphi = 0, 
              shade = 0.75, zlab="Yield", col.axis=37, font.lab=2, col.lab=35,
              contour=("colors"))


```

Changing theta, phi, and r can rotate the graph

```{r}

persp(CR1.rsm, 
      Time ~ Temp, #plot xs
      col=rainbow(50),
      xlabs = c(expression("Time (min)", "Temperature (°C)")), # axis labels
      theta=50, 
      phi=3, r = 180, d=120, 
      border = NULL, 
      ltheta = 0, 
      lphi = 0, 
              shade = 0.75, zlab="Yield", col.axis=37, font.lab=2, col.lab=35,
              contour=("colors"))


```


# Bibliography

```{r warning=FALSE, message=FALSE, echo=FALSE}
#===============================================================================
#BTC.LineZero.Footer.1.1.0
#===============================================================================
#R markdown citation generator.
#===============================================================================
#RLB.Dependencies:
#   magrittr, pacman, stringr
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#BTC.Dependencies:
#   LineZero.Header
#===============================================================================
#Generates citations for each explicitly loaded library.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
str_libraries <- c("r", str_libraries)
for (str_libraries in str_libraries) {
        str_libraries |>
                pacman::p_citation() |>
                print(bibtex = FALSE) |>
                capture.output() %>%
                .[-1:-3] %>% .[. != ""] |>
                stringr::str_squish() |>
                stringr::str_replace("_", "") |>
                cat()
        cat("\n")
}
#===============================================================================
```
