---
title: "COD_week13_1_MGK_BTE3207"
author: "Minsik Kim"
date: "2023-11-20"
output:
        rmdformats::downcute:
        downcute_theme: "chaos"
code_folding: show
fig_width: 6
fig_height: 6
df_print: paged
editor_options: 
        chunk_output_type: inline
markdown: 
        wrap: 72
---
        
        
        
```{r warning=FALSE, message=FALSE, echo=FALSE, results='hide', setup}
#===============================================================================
#BTC.LineZero.Header.1.1.0
#===============================================================================
#R Markdown environment setup and reporting utility.
#===============================================================================
#RLB.Dependencies:
#   knitr, magrittr, pacman, rio, rmarkdown, rmdformats, tibble, yaml
#===============================================================================
#Input for document parameters, libraries, file paths, and options.
#=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=
knitr::opts_chunk$set(message=FALSE, warning = FALSE)

path_working <- "/Users/minsikkim/Dropbox (Personal)/Inha/5_Lectures/Advanced biostatistics/scripts/BTE3207_Advanced_Biostatistics/"
path_library <- "/Library/Frameworks/R.framework/Resources/library"
str_libraries <- c("tidyverse", "pacman", "yaml")



YAML_header <-
        '---
title: "BTE3207 week 13-1"
author: "Minsik Kim"
date: "2032.11.20"
output:
    rmdformats::downcute:
        downcute_theme: "chaos"
        code_folding: hide
        fig_width: 6
        fig_height: 6
---'
seed <- "20230920"

#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Loads libraries, file paths, and other document options.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Boot <- function() {
        .libPaths(path_library)
        
        require(pacman)
        pacman::p_load(c("knitr", "rmarkdown", "rmdformats", "yaml", "lmerTest", "vegan", "phyloseq", "ggplot"))
        
        knitr::opts_knit$set(root.dir = path_working)
        
        str_libraries |> unique() |> sort() -> str_libraries
        pacman::p_load(char = str_libraries)
        
        set.seed(seed)
}
FUN.LineZero.Boot()
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Outputs R environment report.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Report <- function() {
        cat("Line Zero Environment:\n\n")
        paste("R:", pacman::p_version(), "\n") |> cat()
        cat("Libraries:\n")
        for (str_libraries in str_libraries) {
                paste(
                        "    ", str_libraries, ": ", pacman::p_version(package = str_libraries),
                        "\n", sep = ""
                ) |> cat()
        }
        paste("\nOperating System:", pacman::p_detectOS(), "\n") |> cat()
        paste("    Library Path:", path_library, "\n") |> cat()
        paste("    Working Path:", path_working, "\n") |> cat()
        paste("Seed:", seed, "\n\n") |> cat()
        cat("YAML Header:\n")
        cat(YAML_header)
}
FUN.LineZero.Report()


```


# Agenda

This lecture will practice running 

1. simple linear models

2. multiple linear regression

3. Effect modification

4. ANOVA


# Before begin..

Let's load datasets.

## the SBP dataset


```{r}

dataset_sbp <- read.csv(file = "/Users/minsikkim/Dropbox (Personal)/Inha/5_Lectures/Advanced biostatistics/scripts/BTE3207_Advanced_Biostatistics/dataset/sbp_dataset_korea_2013-2014.csv")



head(dataset_sbp)

```

### Data manipulation

Making a new variable `hypertension`

```{r}

dataset_sbp$hypertension <- ifelse(dataset_sbp$SBP > 130 |
                                           dataset_sbp$DBP > 80,
                                   T,
                                   F)


dataset_sbp$history_of_hypertension <- ifelse(dataset_sbp$DIS == 1 |
                                           dataset_sbp$DIS == 2,
                                   T,
                                   F)
dataset_sbp

```

```{r}

dataset_sbp$DIS <- factor(dataset_sbp$DIS, 
                          levels = c(1,2,3,4),
                          labels = c("Hypertension and diabetes",
                                     "Hypertension",
                                     "Diabetes",
                                     "No history"))

set.seed(1)

dataset_sbp_small <- subset(dataset_sbp,
                        row.names(dataset_sbp)%in% sample(x = 1:1000000, size = 1000))


```


### UCLA school dataset

https://stats.idre.ucla.edu/

Dataset is students from 10 handpicked schools, representing a subset of students and schools from a US survey of eight-grade students at 1000 schools (800 public 200 private).

There are quite a lot of variables in the dataset, but we are going start by using just three:

```

 # Variable Type Len Pos Label
-----------------------------------------------------------------------------------------------
15 cstr     Num    8 112
5 homework Num    8  32 Time spent on math homework each week
11 math     Num    8  80 Math score
4 meanses  Num    8  24 Mean SES for the school
7 parented Num    8  48 Parents highest education level

10 percmin  Num    8  72 Percent minority in school
 8 public   Num    8  56 Public school: 1=public, 0=non-public
13 race     Num    8  96 race of student, 1=asian, 2=Hispanic, 3=Black, 4=White, 5=Native American
 9 ratio    Num    8  64 Student-Teacher ratio
18 region   Num    8 136
 1 schid    Num    8   0 School ID
19 schnum   Num    8 144 group(schid)
16 scsize   Num    8 120
14 sctype   Num    8 104 Type of school, 1=public, 2=catholic, 3=Private
                         other religious, 4=Private non-r
 3 ses      Num    8  16 Socioecnonomic Status
12 sex      Num    8  88 Sex: 1=male, 2=female
 2 stuid    Num    8   8 Student ID
17 urban    Num    8 128
 6 white    Num    8  40 Race: 1=white, 0=non-white
 ```
 
 Here,

*homework*: number of hours of homework - Level 1 variable (associated with students)
*math*: score in a math test - expanatory variable
*schnum*: a number for each school - Level 2 variable (associated with school)
*scsize*: school size - Level 2 variable
*public*: school type - Level 2 variable


```{r}

install.packages("haven")
library(haven)

imm10_data <- read_dta("https://stats.idre.ucla.edu/stat/examples/imm/imm10.dta") 
imm10_data
```


# Before analyzing..


You need to double-check if the regression is appropriate or not.

# Simple linear regression

## Step 1

You need to store the linear model first.


```{r}


lm(SBP ~ DBP, data = dataset_sbp_small)

lm(data = dataset_sbp_small, 
                  SBP ~ DBP)

sbp_dbp_fit <- lm(data = dataset_sbp_small, 
                  SBP ~ DBP)

sbp_dbp_fit

sbp_dbp_fit %>% summary

```

You can run ANOVA on the anlaysis result.


```{r}

sbp_dbp_fit %>% anova

```

From the stored data, you can extract multiple summarized informaiton.


```{r}
sbp_dbp_fit

names(sbp_dbp_fit)
# or
sbp_dbp_fit %>% names


```

You can choose the model using

```{r}
sbp_dbp_fit

```

To laod the beta values,

```{r}
sbp_dbp_fit$coefficients
```




Choosing the numeric values in the lm() output can be conducted using double-brackets

```{r}
sbp_dbp_fit$coefficients[[1]]
sbp_dbp_fit$coefficients[[2]]

```

...etc.


## Step 2

You can generate a plot with linear regression



```{r}

plot(SBP ~ DBP, data = dataset_sbp_small)

plot(data = dataset_sbp_small,
     SBP ~ DBP) +
abline(sbp_dbp_fit, col = "red")




```

```{r}

ggplot(data = dataset_sbp_small, 
       aes(x = DBP, y = SBP)) +
        geom_boxplot() 

ggplot(data = dataset_sbp_small, 
       aes(x = DBP, y = SBP)) +
        geom_violin() 



ggplot(data = dataset_sbp_small, 
       aes(x = DBP, y = SBP)) +
        geom_point() +
        geom_smooth(method = "lm")



```


Well, it seems like the data fits well.... but does the model is actually representing the association vetween SBP vs DBP or just the data?

### Wroing example 

Let's assuem that our data is somehow having more data at the lower-end


```{r}

hw_math_fit <- lm(data = imm10_data, math ~ homework)

plot(math ~ homework, data = imm10_data) +
abline(hw_math_fit, col = "red") +
title("Homework vs Math")

```



```{r}
imm10_data$urban
factor(imm10_data$public)

ggplot2::ggplot(imm10_data,
        aes(y = math, 
            x = homework, 
            col = factor(urban),
            shape = factor(public))) +
        geom_point() 
        
```

Do you think the data is normally distributed?

## Double-checking the distribution of data

SBP dataset

```{r}

hist(dataset_sbp_small$SBP)
```
DBP data

```{r}
hist(dataset_sbp_small$DBP)
```
Homework data of urban schools

```{r}

imm10_data$homework %>% hist


imm10_data %>% filter(urban == 1) %>% .$homework %>% hist
```
Urban schools had right skewed data.

Do we need to generate these histograms, every time, for every variables that we are using?



### Using plot() function

Using plot funcion you can explore data at once

```{r}
plot(imm10_data)
imm10_data %>% plot

```

# Using GGrally package

An improved version of those associations, can be double-checked using the below package



```{r}

install.packages("GGally")

library(GGally)

ggpairs(imm10_data)


```

# Adding normal curve to histogram

How can we compare them with normal distribution?

We can add normal curve to the data

```{r}

hist(dataset_sbp_small$SBP, freq = F)
curve(dnorm(x,
            mean= mean(dataset_sbp_small$SBP), 
            sd=sd(dataset_sbp_small$SBP)), 
      col="darkblue", lwd=2, add=TRUE)

```

SBP was quite normal.





```{r}

hist(dataset_sbp_small$DBP, freq = F)
curve(dnorm(x,
            mean= mean(dataset_sbp_small$DBP), 
            sd=sd(dataset_sbp_small$DBP)), 
      col="darkblue", lwd=2, add=TRUE)

```

DBP was somwhat normal..

```{r}

hist(imm10_data$homework, freq = F, breaks = 9)
curve(dnorm(x,
            mean= mean(imm10_data$homework), 
            sd=sd(imm10_data$homework)), 
      col="darkblue", lwd=2, add=TRUE)

```
Not normal at all

# Deciding data is normal

We can generate a quantile of normal distribution and that of data! It woll show the different between normal curve and samples in more intuitive way.

Here, the x-axis is the quantile of a normal distribution of having the same mean and SD with dataset.

```{r}


qqnorm(rnorm(n = 10000), pch = 1, frame = FALSE)
qqline(rnorm(n = 10000),
       col = "red", lwd = 2)


```


## Q-Q plot of normal data


## Q-Q plot of SBP dataset

With SBP data, it looks like this.

```{r}
qqnorm(dataset_sbp_small$SBP, pch = 1, frame = FALSE)
qqline(dataset_sbp_small$SBP,
       col = "red", lwd = 2)
```
Here, the x-axis is the quantile of a normal distribution of having the same mean and SD with SBP dataset.

## Q-Q plot of DBP dataset

With DBP data, it looks like this.

```{r}
qqnorm(dataset_sbp_small$DBP, pch = 1, frame = FALSE)
qqline(dataset_sbp_small$DBP,
       col = "red", lwd = 2)
```
DBP is less normal than SBP.
The error seems to high at higher quantile values.

## Q-Q plot of homework

```{r}
qqnorm(imm10_data$homework, pch = 1, frame = FALSE)
qqline(imm10_data$homework,
       col = "red", lwd = 2)
```
## Q-Q of math test results

```{r}
qqnorm(imm10_data$math, pch = 1, frame = FALSE)
qqline(imm10_data$math,
       col = "red", lwd = 2)
```
Obviously, the data showed unreasonable outcome.

Does it mean that our data is useless?


# Evaluation after running lme

We can also plot the results of the linear model.




```{r}

par(mfrow=c(2,2))
plot(sbp_dbp_fit)
par(mfrow=c(1,1))

```

Here, 

1) residual vs fillted shows *the linearity between two variables*

i.e., it investigates weather the residual is uniform across all the independent variable.

2) Q-Q plot shows *the normality of both data*

This plots the *qunatile* of residuals. The residuals alsho has to have normal distribution. The theoretical residual can be calculated by sqrt(MSE) * z((k-0.375/(n+0.25))).
Where the MSE is mean squares, k is the rank, n is the total sample number, and the z is tne normal distribtion's pnorm().

3) Standardized version of plot 1 (variable omn the same scale).

4) residual v leverage plot. This calculates which one is outlier (based on assumption that the data is normally distributed).We only have 3 outliars for SBP data.


Doing the same thing for imm-dataset,

```{r}
par(mfrow=c(2,2))
plot(hw_math_fit)
par(mfrow=c(1,1))

```

Although the distribution of data was not normal, the residuals were normal. So we can use the dataset.


# Multiple linear regression

Now we are going to run multiple linear regression using R

```{r}
sbp_dbp_fit_mlr <- lm(data = dataset_sbp_small, SBP ~ DBP + BMI + SEX + DIS)
```

Running anova

```{r}
sbp_dbp_fit_mlr %>% anova
```

Gender is irrelevant to our analysis - for now I am going to remove it from the analysis

view summary

```{r}

sbp_dbp_fit_mlr %>% summary

```


Can run the same thing with mlr 

```{r}
par(mfrow=c(2,2))
plot(sbp_dbp_fit_mlr)
par(mfrow=c(1,1))

```
To make `adjusted plot`, You can choose the data as below. 

```{r}

sbp_dbp_fit_mlr$coefficients["(Intercept)"][[1]]
sbp_dbp_fit_mlr$coefficients["DBP"][[1]]

```

Using the adjusted coefficients of the linear model, you can draw linear model.


```{r}


plot(data = dataset_sbp_small, SBP~DBP,
     main = "Simple linear model"
     ) +
        abline(a = sbp_dbp_fit$coefficients["(Intercept)"][[1]],
               b = sbp_dbp_fit$coefficients["DBP"][[1]], 
               col = "blue",
               linewidth = 2) 

plot(data = dataset_sbp_small, SBP~DBP,
     main = "BMI, DIS adjusted linear model\n(of BMI = 0, DIS group 1)"
     ) +
        abline(a = sbp_dbp_fit_mlr$coefficients["(Intercept)"][[1]],
               b = sbp_dbp_fit_mlr$coefficients["DBP"][[1]], 
               col = "red",
               linewidth = 2) 



```


# Effect modificaitons 

## Interaction terms

Interaction term can be added easily

```{r}
sbp_dbp_fit_int <- lm(data = dataset_sbp_small, SBP ~ DBP + BMI + SEX * DIS)
```

Running anova

```{r}
sbp_dbp_fit_int %>% anova
```

Gender is irrelevant to our analysis - for now I am going to remove it from the analysis

view summary

```{r}

sbp_dbp_fit_int %>% summary

```


Can run the same thing with mlr 

```{r}
par(mfrow=c(2,2))
plot(sbp_dbp_fit_mlr)
par(mfrow=c(1,1))

```
To make `adjusted plot`, You can choose the data as below. 

```{r}

sbp_dbp_fit_mlr$coefficients["(Intercept)"][[1]]
sbp_dbp_fit_mlr$coefficients["DBP"][[1]]

```

Using the adjusted coefficients of the linear model, you can draw linear model.

In ggplot package, geom_smooth(method = "lm") will generate linear regression of each data groups at once!

```{r}


ggplot2::ggplot(dataset_sbp_small, aes(y = SBP, x = DBP, col = DIS)) +
        geom_point() +
        geom_smooth(method = "lm")

```

Other example would be...

```{r}


ggplot2::ggplot(imm10_data, aes(y = math, x = homework, col = as.factor(schnum))) +
        geom_point() +
        geom_smooth(method = "lm")

```


## Stratification


### Manucal

Data can be splited using following filter() function


```{r}

data_dis_1 <- dataset_sbp_small %>% filter(as.numeric(dataset_sbp_small$DIS) == 1)
data_dis_2 <- dataset_sbp_small %>% filter(as.numeric(dataset_sbp_small$DIS) == 2)
data_dis_3 <- dataset_sbp_small %>% filter(as.numeric(dataset_sbp_small$DIS) == 3)
data_dis_4 <- dataset_sbp_small %>% filter(as.numeric(dataset_sbp_small$DIS) == 4)

```


```{r}

data_dis_1 %>%
        lm(SBP ~ DBP, data = .) %>% summary

data_dis_2 %>%
        lm(SBP ~ DBP, data = .) %>% summary

data_dis_3 %>%
        lm(SBP ~ DBP, data = .) %>% summary

data_dis_4 %>%
        lm(SBP ~ DBP, data = .) %>% summary


```

QQplot........skipping some


```{r}

data_dis_1 %>%
        lm(SBP ~ DBP, data = .) %>%
        plot()


```

Plotting.. You can do that one by one. Else, the blow code can be used.

```{r}


ggplot2::ggplot(dataset_sbp_small, aes(y = SBP, x = DBP, col = DIS)) +
        geom_point() +
        geom_smooth(method = "lm") +
        facet_wrap(~DIS)

```



# Linaer mixed effect models

```{r}

math_hwk_lmer <- lme4::lmer(data = imm10_data, math ~ homework + (1|schid))

```

```{r}

math_hwk_lmer %>% anova

```

This one does not result in p-values!


```{r}

math_hwk_lmer <- lmerTest::lmer(data = imm10_data, math ~ homework + (1|schid))

```

```{r}

math_hwk_lmer %>% anova

```

`lmerTest` package need to be used instead.


```{r}

plot(math_hwk_lmer)



```
To calculate QQ plot of residuals,

```{r}
qqnorm(resid(math_hwk_lmer), main = "Q-Q residuals", ylab = "Standardized residuals")
qqline(resid(math_hwk_lmer))

```

as the lmer output (`merMod` object) is not a `list`. (They are S4 objects)

# Bibliography

```{r warning=FALSE, message=FALSE, echo=FALSE}
#===============================================================================
#BTC.LineZero.Footer.1.1.0
#===============================================================================
#R markdown citation generator.
#===============================================================================
#RLB.Dependencies:
#   magrittr, pacman, stringr
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#BTC.Dependencies:
#   LineZero.Header
#===============================================================================
#Generates citations for each explicitly loaded library.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
str_libraries <- c("r", str_libraries)
for (str_libraries in str_libraries) {
        str_libraries |>
                pacman::p_citation() |>
                print(bibtex = FALSE) |>
                capture.output() %>%
                .[-1:-3] %>% .[. != ""] |>
                stringr::str_squish() |>
                stringr::str_replace("_", "") |>
                cat()
        cat("\n")
}
#===============================================================================
```
