---
title: "COD_week7_1_MGK_BTE3207"
author: "Minsik Kim"
date: "2024-10-15"
output:
        rmdformats::downcute:
        downcute_theme: "chaos"
code_folding: show
fig_width: 6
fig_height: 6
df_print: paged
editor_options: 
        chunk_output_type: inline
markdown: 
        wrap: 72
---
        
<style>
  /* Default light mode styles */
  .reactable {
    background-color: #ffffff !important; /* Light background */
    color: #000000 !important;            /* Dark text */
    border-color: #cccccc !important;     /* Light border */
  }

</style>
        
```{r warning=FALSE, message=FALSE, echo=FALSE, results='hide', setup}
#===============================================================================
#BTC.LineZero.Header.1.1.0
#===============================================================================
#R Markdown environment setup and reporting utility.
#===============================================================================
#RLB.Dependencies:
#   knitr, magrittr, pacman, rio, rmarkdown, rmdformats, tibble, yaml
#===============================================================================
#Input for document parameters, libraries, file paths, and options.
#=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=
knitr::opts_chunk$set(message=FALSE, warning = FALSE)

path_working <- 
        ifelse(sessionInfo()[1]$R.version$platform == "x86_64-pc-linux-gnu",
               "/mnt/4T_samsung/Dropbox/",
               ifelse(sessionInfo()[1]$R.version$platform == "aarch64-apple-darwin20",
                      "/Volumes/macdrive/Dropbox/", 
                      "/Users/minsikkim/Dropbox/"))
path_library <- 
        ifelse(sessionInfo()[1]$R.version$platform == "x86_64-pc-linux-gnu",
               "/home/bagel/R_lib/",
               "/Library/Frameworks/R.framework/Resources/library/")


str_libraries <- c("tidyverse", "pacman", "yaml", "reactable")



YAML_header <-
        '---
title: "BTE3207 week 7-1"
author: "Minsik Kim"
date: "2024.10.15"
output:
    rmdformats::downcute:
        downcute_theme: "chaos"
        code_folding: hide
        fig_width: 6
        fig_height: 6
---'
seed <- "20241015"

#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Loads libraries, file paths, and other document options.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Boot <- function() {
        .libPaths(path_library)
        
        require(pacman)
        pacman::p_load(c("knitr", "rmarkdown", "rmdformats", "yaml"))
        
        knitr::opts_knit$set(root.dir = path_working)
        
        str_libraries |> unique() |> sort() -> str_libraries
        pacman::p_load(char = str_libraries)
        
        set.seed(seed)
}
FUN.LineZero.Boot()
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Outputs R environment report.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Report <- function() {
        cat("Line Zero Environment:\n\n")
        paste("R:", pacman::p_version(), "\n") |> cat()
        cat("Libraries:\n")
        for (str_libraries in str_libraries) {
                paste(
                        "    ", str_libraries, ": ", pacman::p_version(package = str_libraries),
                        "\n", sep = ""
                ) |> cat()
        }
        paste("\nOperating System:", pacman::p_detectOS(), "\n") |> cat()
        paste("    Library Path:", path_library, "\n") |> cat()
        paste("    Working Path:", path_working, "\n") |> cat()
        paste("Seed:", seed, "\n\n") |> cat()
        cat("YAML Header:\n")
        cat(YAML_header)
}
FUN.LineZero.Report()


```


# Before begin..

Let’s load the SBP (Systolic Blood Pressure) dataset, which we will use throughout this document for various statistical analyses.


```{r}

read.csv(file = "Inha/5_Lectures/2024/Advanced biostatistics/scripts/BTE3207_Advanced_Biostatistics/dataset/sbp_dataset_korea_2013-2014.csv") %>%
        reactable::reactable(., sortable = T)


dataset_sbp <- read.csv(file = "Inha/5_Lectures/2024/Advanced biostatistics/scripts/BTE3207_Advanced_Biostatistics/dataset/sbp_dataset_korea_2013-2014.csv")

head(dataset_sbp)

```

# Creating a New Variable: Hypertension

Logical Variables in R:

Logical variables store TRUE or FALSE values. These variables are useful in decision-making processes (like identifying conditions where blood pressure exceeds a threshold). In R, the result of comparisons such as > or == is always logical (TRUE or FALSE).

Using ifelse() to Create Logical Variables:

The ifelse() function is a vectorized conditional function in R that checks a condition for each element in a vector. It returns one value if the condition is TRUE and another value if it is FALSE.

Syntax: ifelse(condition, value_if_TRUE, value_if_FALSE)

Below, we use ifelse() to create a logical variable called hypertension, which will be TRUE if either the systolic blood pressure (SBP) is greater than 130 or the diastolic blood pressure (DBP) is greater than 80, and FALSE otherwise.


```{r}
# Create 'hypertension' variable based on SBP and DBP

dataset_sbp$hypertension <- ifelse(dataset_sbp$SBP > 130 |
                                           dataset_sbp$DBP > 80,
                                   T,
                                   F)


```

In R, `|` is "OR" and  `&` is "AND".

In the code, `|` is the logical *OR* operator. The condition checks if either SBP > 130 *or* DBP > 80 is true.

	•	If either condition is true, the result will be TRUE.
	•	If both conditions are false, the result will be FALSE.


This new logical variable can now be used to classify individuals as hypertensive or non-hypertensive.



# Understanding Normal Distribution Functions

1. rnorm() – Random Sampling from a Normal Distribution

This function generates random numbers following a normal distribution.



```{r}

# Generate 5 random numbers from a standard normal distribution (mean = 0, sd = 1)
rnorm(5)

# Example with custom mean and standard deviation
# rnorm(100, mean = 100, sd = 1)

```


2. pnorm() – Cumulative Probability for Z-scores (z-score table results)

pnorm() gives the cumulative probability for a given z-score (i.e., the area under the curve up to that point).

```{r}
# Cumulative probabilities for different z-scores
pnorm(2)   # ~97.72% of the data lies below a z-score of 2
pnorm(0)   # 50% of the data lies below the mean
pnorm(1)   # ~84.13% of the data is below a z-score of 1
pnorm(3)   # ~99.87% of the data is below a z-score of 3

```


3. qnorm() – Find Z-scores from Cumulative Probability

qnorm() returns the z-score corresponding to a given cumulative probability.

```{r}
# Z-scores for given probabilities
qnorm(0.97724985)  # Returns ~2, the z-score for 97.72%
qnorm(0.5)         # Returns 0, the z-score for 50%
```

Note: `pnorm()` and `qnorm()` are *reverse* operations of each other.


```{r}

# Calculating density at different points
dnorm(0)  # Height at the mean
dnorm(1)  # Height at z = 1
dnorm(-1) # Height at z = -1

```

`dnorm()` - *probability density function*, the height of normal curve.

```{r}

dnorm(0)

dnorm(1)

dnorm(-1)

```

# Plotting the Normal Distribution

Plotting Discrete Values from dnorm()

```{r}
# Generate sequence and plot dnorm results
x_vals <- seq(-4, 4, by = 0.5)
plot(dnorm(x_vals) ~ x_vals, 
     ylab = "dnorm() result", 
     xlab = "dnorm() input")

```

Ploting of a smooth normal curve

```{r}
# Create a sequence of values and calculate their densities
x <- seq(-4, 4, length = 100)
y <- dnorm(x)

# Plot the normal curve with labeled x-axis
plot(x, y, type = "l", lwd = 2, axes = FALSE, xlab = "", ylab = "")
axis(1, at = -3:3, labels = c("-3σ", "-2σ", "-1σ", "mean", "1σ", "2σ", "3σ"))
abline(v = 2, col = "red")  # Add a vertical line at z = 2
abline(v = -2, col = "red") # Add a vertical line at z = -2

        
```

# Logical Variables and Statistical Tests

Now that we’ve created the hypertension variable as a logical variable, we can use it in statistical tests.

## Single-sample t-test (getting confidence interval of some data)

This tests if the mean of a sample is significantly different from a known value.

```{r}
t.test(c(1, 2, 3))


result <- t.test(dataset_sbp$SBP) # Extract the confidence interval 

result$p.value


result$conf.int


```

## Two-Sample t-test:

This tests if two groups have significantly different means.

Now that we’ve created the hypertension variable as a logical variable, we can use it in statistical tests.

Two-Sample t-Test Example:


```{r}

result <- t.test(SBP ~ SEX, data = dataset_sbp) # Extract the confidence interval 
result
result <- t.test(SBP ~ hypertension, data = dataset_sbp) # Extract the confidence interval 
result
```

# Z-Test for Proportions

## one-sample z-test using z.test()

Test if the proportion of a sample is different from a given proportion.


```{r}
# Install and load the BSDA package

install.packages("BSDA")

library(BSDA)


# Perform a one-sample z-test

result <- z.test(dataset_sbp$hypertension,
                 sigma.x = mean(dataset_sbp$hypertension) *
                         (1-mean(dataset_sbp$hypertension))
                 )


result

```


Hypertention table

```{r}
table(dataset_sbp$SEX)
table(dataset_sbp$DIS)

table(dataset_sbp$SEX, dataset_sbp$hypertension)


```

## Two-Sample Z-Test:

two-sample z-test

```{r}
# Subset the data by sex


library(tidyverse)

dataset_male <- dataset_sbp %>% subset(., .$SEX == 1)
dataset_female <-  dataset_sbp %>% subset(., .$SEX == 2)

# Perform two-sample z-test

result <- z.test(x = dataset_male$hypertension,
                 y = dataset_female$hypertension,
                 sigma.x = mean(dataset_male$hypertension) * (1-mean(dataset_male$hypertension)),
                 sigma.y = mean(dataset_female$hypertension) * (1-mean(dataset_female$hypertension))
                 )


result

```

# Chi-square test for independence

## Two-group Chi-square test

Check if two categorical variables are independent.

```{r}
# Chi-square test between hypertension and sex

result <- chisq.test(x = dataset_sbp$hypertension,
                     y = dataset_sbp$SEX)

result

```

Note: **p**-value is the same as z-score test

## Chi-square test for multiple groups

Check relationships between multiple categorical variables.

```{r}

table(dataset_sbp$SEX, dataset_sbp$DIS)

# Chi-square test between sex and disease status

result <- chisq.test(y = dataset_sbp$DIS,
           x = dataset_sbp$SEX)

result

```

## Fisher's exact test

Fisher’s exact test is used when sample sizes are small.

```{r}



#Author DataFlair

data_frame <- read.csv("https://goo.gl/j6lRXD")

#Reading CSV
data <- table(data_frame$treatment, data_frame$improvement)


fisher.test(data)

# or 

# Perform Fisher's exact test on the hypertension and sex table
fisher.test(table(dataset_sbp$SEX, dataset_sbp$hypertension))


```

```{r}

#Author DataFlair
table(dataset_sbp$SEX,
                  dataset_sbp$hypertension)

fisher.test(table(dataset_sbp$SEX,
                  dataset_sbp$hypertension))

table(dataset_sbp$SEX,
                  dataset_sbp$hypertension) %>%
        fisher.test()

```

As sample number is high, Fisher's exact test results are having the same result as Chi-square / z-test.


## Confidence intervals and statistical tests using epitools

`epitools` pacakge automates all the calculations and tests! Especially for risk ratios and odds ratios.


```{r}
# Install and load epitools
install.packages("epitools")
library(epitools)

# Calculate risk ratio with confidence interval
riskratio.wald(data)

# Calculate odds ratio with confidence interval
oddsratio.wald(data)

```

# Bibliography

```{r warning=FALSE, message=FALSE, echo=FALSE}
#===============================================================================
#BTC.LineZero.Footer.1.1.0
#===============================================================================
#R markdown citation generator.
#===============================================================================
#RLB.Dependencies:
#   magrittr, pacman, stringr
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#BTC.Dependencies:
#   LineZero.Header
#===============================================================================
#Generates citations for each explicitly loaded library.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
str_libraries <- c("r", str_libraries)
for (str_libraries in str_libraries) {
        str_libraries |>
                pacman::p_citation() |>
                print(bibtex = FALSE) |>
                capture.output() %>%
                .[-1:-3] %>% .[. != ""] |>
                stringr::str_squish() |>
                stringr::str_replace("_", "") |>
                cat()
        cat("\n")
}
#===============================================================================
```
